### 2023-04-04 manuscript/manuscript.qmd
Add sections from guidance

---
title: Short Paper
subtitle: A Short Subtitle
author:
  - name: Alice Anonymous
    email: alice@example.com
    affiliations: 
        - id: some-tech
          name: Some Institute of Technology
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name: Bob Security
    email: bob@example.com
    affiliations:
        - id: another-u
          name: Another University
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    note: |
      Another author footnote, this is a very long footnote and it should be a really long footnote. But this footnote is not yet sufficiently long enough to make two lines of footnote text.
  - name: Cat Memes
    email: cat@example.com
    affiliations:
        - ref: another-u
    note: Yet another author footnote.
  - name: Derek Zoolander
    email: derek@example.com
    affilations:
        - ref: some-tech
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
---

# Introduction

# Material and methods

# Theory/calculation (not applicable)

# Results and Discussion

# Conclusions

# Appendix

If there is more than one appendix, they should be identified as A, B, etc. Formulae and equations in appendices should be given separate numbering: Eq. (A.1), Eq. (A.2), etc.; in a subsequent appendix, Eq. (B.1) and so on. Similarly for tables and figures: Table A.1; Fig. A.1, etc.

# Template instructions

Please make sure that your manuscript follows the guidelines in the 
Guide for Authors of the relevant journal. It is not necessary to 
typeset your manuscript in exactly the same way as an article, 
unless you are submitting to a camera-ready copy (CRC) journal.

For detailed instructions regarding the elsevier article class, see   <https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions>

# Bibliography styles

Here are two sample references:  @Feynman1963118 @Dirac1953888.

By default, natbib will be used with the `authoryear` style, set in `classoption` variable in YAML. 
You can sets extra options with `natbiboptions` variable in YAML header. Example

```
natbiboptions: longnamesfirst,angle,semicolon
```

There are various more specific bibliography styles available at
<https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>. 
To use one of these, add it in the header using, for example, `biblio-style: model1-num-names`.

## Using CSL

If `cite-method` is set to `citeproc` in `elsevier_article()`, then pandoc is used for citations instead of `natbib`. In this case, the `csl` option is used to format the references. By default, this template will provide an appropriate style, but alternative `csl` files are available from <https://www.zotero.org/styles?q=elsevier>. These can be downloaded
and stored locally, or the url can be used as in the example header.

# Equations

Here is an equation:
$$ 
  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
  \left(\frac{x}{\beta}\right)^{\alpha-1}
  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
  \alpha,\beta,x > 0 .
$$

Inline equations work as well: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$

# Figures and tables

@fig-meaningless is generated using an R chunk.

```{r}
#| label: fig-meaningless
#| fig-cap: A meaningless scatterplot
#| fig-width: 5
#| fig-height: 5
#| fig-align: center
#| out-width: 50%
#| echo: false
plot(runif(25), runif(25))
```

# Tables coming from R

Tables can also be generated using R chunks, as shown in @tbl-simple example.

```{r}
#| label: tbl-simple
#| tbl-cap: Caption centered above table
#| echo: true
knitr::kable(head(mtcars)[,1:4])
```

# References {-}

### 2023-04-04 manuscript/manuscript.tex
Initial commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  super,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Short Paper},
  pdfauthor={Alice Anonymous; Bob Security; Cat Memes; Derek Zoolander},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Short Paper \\\large{A Short Subtitle} }
\author[1]{Alice Anonymous%
\corref{cor1}%
\fnref{fn1}}
 \ead{alice@example.com} 
\author[2]{Bob Security%
%
\fnref{fn2}}
 \ead{bob@example.com} 
\author[2]{Cat Memes%
%
\fnref{fn3}}
 \ead{cat@example.com} 
\author[]{Derek Zoolander%
%
}
 \ead{derek@example.com}

\affiliation[1]{organization={Some Institute of Technology, Department
Name},addressline={Street Address},city={City},postcode={Postal
Code},postcodesep={}}
\affiliation[2]{organization={Another University, Department
Name},addressline={Street Address},city={City},postcode={Postal
Code},postcodesep={}}

\cortext[cor1]{Corresponding author}
\fntext[fn1]{This is the first author footnote.}
\fntext[fn2]{Another author footnote, this is a very long footnote and
it should be a really long footnote. But this footnote is not yet
sufficiently long enough to make two lines of footnote text.}
\fntext[fn3]{Yet another author footnote.}

\begin{abstract}
This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing
elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget
velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris
sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta
egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet
auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo
molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae.
Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at
tempus elit.
\end{abstract}

\begin{keyword}
    keyword1 \sep 
    keyword2
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, frame hidden, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi

Please make sure that your manuscript follows the guidelines in the
Guide for Authors of the relevant journal. It is not necessary to
typeset your manuscript in exactly the same way as an article, unless
you are submitting to a camera-ready copy (CRC) journal.

For detailed instructions regarding the elsevier article class, see
\url{https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions}

\hypertarget{bibliography-styles}{%
\section{Bibliography styles}\label{bibliography-styles}}

Here are two sample references: \citet{Feynman1963118}
\citet{Dirac1953888}.

By default, natbib will be used with the \texttt{authoryear} style, set
in \texttt{classoption} variable in YAML. You can sets extra options
with \texttt{natbiboptions} variable in YAML header. Example

\begin{verbatim}
natbiboptions: longnamesfirst,angle,semicolon
\end{verbatim}

There are various more specific bibliography styles available at
\url{https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files}.
To use one of these, add it in the header using, for example,
\texttt{biblio-style:\ model1-num-names}.

\hypertarget{using-csl}{%
\subsection{Using CSL}\label{using-csl}}

If \texttt{cite-method} is set to \texttt{citeproc} in
\texttt{elsevier\_article()}, then pandoc is used for citations instead
of \texttt{natbib}. In this case, the \texttt{csl} option is used to
format the references. By default, this template will provide an
appropriate style, but alternative \texttt{csl} files are available from
\url{https://www.zotero.org/styles?q=elsevier}. These can be downloaded
and stored locally, or the url can be used as in the example header.

\hypertarget{equations}{%
\section{Equations}\label{equations}}

Here is an equation: \[ 
  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
  \left(\frac{x}{\beta}\right)^{\alpha-1}
  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
  \alpha,\beta,x > 0 .
\]

Inline equations work as well: \(\sum_{i = 2}^\infty\{\alpha_i^\beta\}\)

\hypertarget{figures-and-tables}{%
\section{Figures and tables}\label{figures-and-tables}}

Figure~\ref{fig-meaningless} is generated using an R chunk.

\begin{figure}

{\centering \includegraphics[width=0.5\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-meaningless-1.pdf}

}

\caption{\label{fig-meaningless}A meaningless scatterplot}

\end{figure}

\hypertarget{tables-coming-from-r}{%
\section{Tables coming from R}\label{tables-coming-from-r}}

Tables can also be generated using R chunks, as shown in
Table~\ref{tbl-simple} example.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{knitr}\SpecialCharTok{::}\FunctionTok{kable}\NormalTok{(}\FunctionTok{head}\NormalTok{(mtcars)[,}\DecValTok{1}\SpecialCharTok{:}\DecValTok{4}\NormalTok{])}
\end{Highlighting}
\end{Shaded}

\hypertarget{tbl-simple}{}
\begin{longtable}[]{@{}lrrrr@{}}
\caption{\label{tbl-simple}Caption centered above table}\tabularnewline
\toprule()
& mpg & cyl & disp & hp \\
\midrule()
\endfirsthead
\toprule()
& mpg & cyl & disp & hp \\
\midrule()
\endhead
Mazda RX4 & 21.0 & 6 & 160 & 110 \\
Mazda RX4 Wag & 21.0 & 6 & 160 & 110 \\
Datsun 710 & 22.8 & 4 & 108 & 93 \\
Hornet 4 Drive & 21.4 & 6 & 258 & 110 \\
Hornet Sportabout & 18.7 & 8 & 360 & 175 \\
Valiant & 18.1 & 6 & 225 & 105 \\
\bottomrule()
\end{longtable}

\renewcommand\refname{References}
  \bibliography{bibliography.bib}

\end{document}

### 2023-04-04 manuscript/manuscript.qmd
Add bibliohgraphies

@@
 date: last-modified
-bibliography: bibliography.bib
+bibliography: [bibliography.bib, introduction.bib, man-qech_air-quality.bib]
 format:
@@
       cite-style: super
+editor: 
+  markdown: 
+    wrap: 72
 ---
@@
 # Introduction
+
+Built in 1964, Malawi's year of independence, Queen Elizabeth Hospital
+(QECH), or Queen's, as it's known locally, is the country's largest
+public hospital. Designed in the Tropical Modernism style of
+architecture popular within Africa in the late colonial period, Queen's
+open-plan form sprawls across a broad swath of central Blantyre, with
+dozens of wards, specialised facilities, and administrative buildings,
+linked together by a bewildering maze of covered walkways. A hive of
+around-the-clock activity, Queen's bustles from dawn to dusk with a
+constant stream of patients, drawn from across the country, and
+maintains a city-like feel even after dark as staff and stay-in family
+members cook and care for their in-patient dependants. Yet, for the
+first-time visitor, it's not the architecture or the crowds that are the
+most striking when entering the grounds, but the smell. Immediately
+noticeable, even from the road outside the hospital, Queen's is
+smothered by an acrid-smelling, white smoke, which hangs over the
+grounds day and night: the product of several constantly burning fires
+spread across the campus. Breathing this air, when a significant amount
+of burning is occurring, primarily at night when the smoke is less
+visible, can be incredibly uncomfortable. Moreover, the tropical
+architecture, designed to let air and light circulate, means that even
+indoors, within patient wards, surgery, and treatment areas, the air
+quality can also be unbearable. Poor air quality associated with
+burning, through released particulate matter (PM), has been linked to
+multiple negative health outcomes like asthma, heart disease, heart
+failure, stroke, and cancer among others (Anderson et al., 2012; Bell &
+HEI Health Review Committee, 2012). In a space of healing and recovery,
+why is such burning occurring, and how can it continue to persist?
 
@@
 
-If there is more than one appendix, they should be identified as A, B, etc.
-Formulae and equations in appendices should be given separate numbering: Eq.
-(A.1), Eq.
-(A.2), etc.; in a subsequent appendix, Eq.
-(B.1) and so on.
-Similarly for tables and figures: Table A.1; Fig.
-A.1, etc.
+If there is more than one appendix, they should be identified as A, B,
+etc. Formulae and equations in appendices should be given separate
+numbering: Eq. (A.1), Eq. (A.2), etc.; in a subsequent appendix, Eq.
+(B.1) and so on. Similarly for tables and figures: Table A.1; Fig. A.1,
+etc.
 
@@
 Guide for Authors of the relevant journal. It is not necessary to
-typeset your manuscript in exactly the same way as an article, 
-unless you are submitting to a camera-ready copy (CRC) journal.
+typeset your manuscript in exactly the same way as an article, unless
+you are submitting to a camera-ready copy (CRC) journal.
 
-For detailed instructions regarding the elsevier article class, see   <https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions>
+For detailed instructions regarding the elsevier article class, see
+<https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions>
 
@@
 
-By default, natbib will be used with the `authoryear` style, set in `classoption` variable in YAML. 
-You can sets extra options with `natbiboptions` variable in YAML header. Example
+By default, natbib will be used with the `authoryear` style, set in
+`classoption` variable in YAML. You can sets extra options with
+`natbiboptions` variable in YAML header. Example
 
@@
 <https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>.
-To use one of these, add it in the header using, for example, `biblio-style: model1-num-names`.
+To use one of these, add it in the header using, for example,
+`biblio-style: model1-num-names`.
 
@@
 
-If `cite-method` is set to `citeproc` in `elsevier_article()`, then pandoc is used for citations instead of `natbib`.
-In this case, the `csl` option is used to format the references.
-By default, this template will provide an appropriate style, but alternative `csl` files are available from <https://www.zotero.org/styles?q=elsevier>.
-These can be downloaded
-and stored locally, or the url can be used as in the example header.
+If `cite-method` is set to `citeproc` in `elsevier_article()`, then
+pandoc is used for citations instead of `natbib`. In this case, the
+`csl` option is used to format the references. By default, this template
+will provide an appropriate style, but alternative `csl` files are
+available from <https://www.zotero.org/styles?q=elsevier>. These can be
+downloaded and stored locally, or the url can be used as in the example
+header.
 
@@
 
-Here is an equation:
-$$ 
+Here is an equation: $$ 
   f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
@@
 
-Tables can also be generated using R chunks, as shown in @tbl-simple example.
+Tables can also be generated using R chunks, as shown in @tbl-simple
+example.
 
@@
 
-# References {-}
+# References {.unnumbered}

### 2023-04-05 manuscript/manuscript.qmd
Fix reference style

@@
 date: last-modified
-bibliography: [bibliography.bib, introduction.bib, man-qech_air-quality.bib]
+bibliography: [bibliography.bib, introduction.bib, man-qech-air-quality.bib]
 format:
@@
     journal:
-      name: Journal Name
-      formatting: preprint
+      name: Habitat International
+      formatting: review
       model: 3p
-      cite-style: super
+      cite-style: authoryear
+      highlights:
+        - Highlight 1
+        - Highlight 2
+        - Highlight 3
 editor: 
@@
 multiple negative health outcomes like asthma, heart disease, heart
-failure, stroke, and cancer among others (Anderson et al., 2012; Bell &
-HEI Health Review Committee, 2012). In a space of healing and recovery,
+failure, stroke, and cancer among others [@anderson2012clearing; @bell2012assessment].
+In a space of healing and recovery,
 why is such burning occurring, and how can it continue to persist?

### 2023-04-05 manuscript/manuscript.tex
Fix reference style

@@
 \documentclass[
-  super,
-  preprint,
+  authoryear,
+  review,
   3p]{elsarticle}
@@
 \makeatother
-\journal{Journal Name}
+\journal{Habitat International}
 \ifLuaTeX
@@
 \usepackage[]{natbib}
-\bibliographystyle{elsarticle-num}
+\bibliographystyle{elsarticle-harv}
 \IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
@@
 
+\begin{highlights}
+\item Highlight 1\item Highlight 2\item Highlight 3
+\end{highlights}
+
 \begin{keyword}
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, frame hidden, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable]}{\end{tcolorbox}}\fi
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+Built in 1964, Malawi's year of independence, Queen Elizabeth Hospital
+(QECH), or Queen's, as it's known locally, is the country's largest
+public hospital. Designed in the Tropical Modernism style of
+architecture popular within Africa in the late colonial period, Queen's
+open-plan form sprawls across a broad swath of central Blantyre, with
+dozens of wards, specialised facilities, and administrative buildings,
+linked together by a bewildering maze of covered walkways. A hive of
+around-the-clock activity, Queen's bustles from dawn to dusk with a
+constant stream of patients, drawn from across the country, and
+maintains a city-like feel even after dark as staff and stay-in family
+members cook and care for their in-patient dependants. Yet, for the
+first-time visitor, it's not the architecture or the crowds that are the
+most striking when entering the grounds, but the smell. Immediately
+noticeable, even from the road outside the hospital, Queen's is
+smothered by an acrid-smelling, white smoke, which hangs over the
+grounds day and night: the product of several constantly burning fires
+spread across the campus. Breathing this air, when a significant amount
+of burning is occurring, primarily at night when the smoke is less
+visible, can be incredibly uncomfortable. Moreover, the tropical
+architecture, designed to let air and light circulate, means that even
+indoors, within patient wards, surgery, and treatment areas, the air
+quality can also be unbearable. Poor air quality associated with
+burning, through released particulate matter (PM), has been linked to
+multiple negative health outcomes like asthma, heart disease, heart
+failure, stroke, and cancer among others
+\citep{anderson2012clearing, bell2012assessment}. In a space of healing
+and recovery, why is such burning occurring, and how can it continue to
+persist?
+
+\hypertarget{material-and-methods}{%
+\section{Material and methods}\label{material-and-methods}}
+
+\hypertarget{theorycalculation-not-applicable}{%
+\section{Theory/calculation (not
+applicable)}\label{theorycalculation-not-applicable}}
+
+\hypertarget{results-and-discussion}{%
+\section{Results and Discussion}\label{results-and-discussion}}
+
+\hypertarget{conclusions}{%
+\section{Conclusions}\label{conclusions}}
+
+\hypertarget{appendix}{%
+\section{Appendix}\label{appendix}}
+
+If there is more than one appendix, they should be identified as A, B,
+etc. Formulae and equations in appendices should be given separate
+numbering: Eq. (A.1), Eq. (A.2), etc.; in a subsequent appendix, Eq.
+(B.1) and so on. Similarly for tables and figures: Table A.1; Fig. A.1,
+etc.
+
+\hypertarget{template-instructions}{%
+\section{Template instructions}\label{template-instructions}}
 
@@
 \renewcommand\refname{References}
-  \bibliography{bibliography.bib}
+  \bibliography{bibliography.bib,introduction.bib,man-qech-air-quality.bib}
 

### 2023-04-05 manuscript/manuscript.qmd
Transfer introduction, update references

@@
 
+Queen’s, simply, produces an immense amount of waste that it is unable to efficiently manage.
+The dozens of wards, offices, and kitchens, serving the hundreds of patients and staff, generates considerable infectious medical waste, mixed domestic and office waste, and kitchen waste, which needs to be removed and disposed of daily.
+Some individual wards implement waste separation, binning medical wastes separately from general waste, however separation ends at the point of collection.
+Queen’s is not served by regular municipal waste collection.
+What municipal collection that is done is ad-hoc, sporadic, and based on the hospital’s available financial resources.
+Rather, most hospital waste (infectious and non-infectious) is gathered by grounds staff and burnt at the hospital’s incinerator, located at a central point within the hospital campus  [@kalina2020this].
+However, the vast majority of the deposited waste is not incinerated, as a new incinerator provided by the Ministry of Health in late 2019, is only able to handle a small percentage of the hospital’s waste.
+Furthermore, the incinerator is frequently inoperable, under repair, or without power,.
+Rather, the waste is openly burned, in a constantly smouldering pile to the side of the incinerator building, sending up the aforementioned clouds of smoke, which choke the grounds (Figure 1).
+Innumerable other, smaller, fires, scattered across the QECH campus (Figure 2) contribute their part, as grounds and maintenance staff habitually burn leaves and other garden refuse.
+Furthermore, caregivers who reside separately on hospital grounds, cook for themselves throughout the day and night using biomass (charcoal or wood), and burn their own domestic waste.
+As a result, air quality at the hospital is a constant source of discussion, with patients, staff, and caregivers struggling to cope, when just breathing is, as the quote used in the title of this article, ‘unbearable.’
+
+Challenges with solid waste management (SWM) are not unique to Queen’s, and remain persistent globally, particularly within African nations, and the Global South more broadly  [@bundhoo2018solid; @sergekubanza2022exploring; @muheirwe2022paradox; @srivastava2015urban].
+Explanations for these disparities have centred on the on waste generation outstripping waste management infrastructure [@muheirwe2022paradox], and the high costs of waste management systems, which often prove unaffordable for many low-and-middle-income countries (LMICs) [@guerrero2013solid; @sergekubanza2022exploring].
+Waste collection is an important step within SWM systems, and is a common barrier within LMIC contexts [@bundhoo2018solid].
+Waste inequalities between high-income countries (HICs) and low-income countries (LICs) is often most pernicious and visible at the point of collection, with some HICs achieving 100% collection rates, while some African nations have rates well below 50% [@shi2021unbalanced].
+Inequalities persist beyond collection, however.
+Even once waste is collected, municipalities in LMICs may burn or dump it in non-sanitary landfills due to the absence of further treatment or disposal options [@cook2021globala].
+According to @kaza2018what, 90% of waste in LIC is disposed of in unregulated dumps or openly burned, while the quantity of waste generation in these countries is expected to triple by 2050.
+As a result, open waste burning, in nations like Malawi, can be expected to become more prevalent.
+
+```{r}
+# Figure 1: Open burning at QECH, with the old incinerator building in the background (Authors).
+knitr::include_graphics(here::here("manuscript/img/image-replace-tree.png"))
+```
+
+Although hospital waste is among the common types of solid waste openly burned in LICs (in addition to municipal solid waste, sewage sludge, market or commercial waste, agricultural residues), the open burning of hospital waste has not garnered as much discussion as the burning of other waste fractions [@cook2021globala].
+The World Health Organisation (WHO) defines hospital, or health-care waste, as waste generated by health-care activities, of which approximately 85% is general, non-hazardous waste, but the remaining 15% can be considered hazardous, and may be toxic, infectious, or radioactive.
+These infectious fractions are commonly incinerated in both high and middle-income countries, however in the absence of capacity, or costly incineration or autoclave infrastructure, open burning is often a last resort option for disposal in LIC contexts [@who2018healthcare].
+The open burning of biomedical waste may reduce infection risk from potentially harmful pathogens, however, such waste may contain sharps, radioactive waste, mercury-containing instruments, and plastics, the open burning of which may result in the emission of dioxins, furans, and toxic particulate matter (PM) [@who2018healthcare; @cook2021globala].
+
+Poor air quality in in African cities can be attributed to multiple factors, including vehicular emissions, dust, or the use of low-quality fuels for cooking, lighting, and heating, however, the open burning of waste has been identified as one of the key contributors to air pollution within urban areas [@amegah2018proliferation; @bulto2020impact].
+Although open trashing burning is a problem in LICs broadly, for instance, @wiedinmyer2014global estimate that open trash burning contributes to 29% of the total global PM2.5 anthropogenic emissions, it is of particular concern in Sub-Saharan African (SSA) cities, which are home to 19 of the world’s 50 biggest dumpsites  (United Nations Development Programme, 2021).
+Particulate matter, or PM, is not specific to trash burning, but rather the product of any type of incomplete combustion and is one of the most commonly measured indicators for air quality.
+PM is measured in terms of its particle size and generally classified into PM with particle sizes less than 10μm (PM10) and those with sizes less than 2.5μm (PM2.5), while one way of measuring the severity of air pollution is by expressing the mass of PM per volume of air (μg PMx/m3 air).
+PM has been linked to multiple negative health outcomes.
+Amongst particulate air pollution, PM10 and PM2.5 are of particular concern as they can penetrate deep into the lungs, and their exposure has been associated with asthma, heart disease, heart failure, stroke, and cancer [@mohapatra2014effect; @xing2016impact].
+As a result, the health impacts of ambient air particle pollution may be significant, especially for the most vulnerable.
+For instance, the @ihme2019ambient estimates PM related air pollution resulted in 4.14 million deaths worldwide (in 2019), of which, 394,000 occurred in Africa.
+Yet, despite its importance, the air quality and health picture in Africa is worryingly incomplete.
+For instance, the WHO only collects data from 10 African countries (covering 39 cities) [@worldhealthorganization2016ambient].
+Likewise, robust, epidemiological data from and for the African context is also lacking, though emerging data suggests strong correlations between PM and poor health outcomes [@coker2018narrative; @heft-neal2018robust].
+Moving forward, it is not just the air at Queen’s which lacks clarity.
+
+Utilising a mixed-methods approach, including a network of custom-designed air quality monitors and qualitative fieldwork with hospital staff, patients, and caregivers, the purpose of this investigation was to gain a multi-dimensional understanding of the impacts of the open burning of the various solid and organic waste fractions at QECH.
+Specifically, this work aimed to intensively and longitudinally measure the air quality, in particulate matter (PM), at multiple locations at and surrounding the central burning point.
+Furthermore, we also aimed to qualitatively understand how affected individuals perceive air quality at QECH and understand potential health impacts.
+
+This research responds to a several specific gaps within the body of academic air quality literature: Although there has been ample discussion of air quality challenges in African cities [@mbow-diokhane2019air; @miralvarez2020scoping; @petkova2013particulate; @singh2021aira], there is a paucity of data on air quality impacts linked to trash burning.
+In particular, within African cities, there has been a total dearth of scholarship on the burning of medical waste, or on air quality within hospital contexts.
+Furthermore, while the voices of the citizens of Northern cities [@deguen2012new; @howel2002urban; @reames2019people] have been well documented on localised air quality issues, the voices of African urban dwellers have not been given the same consideration, and there remains a pressing need to understand how these populations experience the impacts of the open burning of trash within their communities and understand potential risks.
+
+Findings suggest that particulate matter concentrations are routinely above the WHO guidelines and are especially worrisome at several locations.
+Over the course of the two-month study, the hazardous limits for both PM10 and PM2.5 were exceeded at all locations except for the Lions Sight hospital.
+The limits were exceeded fewer than 50 times at five of the locations, but the monitors at the Lighthouse Clinic and the Guardian Shelter recorded more than 50 instances above the hazardous limits.
+The extremely hazardous air quality at the Guardian shelter is mostly from cooking over open fires, while the PM at the Lighthouse Clinic is a function of being located directly adjacent to the smouldering waste pile.
+
+Interviews show that patients, staff, and caregivers alike, are keenly aware that the air quality at Queen’s is poor, with most respondents reporting frequent respiratory-related illness.
+Many also linked the smoke to potential long-term health complications and expressed a belief that the pollution could be contributing to other, more potentially life-threatening diseases, such as asthma, cancer, and tuberculosis.
+Moreover, the tropical design of hospital buildings have rendered most coping mechanisms ineffective, with most staff only finding relief at home: relief that is not available to the hundreds of patients and caregivers who sleep at QECH or are unable to leave.
+
+In many contexts, waste burning is a necessary, but sporadic event with limited exposure.
+However, in the context of a large hospital, serving the sickest and poorest, this study shows that the consistent, and highly toxic smoke produced is an unrelenting and unnecessary burden that must be addressed, lest the patients leave in worse condition than in which they arrived.
+
 # Material and methods

### 2023-04-05 manuscript/manuscript.qmd
Narrative transfer completed

@@
 
-Built in 1964, Malawi's year of independence, Queen Elizabeth Hospital
-(QECH), or Queen's, as it's known locally, is the country's largest
-public hospital. Designed in the Tropical Modernism style of
-architecture popular within Africa in the late colonial period, Queen's
-open-plan form sprawls across a broad swath of central Blantyre, with
-dozens of wards, specialised facilities, and administrative buildings,
-linked together by a bewildering maze of covered walkways. A hive of
-around-the-clock activity, Queen's bustles from dawn to dusk with a
-constant stream of patients, drawn from across the country, and
-maintains a city-like feel even after dark as staff and stay-in family
-members cook and care for their in-patient dependants. Yet, for the
-first-time visitor, it's not the architecture or the crowds that are the
-most striking when entering the grounds, but the smell. Immediately
-noticeable, even from the road outside the hospital, Queen's is
-smothered by an acrid-smelling, white smoke, which hangs over the
-grounds day and night: the product of several constantly burning fires
-spread across the campus. Breathing this air, when a significant amount
-of burning is occurring, primarily at night when the smoke is less
-visible, can be incredibly uncomfortable. Moreover, the tropical
-architecture, designed to let air and light circulate, means that even
-indoors, within patient wards, surgery, and treatment areas, the air
-quality can also be unbearable. Poor air quality associated with
-burning, through released particulate matter (PM), has been linked to
-multiple negative health outcomes like asthma, heart disease, heart
-failure, stroke, and cancer among others [@anderson2012clearing; @bell2012assessment].
-In a space of healing and recovery,
-why is such burning occurring, and how can it continue to persist?
+Built in 1964, Malawi's year of independence, Queen Elizabeth Hospital (QECH), or Queen's, as it's known locally, is the country's largest public hospital.
+Designed in the Tropical Modernism style of architecture popular within Africa in the late colonial period, Queen's open-plan form sprawls across a broad swath of central Blantyre, with dozens of wards, specialised facilities, and administrative buildings, linked together by a bewildering maze of covered walkways.
+A hive of around-the-clock activity, Queen's bustles from dawn to dusk with a constant stream of patients, drawn from across the country, and maintains a city-like feel even after dark as staff and stay-in family members cook and care for their in-patient dependants.
+Yet, for the first-time visitor, it's not the architecture or the crowds that are the most striking when entering the grounds, but the smell.
+Immediately noticeable, even from the road outside the hospital, Queen's is smothered by an acrid-smelling, white smoke, which hangs over the grounds day and night: the product of several constantly burning fires spread across the campus.
+Breathing this air, when a significant amount of burning is occurring, primarily at night when the smoke is less visible, can be incredibly uncomfortable.
+Moreover, the tropical architecture, designed to let air and light circulate, means that even indoors, within patient wards, surgery, and treatment areas, the air quality can also be unbearable.
+Poor air quality associated with burning, through released particulate matter (PM), has been linked to multiple negative health outcomes like asthma, heart disease, heart failure, stroke, and cancer among others [@anderson2012clearing; @bell2012assessment].
+In a space of healing and recovery, why is such burning occurring, and how can it continue to persist?
 
-Queen’s, simply, produces an immense amount of waste that it is unable to efficiently manage.
+Queen's, simply, produces an immense amount of waste that it is unable to efficiently manage.
 The dozens of wards, offices, and kitchens, serving the hundreds of patients and staff, generates considerable infectious medical waste, mixed domestic and office waste, and kitchen waste, which needs to be removed and disposed of daily.
 Some individual wards implement waste separation, binning medical wastes separately from general waste, however separation ends at the point of collection.
-Queen’s is not served by regular municipal waste collection.
-What municipal collection that is done is ad-hoc, sporadic, and based on the hospital’s available financial resources.
-Rather, most hospital waste (infectious and non-infectious) is gathered by grounds staff and burnt at the hospital’s incinerator, located at a central point within the hospital campus  [@kalina2020this].
-However, the vast majority of the deposited waste is not incinerated, as a new incinerator provided by the Ministry of Health in late 2019, is only able to handle a small percentage of the hospital’s waste.
+Queen's is not served by regular municipal waste collection.
+What municipal collection that is done is ad-hoc, sporadic, and based on the hospital's available financial resources.
+Rather, most hospital waste (infectious and non-infectious) is gathered by grounds staff and burnt at the hospital's incinerator, located at a central point within the hospital campus [@kalina2020this].
+However, the vast majority of the deposited waste is not incinerated, as a new incinerator provided by the Ministry of Health in late 2019, is only able to handle a small percentage of the hospital's waste.
 Furthermore, the incinerator is frequently inoperable, under repair, or without power,.
@@
 Furthermore, caregivers who reside separately on hospital grounds, cook for themselves throughout the day and night using biomass (charcoal or wood), and burn their own domestic waste.
-As a result, air quality at the hospital is a constant source of discussion, with patients, staff, and caregivers struggling to cope, when just breathing is, as the quote used in the title of this article, ‘unbearable.’
+As a result, air quality at the hospital is a constant source of discussion, with patients, staff, and caregivers struggling to cope, when just breathing is, as the quote used in the title of this article, 'unbearable.'
 
-Challenges with solid waste management (SWM) are not unique to Queen’s, and remain persistent globally, particularly within African nations, and the Global South more broadly  [@bundhoo2018solid; @sergekubanza2022exploring; @muheirwe2022paradox; @srivastava2015urban].
+Challenges with solid waste management (SWM) are not unique to Queen's, and remain persistent globally, particularly within African nations, and the Global South more broadly [@bundhoo2018solid; @sergekubanza2022exploring; @muheirwe2022paradox; @srivastava2015urban].
 Explanations for these disparities have centred on the on waste generation outstripping waste management infrastructure [@muheirwe2022paradox], and the high costs of waste management systems, which often prove unaffordable for many low-and-middle-income countries (LMICs) [@guerrero2013solid; @sergekubanza2022exploring].
@@
 Poor air quality in in African cities can be attributed to multiple factors, including vehicular emissions, dust, or the use of low-quality fuels for cooking, lighting, and heating, however, the open burning of waste has been identified as one of the key contributors to air pollution within urban areas [@amegah2018proliferation; @bulto2020impact].
-Although open trashing burning is a problem in LICs broadly, for instance, @wiedinmyer2014global estimate that open trash burning contributes to 29% of the total global PM2.5 anthropogenic emissions, it is of particular concern in Sub-Saharan African (SSA) cities, which are home to 19 of the world’s 50 biggest dumpsites  (United Nations Development Programme, 2021).
+Although open trashing burning is a problem in LICs broadly, for instance, @wiedinmyer2014global estimate that open trash burning contributes to 29% of the total global PM2.5 anthropogenic emissions, it is of particular concern in Sub-Saharan African (SSA) cities, which are home to 19 of the world's 50 biggest dumpsites (United Nations Development Programme, 2021).
 Particulate matter, or PM, is not specific to trash burning, but rather the product of any type of incomplete combustion and is one of the most commonly measured indicators for air quality.
@@
 Likewise, robust, epidemiological data from and for the African context is also lacking, though emerging data suggests strong correlations between PM and poor health outcomes [@coker2018narrative; @heft-neal2018robust].
-Moving forward, it is not just the air at Queen’s which lacks clarity.
+Moving forward, it is not just the air at Queen's which lacks clarity.
 
@@
 
-Interviews show that patients, staff, and caregivers alike, are keenly aware that the air quality at Queen’s is poor, with most respondents reporting frequent respiratory-related illness.
+Interviews show that patients, staff, and caregivers alike, are keenly aware that the air quality at Queen's is poor, with most respondents reporting frequent respiratory-related illness.
 Many also linked the smoke to potential long-term health complications and expressed a belief that the pollution could be contributing to other, more potentially life-threatening diseases, such as asthma, cancer, and tuberculosis.
@@
 
+This study utilised a mixed-methods approach to both quantitatively measure the air quality at eight locations around QECH, and to qualitatively investigate the perceived impacts amongst staff and caregivers.
+Although the work was conducted over one sustained period in late 2019, it must be contextualised within five years' experience of research and activism within QECH by the authors.
+All relevant permissions were obtained from QECH beforehand, through consultations with administration and staff, and the research was approved by the National Committee On Research in The Social Sciences And Humanities in Malawi, Protocol NO.
+P.03/19/356.
+
+## Study site
+
+The sensors were located within the QECH campus in consultation with hospital management.
+The goal was to locate sensors across the campus with a range of distances and directions from the incinerator, though the final decision was based on accessibility, the permission of the unit's head, and convenience.
+Each of the locations is briefly described below and shown in Figure 2:
+
+-   **Lions Sight First Eye Hospital**, often shortened to **Lions Sight**, is the largest eye hospital in Malawi and serves as the main teaching eye hospital for the Kamuzu University of Health Sciences (formerly College of Medicine).
+It is staffed by 5 consultants and a team of clinical officers and nurses.
+It provides a mix of public (no cost) and private (at cost) services.
+
+-   **The Blantyre Malaria Project (BMP)**, established by Professors Terrie Taylor and Malcolm Molyneux, has carried out clinical research and patient care in the area of paediatric malaria, specifically cerebral malaria, for more than 30 years.
+Its research infrastructure includes an administrative team, an inpatient research ward, an MRI centre, and an outpatient research clinic in Ndirande township and in other districts outside of Blantyre.
+
+-   **Ward 6B** is a male ward for trauma and orthopaedic patients.
+It was established as part of the original hospital design.
+It is a 60-bed ward with a dedicated nursing station and treatment rooms.
+
+-   **The Blantyre Lighthouse Trust Clinic**, also known as **Umodzi Family Centre** or simply, the **Lighthouse Clinic**, is one of four operating across Malawi.
+The clinics work in collaboration with the Ministry of Health (MOH) to provide integrated HIV testing, treatment and care for people living with HIV.
+
+-   The **Moyo Nutritional Rehabilitation and Research Unit**, known on campus as **uMoyo**, is a 57-bed nutritional rehabilitation unit (NRU) for treating infants and children with severe malnutrition and acute illnesses.
+It is one of 104 operational NRUs in Malawi.
+It also has an outpatient therapeutic feeding program (OTP) for children with malnutrition who can be treated outside of the hospital.
+
+-   **Mercy James Institute for Pediatric Surgery and Intensive Care (MJC)**, shortened on campus to **Mercy James**, was opened in 2017 by the NGO Raising Malawi, which was founded by Madonna.
+With 3 operating rooms and 50 beds, it is the first and only first dedicated paediatric surgery hospital in Malawi.
+Prior to its opening, QECH had fewer than 10 intensive care beds.
+
+-   The **Administration Building** is the operational hub of the campus, containing mostly offices and meeting rooms.
+
+-   When a patient arrives at QECH, she or he will usually arrive with a guardian: someone to cook for them, buy medicine, do their laundry, and help them bathe.
+The **Guardian Shelter** is a gender-separated concrete floor shelter for sleeping along with a cooking pavilion and toilet/shower block are maintained by a local NGO.
+Though simple, the facility is secure and offers one of the only free/accessible toilet facilities for visitors on the QECH campus.
+
+```{r}
+# map
+# Figure 2: Sensor locations within QECH
+```
+
+## Waste management
+
+As the largest public hospital in the country, Queen's produces an immense amount of waste.
+Hospital staff, such as doctors and nurses, primarily generate and handle medical waste (dry, wet, sharps), which is disposed of in separate waste bins labelled accordingly within the ward.
+Hospital visitors (caretakers and patients) generate food waste and are advised by staff to discard such waste in waste bins outside, on hospital grounds.
+Cleaning staff are responsible for emptying bins and maintaining the cleanliness of indoor and outdoor spaces.
+Collected bags are temporarily stored in the sluice room, and cleaners during the night shift are responsible for transporting the waste on wheelchairs or trolleys to the incinerator for burning by waste disposers.
+
+Yet, according to respondents and our own observations, waste management procedures within the hospital are fraught with challenges which impact the implementation of best practices.
+These include the lack of a coherent hospital-wide waste management policy, insufficient waste management training for staff, and a persistent lack of material and financial resources.
+For instance, although hospital administration was able to articulate a set of standard operating principles (SoPs) for waste management, below the administration, doctors and nurses were generally limited in their understanding about its contents.
+Furthermore, cleaners, who are responsible for day-to-day cleansing and waste disposal, yet rate at the bottom of the institutional hierarchy at Queens, were unable to access or analyse hospital policy, and generally tended to rely on "experience" and "human judgement" when handling potentially hazardous waste.
+Moreover, although nearly all doctors and nurses had prior education on waste management best practices, cleaners began their waste handling duties without any prior waste management experience, and although there is an orientation programme for new hires, there is no regular refresher training or capacity building regarding waste for existing staff, at any level.
+
+Despite its national prominence, and international reputation for academic medical research, Queen's is plagued by constant shortages of human and material resources.
+Yet, these shortages are not shared equally across the space, with a sharp division in cleanliness, resources, and quality of care between public wards, supported by the Malawi Ministry of Health, and the private wads with international funding, such as the aforementioned Mercy James Institute.
+For staff, these shortages have manifested in poor salaries, missed pay cheques, failing infrastructure and limited supplies [@kalina2020this].
+Regarding waste management, these shortages have manifested in a chronic shortage of black bags, waste containers and cleaning staff.
+As a result, due to a lack of bags and storage containers, there is minimal segregation of waste, and highly infectious items (bandages, gloves, syringes, and other sharps) are disposed of along with non-hazardous waste (food, newspapers, packaging).
+As a result, rather than facilitating different disposal pathways for hazardous and non-hazardous waste, all waste is mixed, necessitating more burning than would be necessary if non-hazardous waste was diverted, and contributing to workplace risk, with reports of accidental exposure to potentially infected waste, including accidental pokes from improperly disposed of sharps, common amongst hospital staff.
+Regardless, separation ends at the point of collection.
+The hospital is not serviced by regular municipal waste collection, and as a result, nearly all of the waste produced by the hospital, hazardous and non-hazardous, has to be disposed of on-site, either within the limited capacity of the hospital incinerator, or more commonly, through open burning.
+As a result, the constantly smouldering pile of waste puts forth a continuous cloud of grey smoke, which mingling with the dozens of other fires on the grounds, from cooking fires and burning garden waste, blankets Queen's in a permanent cloud of foul smelling haze.
+
+## Measuring Air Quality
+
+### Hardware and Software
+
+The PM monitoring device centres around a Raspberry Pi 3 Model B single-board computer complete with an operating system and storage space.
+The added advantage of storage space is that the data are not lost if the internet connection breaks down.
+Because the Pi can only read digital signals, we needed to include an analog-to-digital (ADC) converter.
+The amount of dust is measured by a Nova PM SDS011 High Precision Laser sensor.
+This sensor measures particles at 2.5 and 10 micrometres in diameter in mg/m3.
+A list of the specific hardware and software components can be found in Appendix TODO.
+
+### Installation
+
+The choice of where the sensors were installed was done in collaboration with the QECH administration, but the position of the sensors at the individual buildings were decided by the research team.
+Four air quality sensors were installed on the outside of buildings (Guardian Shelter, Mercy James, Malaria Project, Lighthouse Clinic) while the remaining four sensors were installed inside (Ward 6B, Administration, Lions Sight, uMoyo).
+All of the sensors were mounted to a wall at a height of approximately two metres from the ground with the help of QECH maintenance department personnel.
+This height was chosen to prevent the public from tampering with the equipment when they visit the hospital.
+Additionally, we assured that each pipe that inhaled ambient air was freely protruding in the building (room) in order to capture the air quality.
+
+Because connecting each unit to a power supply was not possible, each unit was equipped with an external battery (5000mAh) which powered the sensor unit for three days continuously.
+To avoid any down time, we changed the batteries every two days.
+Data were also collected from the sensor at the time of battery replacement.
+The sensors were capable of being connected to WIFI, a feature which enabled us to download the data into our computers wirelessly.
+
+### Qualitative Methods
+
+Qualitative data collection consisted of 26 interviews with caregivers and hospital staff (including janitorial and maintenance staff, nurses, doctors, and administrators) conducted in October and November 2019, as well as an additional 31 interviews conducted with staff, patients, and caregivers in January and February 2021.
+Interviews were semi-structured, and included questions on hospital waste management practices, perceived best practices, and workplace health, safety, and risk.
+However, the semi-structured nature of the interviews allowed for the participants to raise topics of interest, and for unexpected themes to emerge.
+These interviews are supplemented by participatory observation recorded within detailed field notes, and supported by nearly a decade of sustained involvement and activism by the authors in the waste management practices of the hospital.
+
+Interview respondents were chosen using a purposive or judgement sampling regimen, i.e.
+a subjective sampling method in which respondents are selected based on their ability to effectively contribute to the study's research objectives [@kitchin2013conducting].
+All respondents were purposively selected based on availability, willingness to provide written, informed consent, and their individual insight into the questions posed within the study.
+Interviews were conducted in the local language (Chichewa), audio recorded, and transcribed into English.
+Participation was voluntary, and responses were recorded anonymously.
+Written, informed consent was obtained from each respondent prior to the interview.
+The study was approved by the National Committee on Research in the Social Sciences and Humanities (NCRSH) of Malawi; Protocol NO.
+P.03/19/356.
+Collected data were analysed thematically and was coded within the software programme Nvivo, which organises materials and assists with the coding process.
+
+### Limitations
+
+Though extensive and useful in understanding the conditions at QECH, the data set would benefit from a comparison with background values.
+As trash-burning is extensive throughout the city, there were no obvious locations that could be confidently used as representative background values.
+And although there were several possible areas upwind and outside of town, the logistics and safety involved in accessing them to change the batteries meant that the measured values at Queens could not be compared to a stable background concentration.
+Knowing the wind direction and velocity would have helped in better identifying the potential sources of the measured PM.
+However, the necessary equipment was not easily available in Blantyre at the time, and as an exploratory study, the immediate hazards for the staff and patients at QECH are relevant regardless of the burning source.
+
+Knowing the exact burning locations and burning times, especially at the incinerator and the Guardian Shelter, would have helped in better identifying the sources and movements of the plumes.
+However, given the size of the campus and it's 24-hour schedule, a much larger research team would have been required to quantify all the burning.
+Finally, The humidity, and the potential for interference was not accounted for; the measurements were not adjusted for humidity which could affect the values, though not enough to significantly alter the key findings and need for immediate action.
+
 # Theory/calculation (not applicable)
@@
 # Results and Discussion
+
+The Malawi Bureau of Standards is the national agency responsible for setting and publishing all standards in the country.
+At the time of writing, the official website (www.mbsmw.org) was not available.
+However, published work that references Malawian standards [@kutlarjoss2017time; @mapoma2014air] indicate that the maximum 24-h PM10 value is 25 μg/m3 and the maximum annual PM2.5 value is 8 μg/m3.
+There is no daily maximum value for PM2.5.
+It should be noted that both of these values are at, or below, the level-4 interim WHO targets.
+
+Therefore, for international comparisons and due to the lack of comprehensive Malawian Standards, we use the WHO recommendations as a basis of comparison for the measured values [@worldhealthorganization2021who].
+The relevant particulate matter targets are presented in Table 1.
+
+```{r}
+# Table 1. Particulate matter targets from @worldhealthorganization2021who
+knitr::include_graphics(here::here("img/image-replace-table.png"))
+
+```
+
+## Particulate Measurements
+
+### Peaks
+
+Due to localised variation (air movement in the immediate area) and the density of the observations, the plotted 5-minute data obscure persistent trends.
+The full set of 5-minute data are plotted and presented in Appendix XA1 (TODO).
+However, the number of measured values that exceeded the hazardous limit for both parameters are summarised in Table 2.
+
+```{r}
+# Table 2.
+Peaks for both PM2.5 and PM10 are the number of data points above the WHO interim target 1 (annual) which is the least stringent;  the number of observations recorded (count) is provided for reference.
+
+```
+
+Over the course of the two-month study, the hazardous limits for both parameters were exceeded at all locations except for the Lions Sight hospital.
+The limits were exceeded fewer than 50 times at five of the locations, and only the monitors at the Lighthouse Clinic and the Guardian Shelter recorded more than 50 instances above the hazardous limits.
+At the Lighthouse Clinic, the PM10 limit was exceeded almost twice as often as the PM2.5 limit, while at the Guardian Shelter, the reverse is true.
+Because the Lighthouse Clinic is within 50 m of the incinerator, the number and predominance of PM10 peaks is characteristic of incomplete combustion and dust that is typical for the area and the open burning that occurs there.
+The residents at the Guardian Shelter however are mostly cooking within contained clay stoves that have been designed to improve combustion of the coal and wood that is burned within them.
+The fewer number of PM10 peaks is a testament to this intervention, though the frequency of PM2.5 peaks is still beyond acceptable.
+
+### 24-hour averages
+
+The 24-hour averaged values (logged every 5 minutes) are presented for each location, for both PM10 and PM2.5 in Figure 3.
+As well as dampening local variation and peaks, the WHO air quality guidelines are also based on 24-hour averages which are the standard against which the health risks can be judged.
+
+Overall, PM2.5 values remained below 100 μg/m3 at 6 of 8 locations (Administration, Lions Sight, Malaria Project, uMoyo, Ward 6B, and Mercy James); PM10 values were consistently below 100μg/m3 at the same locations, but with several average values extending slightly above, and then infrequently.
+
+The daily averages at both the Lighthouse Clinic and the Guardian Shelter are both consistently higher for both parameters and the two averages closely followed the same general trends.
+Though the peaks at the Lighthouse Clinic were higher than the Guardian Shelter, the low values were consistently lower, indicating more times of little or no burning, unlike the Guardian Shelter emissions which were relatively constant.
+However, further analysis of 12 hour averages (8:00-15:59 (working hours) and 16:00-7:59 (evening)) did not indicate clear differences between the time periods; stated differently, the data did not clearly point to more or less burning in the day or at night.
+
+```{r}
+# Figure 3. Average 24-hour PM10 and PM2.5 at 8 monitoring stations over 2 months
+
+```
+
+### PM ratios
+
+Given that the type of fuel and the type of burning (contained vs.
+open) produce very different particulate "fingerprints", the ratio of PM10 to PM2.5 values were examined to determine if a clear difference between locations, and therefore source, could be identified.
+The results are presented in Figure 4.
+
+```{r}
+# Figure 4. PM10:PM2.5 values for each location.  Each line shows the minimum and maximum ratio for the day based on ratios calculated hourly.  
+```
+
+The results presented in Figure 5 illustrate both the presence of PM10 relative to PM2.5 as well as the range of values over which that ratio is observed.
+
+Unlike the peak values presented in Table 2, the values presented in Figure 5 indicate the relative presence of larger (PM10) particles to smaller ones (PM2.5) regardless of their concentration: the ratio of two low concentrations can be the same as the ratio of two high concentrations and is therefore more indicative of source than of proximity.
+
+In general, the calculated ratios are concentrated between 1.25 with few values exceeding 7.5.
+Three from each the Guardian Shelter and the Ward 6B exceeded 7.5 though the most values exceeding 5 were at the Guardian Shelter.
+It is interesting to note that despite the large daily variations at the Lighthouse Clinic, the parameter ratios there were fairly consistent with few spikes, and only 2 days with ratios above 5.
+
+An examination of the range of ratios is used to better understand the variability within a single day.
+For example, several of the calculated values at the Guardian Shelter have a range of more than 5, and one day had a range of close to 10 i.e.
+the minimum ratio recorded for that day was close to 1 and the maximum value was over 10.
+The composition of emissions recorded at that location ranged significantly, and likely reflects a range of burning styles and/or fuel type.
+
+## Exposure Data
+
+The relative amount of time exposed to a given category of air quality is shown in Figure 5. Each square represents 1% of the measured time.
+
+```{r}
+# Figure 5. Percentage of measured values categorised according to WHO 2021 targets
+```
+
+Unsurprisingly, a majority of the air quality at the Lighthouse Clinic and the Guardian Shelter falls at, or below, interim target 1.
+What is less expected is that even though there are fewer days of hazardous or very bad air quality at the other locations, most locations do not have target 4 values for even half of the time.
+Mercy James, the Malaria Project and Lions Sight have the largest percentages of days with air quality meeting the interim target 4.
+Interim target 2 rated air occupies an unexpectedly large proportion of the days at all locations, indicating that although there are a few very toxic, and a few quite good days, the majority of the air is actually neither, but still cause for concern.
+
+## "The Air We Breathe is Not Good": Perspectives from QECH
+
+As the previous section has described, the open incineration of waste at QECH has created hazardous conditions for those occupying the space.
+These risks were not lost on staff and caregivers, as interviews demonstrated broad and near universal awareness of the poor air quality within the hospital grounds.
+Overwhelmingly, within interviews, both caregivers and staff were quick to decry the poor quality of the air, generally without prompting.
+The few exceptions were those staff posted on the peripheries of the hospital grounds, at a distance from the spaces of incineration.
+However, even those who did not experience the impacts of the burning, were aware of it, and considered themselves fortunate to be posted in a section of the hospital where it was less of a problem.
+Furthermore, according to caregivers and staff who work night shifts, air quality can be particularly bad at night and in the early morning, despite the smoke being less visible, due to the habit of janitorial staff concentrating their burning during the late hours.
+The poor air quality on hospital grounds is also a frequent cause for complaint by patients and visitors, with nearly every staff member interviewed being able to recall having received a complaint, and in turn, complaining to administration.
+One of the staff members responsible for the burning said that he personally, had received hundreds of complaints, but was powerless to affect meaningful change, aside from burning at different hours, until the incinerator could be repaired.
+
+Nonetheless, despite this consensus that air quality caused by the burning of waste within the hospital was a problem, significant differences emerged between respondents over their understandings of potential impacts, the effectiveness of various coping mechanisms, and their problematisations linked to the burning of specific waste materials.
+In addition, interviews revealed that these understandings were informed by a significant amount of misinformation, even amongst trained medical staff, which may contribute to them being less able to mitigate potential risks for themselves and those who rely on their care.
+
+### Impacts, Problematisations, and Misinformation
+
+The poor air quality within QECH was responsible, according to respondents, for a wide array of health impacts.
+The most common ones cited included: coughing and sneezing, sore throat, stinging eyes, breathing difficulties, and persistent cold and flu.
+Nausea was also mentioned, but was not a commonly described impact.
+Only one respondent, of the 26 total interviewed, did not describe lingering health impacts which they could ascribe to the smoke, however, they also described having chronic eye irritation, but did not believe the smoke was a contributing factor.
+
+In addition to these impacts, which respondents bear on a daily basis, many also believed that the smoke could contribute to a number of more serious, long-term health complications.
+For instance, nearly a quarter of respondents raised concerns of the potential impact that the smoke could have on patients or staff with asthma.
+Others flagged poor air quality as a potential risk factor for certain cancers, lung disease, or heart problems.
+For a few, the smoke posed an unknown danger, they were not sure what types of impacts it could have, but they were sure it was harmful in some way.
+
+Also, understandably, given the large tuberculosis ward present within the hospital grounds, and the high prevalence of the disease within Malawi [@worldhealthorganization2019global] there was significant concern (more than half of respondents) about the impact the poor air quality could have on the infected.
+However, there also persisted a belief among several respondents, including several nurses, that the smoke could be a cause of the disease itself.
+As one staff member (20/11/2019) stated, "I believe breathing this air for a long time can cause Tuberculosis." This, however, was only one of the few instances of misinformation which staff members held regarding air quality and health.
+Another example, voiced by several respondents, included a belief that some staff members were immune to ill-health impacts of the smoke, because they had received vaccinations from the hospital.
+Nonetheless, they were concerned about the impacts of the smoke on patients and visitors, as one staff member (18/11/2019) expressed:
+
+> Personally, I have never experienced \[eye discomfort\] because I get vaccinated and I am protected including other staff.
+However, we realize that the air can badly affect other people and patients who come to this place.
+
+Another interesting misconception that emerged, which may be tied partly to translation and transcription, was a different cultural understanding of smoke versus smells.
+More than half of the respondents appeared to conflate the two, with some expressing a belief that it was the odour of what was being burnt that was harmful, rather than the smoke being given off.
+This has led many to specifically problematise the burning of certain wastes, such as plastics, medicines, and other medical wastes, which give off distinctive or less pleasant odours, as opposed to the burning of other items, such as garden refuse, which may produce significant smoke, and contribute to higher recorded values of particulate matter, but produce a less pungent, or more normalised, odour.
+
+### Coping Mechanisms
+
+Finally, in order to manage the impacts of QECH's persistently poor air quality, staff and visitors reported having developed a number of coping mechanisms, designed to help them get through their daily routines.
+These included staying indoors, blocking doors and windows, and taking breaks away from hospital grounds in order to catch some breaths of fresher air.
+However, for janitorial staff, inside was not necessarily better, as several reported their indoor workspaces as being also insufferable for long periods of time from the smell of cleaning agents and other chemicals.
+Furthermore, amongst respondents there was a general disagreement over the effectiveness of personal protective equipment (PPE), such as face masks, towards mitigating the impacts of the smoke.
+A few staff described pleading to hospital administration for such equipment, to no avail.
+However, other staff members, who do have access to PPE, noted that even face masks do little to mitigate the impacts of the smoke, describing them as ineffective.
+
+Most staff, however, have been unable to find anyway to mitigate the impacts of the smoke, and only found relief once they reached home at the end of their shift, as one of the janitorial staff described, "we only feel safe when we are home".
+Of course, this relief is not available for the hundreds of patients and caregivers who sleep at QECH or are unable to leave.
+Ultimately, most place their hope in the construction of the new incinerator (which had not yet been activated at the time of the interviews), and biding their time as construction drags on; coping as best they can.
+As one staff member (25/11/2019) described,"we are just hoping we will start breathing good air soon, when the new incinerator is opened."
 
@@
 
-If there is more than one appendix, they should be identified as A, B,
-etc. Formulae and equations in appendices should be given separate
-numbering: Eq. (A.1), Eq. (A.2), etc.; in a subsequent appendix, Eq.
-(B.1) and so on. Similarly for tables and figures: Table A.1; Fig. A.1,
-etc.
-
-# Template instructions
-
-Please make sure that your manuscript follows the guidelines in the
-Guide for Authors of the relevant journal. It is not necessary to
-typeset your manuscript in exactly the same way as an article, unless
-you are submitting to a camera-ready copy (CRC) journal.
-
-For detailed instructions regarding the elsevier article class, see
-<https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions>
-
-# Bibliography styles
-
-Here are two sample references: @Feynman1963118 @Dirac1953888.
-
-By default, natbib will be used with the `authoryear` style, set in
-`classoption` variable in YAML. You can sets extra options with
-`natbiboptions` variable in YAML header. Example
-
-```         
-natbiboptions: longnamesfirst,angle,semicolon
-```
-
-There are various more specific bibliography styles available at
-<https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>.
-To use one of these, add it in the header using, for example,
-`biblio-style: model1-num-names`.
-
-## Using CSL
-
-If `cite-method` is set to `citeproc` in `elsevier_article()`, then
-pandoc is used for citations instead of `natbib`. In this case, the
-`csl` option is used to format the references. By default, this template
-will provide an appropriate style, but alternative `csl` files are
-available from <https://www.zotero.org/styles?q=elsevier>. These can be
-downloaded and stored locally, or the url can be used as in the example
-header.
-
-# Equations
-
-Here is an equation: $$ 
-  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
-  \left(\frac{x}{\beta}\right)^{\alpha-1}
-  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
-  \alpha,\beta,x > 0 .
-$$
-
-Inline equations work as well: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$
+If there is more than one appendix, they should be identified as A, B, etc.
+Formulae and equations in appendices should be given separate numbering: Eq.
+(A.1), Eq.
+(A.2), etc.; in a subsequent appendix, Eq.
+(B.1) and so on.
+Similarly for tables and figures: Table A.1; Fig.
+A.1, etc.
 
@@
 
-Tables can also be generated using R chunks, as shown in @tbl-simple
-example.
+Tables can also be generated using R chunks, as shown in @tbl-simple example.
 

### 2023-04-05 manuscript/manuscript.qmd
Fix YAML
Finalise complete reference list

@@
 date: last-modified
-bibliography: [bibliography.bib, introduction.bib, man-qech-air-quality.bib]
+bibliography: man-qechairquality.bib
 format:
@@
 # Table 1. Particulate matter targets from @worldhealthorganization2021who
-knitr::include_graphics(here::here("img/image-replace-table.png"))
+knitr::include_graphics(here::here("manuscript/img/image-replace-table.png"))
 

### 2023-04-05 manuscript/manuscript.tex
Finalise complete reference list

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, boxrule=0pt, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 
+Queen's, simply, produces an immense amount of waste that it is unable
+to efficiently manage. The dozens of wards, offices, and kitchens,
+serving the hundreds of patients and staff, generates considerable
+infectious medical waste, mixed domestic and office waste, and kitchen
+waste, which needs to be removed and disposed of daily. Some individual
+wards implement waste separation, binning medical wastes separately from
+general waste, however separation ends at the point of collection.
+Queen's is not served by regular municipal waste collection. What
+municipal collection that is done is ad-hoc, sporadic, and based on the
+hospital's available financial resources. Rather, most hospital waste
+(infectious and non-infectious) is gathered by grounds staff and burnt
+at the hospital's incinerator, located at a central point within the
+hospital campus \citep{kalina2020this}. However, the vast majority of
+the deposited waste is not incinerated, as a new incinerator provided by
+the Ministry of Health in late 2019, is only able to handle a small
+percentage of the hospital's waste. Furthermore, the incinerator is
+frequently inoperable, under repair, or without power,. Rather, the
+waste is openly burned, in a constantly smouldering pile to the side of
+the incinerator building, sending up the aforementioned clouds of smoke,
+which choke the grounds (Figure 1). Innumerable other, smaller, fires,
+scattered across the QECH campus (Figure 2) contribute their part, as
+grounds and maintenance staff habitually burn leaves and other garden
+refuse. Furthermore, caregivers who reside separately on hospital
+grounds, cook for themselves throughout the day and night using biomass
+(charcoal or wood), and burn their own domestic waste. As a result, air
+quality at the hospital is a constant source of discussion, with
+patients, staff, and caregivers struggling to cope, when just breathing
+is, as the quote used in the title of this article, `unbearable.'
+
+Challenges with solid waste management (SWM) are not unique to Queen's,
+and remain persistent globally, particularly within African nations, and
+the Global South more broadly
+\citep{bundhoo2018solid, sergekubanza2022exploring, muheirwe2022paradox, srivastava2015urban}.
+Explanations for these disparities have centred on the on waste
+generation outstripping waste management infrastructure
+\citep{muheirwe2022paradox}, and the high costs of waste management
+systems, which often prove unaffordable for many low-and-middle-income
+countries (LMICs) \citep{guerrero2013solid, sergekubanza2022exploring}.
+Waste collection is an important step within SWM systems, and is a
+common barrier within LMIC contexts \citep{bundhoo2018solid}. Waste
+inequalities between high-income countries (HICs) and low-income
+countries (LICs) is often most pernicious and visible at the point of
+collection, with some HICs achieving 100\% collection rates, while some
+African nations have rates well below 50\% \citep{shi2021unbalanced}.
+Inequalities persist beyond collection, however. Even once waste is
+collected, municipalities in LMICs may burn or dump it in non-sanitary
+landfills due to the absence of further treatment or disposal options
+\citep{cook2021globala}. According to \citet{kaza2018what}, 90\% of
+waste in LIC is disposed of in unregulated dumps or openly burned, while
+the quantity of waste generation in these countries is expected to
+triple by 2050. As a result, open waste burning, in nations like Malawi,
+can be expected to become more prevalent.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Figure 1: Open burning at QECH, with the old incinerator building in the background (Authors).}
+\NormalTok{knitr}\SpecialCharTok{::}\FunctionTok{include\_graphics}\NormalTok{(here}\SpecialCharTok{::}\FunctionTok{here}\NormalTok{(}\StringTok{"manuscript/img/image{-}replace{-}tree.png"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{img/image-replace-tree.png}
+
+}
+
+\end{figure}
+
+Although hospital waste is among the common types of solid waste openly
+burned in LICs (in addition to municipal solid waste, sewage sludge,
+market or commercial waste, agricultural residues), the open burning of
+hospital waste has not garnered as much discussion as the burning of
+other waste fractions \citep{cook2021globala}. The World Health
+Organisation (WHO) defines hospital, or health-care waste, as waste
+generated by health-care activities, of which approximately 85\% is
+general, non-hazardous waste, but the remaining 15\% can be considered
+hazardous, and may be toxic, infectious, or radioactive. These
+infectious fractions are commonly incinerated in both high and
+middle-income countries, however in the absence of capacity, or costly
+incineration or autoclave infrastructure, open burning is often a last
+resort option for disposal in LIC contexts \citep{who2018healthcare}.
+The open burning of biomedical waste may reduce infection risk from
+potentially harmful pathogens, however, such waste may contain sharps,
+radioactive waste, mercury-containing instruments, and plastics, the
+open burning of which may result in the emission of dioxins, furans, and
+toxic particulate matter (PM)
+\citep{who2018healthcare, cook2021globala}.
+
+Poor air quality in in African cities can be attributed to multiple
+factors, including vehicular emissions, dust, or the use of low-quality
+fuels for cooking, lighting, and heating, however, the open burning of
+waste has been identified as one of the key contributors to air
+pollution within urban areas
+\citep{amegah2018proliferation, bulto2020impact}. Although open trashing
+burning is a problem in LICs broadly, for instance,
+\citet{wiedinmyer2014global} estimate that open trash burning
+contributes to 29\% of the total global PM2.5 anthropogenic emissions,
+it is of particular concern in Sub-Saharan African (SSA) cities, which
+are home to 19 of the world's 50 biggest dumpsites (United Nations
+Development Programme, 2021). Particulate matter, or PM, is not specific
+to trash burning, but rather the product of any type of incomplete
+combustion and is one of the most commonly measured indicators for air
+quality. PM is measured in terms of its particle size and generally
+classified into PM with particle sizes less than 10μm (PM10) and those
+with sizes less than 2.5μm (PM2.5), while one way of measuring the
+severity of air pollution is by expressing the mass of PM per volume of
+air (μg PMx/m3 air). PM has been linked to multiple negative health
+outcomes. Amongst particulate air pollution, PM10 and PM2.5 are of
+particular concern as they can penetrate deep into the lungs, and their
+exposure has been associated with asthma, heart disease, heart failure,
+stroke, and cancer \citep{mohapatra2014effect, xing2016impact}. As a
+result, the health impacts of ambient air particle pollution may be
+significant, especially for the most vulnerable. For instance, the
+\citet{ihme2019ambient} estimates PM related air pollution resulted in
+4.14 million deaths worldwide (in 2019), of which, 394,000 occurred in
+Africa. Yet, despite its importance, the air quality and health picture
+in Africa is worryingly incomplete. For instance, the WHO only collects
+data from 10 African countries (covering 39 cities)
+\citep{worldhealthorganization2016ambient}. Likewise, robust,
+epidemiological data from and for the African context is also lacking,
+though emerging data suggests strong correlations between PM and poor
+health outcomes \citep{coker2018narrative, heft-neal2018robust}. Moving
+forward, it is not just the air at Queen's which lacks clarity.
+
+Utilising a mixed-methods approach, including a network of
+custom-designed air quality monitors and qualitative fieldwork with
+hospital staff, patients, and caregivers, the purpose of this
+investigation was to gain a multi-dimensional understanding of the
+impacts of the open burning of the various solid and organic waste
+fractions at QECH. Specifically, this work aimed to intensively and
+longitudinally measure the air quality, in particulate matter (PM), at
+multiple locations at and surrounding the central burning point.
+Furthermore, we also aimed to qualitatively understand how affected
+individuals perceive air quality at QECH and understand potential health
+impacts.
+
+This research responds to a several specific gaps within the body of
+academic air quality literature: Although there has been ample
+discussion of air quality challenges in African cities
+\citep{mbow-diokhane2019air, miralvarez2020scoping, petkova2013particulate, singh2021aira},
+there is a paucity of data on air quality impacts linked to trash
+burning. In particular, within African cities, there has been a total
+dearth of scholarship on the burning of medical waste, or on air quality
+within hospital contexts. Furthermore, while the voices of the citizens
+of Northern cities
+\citep{deguen2012new, howel2002urban, reames2019people} have been well
+documented on localised air quality issues, the voices of African urban
+dwellers have not been given the same consideration, and there remains a
+pressing need to understand how these populations experience the impacts
+of the open burning of trash within their communities and understand
+potential risks.
+
+Findings suggest that particulate matter concentrations are routinely
+above the WHO guidelines and are especially worrisome at several
+locations. Over the course of the two-month study, the hazardous limits
+for both PM10 and PM2.5 were exceeded at all locations except for the
+Lions Sight hospital. The limits were exceeded fewer than 50 times at
+five of the locations, but the monitors at the Lighthouse Clinic and the
+Guardian Shelter recorded more than 50 instances above the hazardous
+limits. The extremely hazardous air quality at the Guardian shelter is
+mostly from cooking over open fires, while the PM at the Lighthouse
+Clinic is a function of being located directly adjacent to the
+smouldering waste pile.
+
+Interviews show that patients, staff, and caregivers alike, are keenly
+aware that the air quality at Queen's is poor, with most respondents
+reporting frequent respiratory-related illness. Many also linked the
+smoke to potential long-term health complications and expressed a belief
+that the pollution could be contributing to other, more potentially
+life-threatening diseases, such as asthma, cancer, and tuberculosis.
+Moreover, the tropical design of hospital buildings have rendered most
+coping mechanisms ineffective, with most staff only finding relief at
+home: relief that is not available to the hundreds of patients and
+caregivers who sleep at QECH or are unable to leave.
+
+In many contexts, waste burning is a necessary, but sporadic event with
+limited exposure. However, in the context of a large hospital, serving
+the sickest and poorest, this study shows that the consistent, and
+highly toxic smoke produced is an unrelenting and unnecessary burden
+that must be addressed, lest the patients leave in worse condition than
+in which they arrived.
+
 \hypertarget{material-and-methods}{%
 \section{Material and methods}\label{material-and-methods}}
+
+This study utilised a mixed-methods approach to both quantitatively
+measure the air quality at eight locations around QECH, and to
+qualitatively investigate the perceived impacts amongst staff and
+caregivers. Although the work was conducted over one sustained period in
+late 2019, it must be contextualised within five years' experience of
+research and activism within QECH by the authors. All relevant
+permissions were obtained from QECH beforehand, through consultations
+with administration and staff, and the research was approved by the
+National Committee On Research in The Social Sciences And Humanities in
+Malawi, Protocol NO. P.03/19/356.
+
+\hypertarget{study-site}{%
+\subsection{Study site}\label{study-site}}
+
+The sensors were located within the QECH campus in consultation with
+hospital management. The goal was to locate sensors across the campus
+with a range of distances and directions from the incinerator, though
+the final decision was based on accessibility, the permission of the
+unit's head, and convenience. Each of the locations is briefly described
+below and shown in Figure 2:
+
+\begin{itemize}
+\item
+  \textbf{Lions Sight First Eye Hospital}, often shortened to
+  \textbf{Lions Sight}, is the largest eye hospital in Malawi and serves
+  as the main teaching eye hospital for the Kamuzu University of Health
+  Sciences (formerly College of Medicine). It is staffed by 5
+  consultants and a team of clinical officers and nurses. It provides a
+  mix of public (no cost) and private (at cost) services.
+\item
+  \textbf{The Blantyre Malaria Project (BMP)}, established by Professors
+  Terrie Taylor and Malcolm Molyneux, has carried out clinical research
+  and patient care in the area of paediatric malaria, specifically
+  cerebral malaria, for more than 30 years. Its research infrastructure
+  includes an administrative team, an inpatient research ward, an MRI
+  centre, and an outpatient research clinic in Ndirande township and in
+  other districts outside of Blantyre.
+\item
+  \textbf{Ward 6B} is a male ward for trauma and orthopaedic patients.
+  It was established as part of the original hospital design. It is a
+  60-bed ward with a dedicated nursing station and treatment rooms.
+\item
+  \textbf{The Blantyre Lighthouse Trust Clinic}, also known as
+  \textbf{Umodzi Family Centre} or simply, the \textbf{Lighthouse
+  Clinic}, is one of four operating across Malawi. The clinics work in
+  collaboration with the Ministry of Health (MOH) to provide integrated
+  HIV testing, treatment and care for people living with HIV.
+\item
+  The \textbf{Moyo Nutritional Rehabilitation and Research Unit}, known
+  on campus as \textbf{uMoyo}, is a 57-bed nutritional rehabilitation
+  unit (NRU) for treating infants and children with severe malnutrition
+  and acute illnesses. It is one of 104 operational NRUs in Malawi. It
+  also has an outpatient therapeutic feeding program (OTP) for children
+  with malnutrition who can be treated outside of the hospital.
+\item
+  \textbf{Mercy James Institute for Pediatric Surgery and Intensive Care
+  (MJC)}, shortened on campus to \textbf{Mercy James}, was opened in
+  2017 by the NGO Raising Malawi, which was founded by Madonna. With 3
+  operating rooms and 50 beds, it is the first and only first dedicated
+  paediatric surgery hospital in Malawi. Prior to its opening, QECH had
+  fewer than 10 intensive care beds.
+\item
+  The \textbf{Administration Building} is the operational hub of the
+  campus, containing mostly offices and meeting rooms.
+\item
+  When a patient arrives at QECH, she or he will usually arrive with a
+  guardian: someone to cook for them, buy medicine, do their laundry,
+  and help them bathe. The \textbf{Guardian Shelter} is a
+  gender-separated concrete floor shelter for sleeping along with a
+  cooking pavilion and toilet/shower block are maintained by a local
+  NGO. Though simple, the facility is secure and offers one of the only
+  free/accessible toilet facilities for visitors on the QECH campus.
+\end{itemize}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# map}
+\CommentTok{\# Figure 2: Sensor locations within QECH}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{waste-management}{%
+\subsection{Waste management}\label{waste-management}}
+
+As the largest public hospital in the country, Queen's produces an
+immense amount of waste. Hospital staff, such as doctors and nurses,
+primarily generate and handle medical waste (dry, wet, sharps), which is
+disposed of in separate waste bins labelled accordingly within the ward.
+Hospital visitors (caretakers and patients) generate food waste and are
+advised by staff to discard such waste in waste bins outside, on
+hospital grounds. Cleaning staff are responsible for emptying bins and
+maintaining the cleanliness of indoor and outdoor spaces. Collected bags
+are temporarily stored in the sluice room, and cleaners during the night
+shift are responsible for transporting the waste on wheelchairs or
+trolleys to the incinerator for burning by waste disposers.
+
+Yet, according to respondents and our own observations, waste management
+procedures within the hospital are fraught with challenges which impact
+the implementation of best practices. These include the lack of a
+coherent hospital-wide waste management policy, insufficient waste
+management training for staff, and a persistent lack of material and
+financial resources. For instance, although hospital administration was
+able to articulate a set of standard operating principles (SoPs) for
+waste management, below the administration, doctors and nurses were
+generally limited in their understanding about its contents.
+Furthermore, cleaners, who are responsible for day-to-day cleansing and
+waste disposal, yet rate at the bottom of the institutional hierarchy at
+Queens, were unable to access or analyse hospital policy, and generally
+tended to rely on ``experience'' and ``human judgement'' when handling
+potentially hazardous waste. Moreover, although nearly all doctors and
+nurses had prior education on waste management best practices, cleaners
+began their waste handling duties without any prior waste management
+experience, and although there is an orientation programme for new
+hires, there is no regular refresher training or capacity building
+regarding waste for existing staff, at any level.
+
+Despite its national prominence, and international reputation for
+academic medical research, Queen's is plagued by constant shortages of
+human and material resources. Yet, these shortages are not shared
+equally across the space, with a sharp division in cleanliness,
+resources, and quality of care between public wards, supported by the
+Malawi Ministry of Health, and the private wads with international
+funding, such as the aforementioned Mercy James Institute. For staff,
+these shortages have manifested in poor salaries, missed pay cheques,
+failing infrastructure and limited supplies \citep{kalina2020this}.
+Regarding waste management, these shortages have manifested in a chronic
+shortage of black bags, waste containers and cleaning staff. As a
+result, due to a lack of bags and storage containers, there is minimal
+segregation of waste, and highly infectious items (bandages, gloves,
+syringes, and other sharps) are disposed of along with non-hazardous
+waste (food, newspapers, packaging). As a result, rather than
+facilitating different disposal pathways for hazardous and non-hazardous
+waste, all waste is mixed, necessitating more burning than would be
+necessary if non-hazardous waste was diverted, and contributing to
+workplace risk, with reports of accidental exposure to potentially
+infected waste, including accidental pokes from improperly disposed of
+sharps, common amongst hospital staff. Regardless, separation ends at
+the point of collection. The hospital is not serviced by regular
+municipal waste collection, and as a result, nearly all of the waste
+produced by the hospital, hazardous and non-hazardous, has to be
+disposed of on-site, either within the limited capacity of the hospital
+incinerator, or more commonly, through open burning. As a result, the
+constantly smouldering pile of waste puts forth a continuous cloud of
+grey smoke, which mingling with the dozens of other fires on the
+grounds, from cooking fires and burning garden waste, blankets Queen's
+in a permanent cloud of foul smelling haze.
+
+\hypertarget{measuring-air-quality}{%
+\subsection{Measuring Air Quality}\label{measuring-air-quality}}
+
+\hypertarget{hardware-and-software}{%
+\subsubsection{Hardware and Software}\label{hardware-and-software}}
+
+The PM monitoring device centres around a Raspberry Pi 3 Model B
+single-board computer complete with an operating system and storage
+space. The added advantage of storage space is that the data are not
+lost if the internet connection breaks down. Because the Pi can only
+read digital signals, we needed to include an analog-to-digital (ADC)
+converter. The amount of dust is measured by a Nova PM SDS011 High
+Precision Laser sensor. This sensor measures particles at 2.5 and 10
+micrometres in diameter in mg/m3. A list of the specific hardware and
+software components can be found in Appendix TODO.
+
+\hypertarget{installation}{%
+\subsubsection{Installation}\label{installation}}
+
+The choice of where the sensors were installed was done in collaboration
+with the QECH administration, but the position of the sensors at the
+individual buildings were decided by the research team. Four air quality
+sensors were installed on the outside of buildings (Guardian Shelter,
+Mercy James, Malaria Project, Lighthouse Clinic) while the remaining
+four sensors were installed inside (Ward 6B, Administration, Lions
+Sight, uMoyo). All of the sensors were mounted to a wall at a height of
+approximately two metres from the ground with the help of QECH
+maintenance department personnel. This height was chosen to prevent the
+public from tampering with the equipment when they visit the hospital.
+Additionally, we assured that each pipe that inhaled ambient air was
+freely protruding in the building (room) in order to capture the air
+quality.
+
+Because connecting each unit to a power supply was not possible, each
+unit was equipped with an external battery (5000mAh) which powered the
+sensor unit for three days continuously. To avoid any down time, we
+changed the batteries every two days. Data were also collected from the
+sensor at the time of battery replacement. The sensors were capable of
+being connected to WIFI, a feature which enabled us to download the data
+into our computers wirelessly.
+
+\hypertarget{qualitative-methods}{%
+\subsubsection{Qualitative Methods}\label{qualitative-methods}}
+
+Qualitative data collection consisted of 26 interviews with caregivers
+and hospital staff (including janitorial and maintenance staff, nurses,
+doctors, and administrators) conducted in October and November 2019, as
+well as an additional 31 interviews conducted with staff, patients, and
+caregivers in January and February 2021. Interviews were
+semi-structured, and included questions on hospital waste management
+practices, perceived best practices, and workplace health, safety, and
+risk. However, the semi-structured nature of the interviews allowed for
+the participants to raise topics of interest, and for unexpected themes
+to emerge. These interviews are supplemented by participatory
+observation recorded within detailed field notes, and supported by
+nearly a decade of sustained involvement and activism by the authors in
+the waste management practices of the hospital.
+
+Interview respondents were chosen using a purposive or judgement
+sampling regimen, i.e.~a subjective sampling method in which respondents
+are selected based on their ability to effectively contribute to the
+study's research objectives \citep{kitchin2013conducting}. All
+respondents were purposively selected based on availability, willingness
+to provide written, informed consent, and their individual insight into
+the questions posed within the study. Interviews were conducted in the
+local language (Chichewa), audio recorded, and transcribed into English.
+Participation was voluntary, and responses were recorded anonymously.
+Written, informed consent was obtained from each respondent prior to the
+interview. The study was approved by the National Committee on Research
+in the Social Sciences and Humanities (NCRSH) of Malawi; Protocol NO.
+P.03/19/356. Collected data were analysed thematically and was coded
+within the software programme Nvivo, which organises materials and
+assists with the coding process.
+
+\hypertarget{limitations}{%
+\subsubsection{Limitations}\label{limitations}}
+
+Though extensive and useful in understanding the conditions at QECH, the
+data set would benefit from a comparison with background values. As
+trash-burning is extensive throughout the city, there were no obvious
+locations that could be confidently used as representative background
+values. And although there were several possible areas upwind and
+outside of town, the logistics and safety involved in accessing them to
+change the batteries meant that the measured values at Queens could not
+be compared to a stable background concentration. Knowing the wind
+direction and velocity would have helped in better identifying the
+potential sources of the measured PM. However, the necessary equipment
+was not easily available in Blantyre at the time, and as an exploratory
+study, the immediate hazards for the staff and patients at QECH are
+relevant regardless of the burning source.
+
+Knowing the exact burning locations and burning times, especially at the
+incinerator and the Guardian Shelter, would have helped in better
+identifying the sources and movements of the plumes. However, given the
+size of the campus and it's 24-hour schedule, a much larger research
+team would have been required to quantify all the burning. Finally, The
+humidity, and the potential for interference was not accounted for; the
+measurements were not adjusted for humidity which could affect the
+values, though not enough to significantly alter the key findings and
+need for immediate action.
 
@@
 \section{Results and Discussion}\label{results-and-discussion}}
+
+The Malawi Bureau of Standards is the national agency responsible for
+setting and publishing all standards in the country. At the time of
+writing, the official website (www.mbsmw.org) was not available.
+However, published work that references Malawian standards
+\citep{kutlarjoss2017time, mapoma2014air} indicate that the maximum 24-h
+PM10 value is 25 μg/m3 and the maximum annual PM2.5 value is 8 μg/m3.
+There is no daily maximum value for PM2.5. It should be noted that both
+of these values are at, or below, the level-4 interim WHO targets.
+
+Therefore, for international comparisons and due to the lack of
+comprehensive Malawian Standards, we use the WHO recommendations as a
+basis of comparison for the measured values
+\citep{worldhealthorganization2021who}. The relevant particulate matter
+targets are presented in Table 1.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Table 1. Particulate matter targets from @worldhealthorganization2021who}
+\NormalTok{knitr}\SpecialCharTok{::}\FunctionTok{include\_graphics}\NormalTok{(here}\SpecialCharTok{::}\FunctionTok{here}\NormalTok{(}\StringTok{"manuscript/img/image{-}replace{-}table.png"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{img/image-replace-table.png}
+
+}
+
+\end{figure}
+
+\hypertarget{particulate-measurements}{%
+\subsection{Particulate Measurements}\label{particulate-measurements}}
+
+\hypertarget{peaks}{%
+\subsubsection{Peaks}\label{peaks}}
+
+Due to localised variation (air movement in the immediate area) and the
+density of the observations, the plotted 5-minute data obscure
+persistent trends. The full set of 5-minute data are plotted and
+presented in Appendix XA1 (TODO). However, the number of measured values
+that exceeded the hazardous limit for both parameters are summarised in
+Table 2.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Table 2.
+Peaks for both PM2.5 and PM10 are the number of data points above the WHO interim target 1 (annual) which is the least stringent;  the number of observations recorded (count) is provided for reference.
+}
+\end{Highlighting}
+\end{Shaded}
+
+Over the course of the two-month study, the hazardous limits for both
+parameters were exceeded at all locations except for the Lions Sight
+hospital. The limits were exceeded fewer than 50 times at five of the
+locations, and only the monitors at the Lighthouse Clinic and the
+Guardian Shelter recorded more than 50 instances above the hazardous
+limits. At the Lighthouse Clinic, the PM10 limit was exceeded almost
+twice as often as the PM2.5 limit, while at the Guardian Shelter, the
+reverse is true. Because the Lighthouse Clinic is within 50 m of the
+incinerator, the number and predominance of PM10 peaks is characteristic
+of incomplete combustion and dust that is typical for the area and the
+open burning that occurs there. The residents at the Guardian Shelter
+however are mostly cooking within contained clay stoves that have been
+designed to improve combustion of the coal and wood that is burned
+within them. The fewer number of PM10 peaks is a testament to this
+intervention, though the frequency of PM2.5 peaks is still beyond
+acceptable.
+
+\hypertarget{hour-averages}{%
+\subsubsection{24-hour averages}\label{hour-averages}}
+
+The 24-hour averaged values (logged every 5 minutes) are presented for
+each location, for both PM10 and PM2.5 in Figure 3. As well as dampening
+local variation and peaks, the WHO air quality guidelines are also based
+on 24-hour averages which are the standard against which the health
+risks can be judged.
+
+Overall, PM2.5 values remained below 100 μg/m3 at 6 of 8 locations
+(Administration, Lions Sight, Malaria Project, uMoyo, Ward 6B, and Mercy
+James); PM10 values were consistently below 100μg/m3 at the same
+locations, but with several average values extending slightly above, and
+then infrequently.
+
+The daily averages at both the Lighthouse Clinic and the Guardian
+Shelter are both consistently higher for both parameters and the two
+averages closely followed the same general trends. Though the peaks at
+the Lighthouse Clinic were higher than the Guardian Shelter, the low
+values were consistently lower, indicating more times of little or no
+burning, unlike the Guardian Shelter emissions which were relatively
+constant. However, further analysis of 12 hour averages (8:00-15:59
+(working hours) and 16:00-7:59 (evening)) did not indicate clear
+differences between the time periods; stated differently, the data did
+not clearly point to more or less burning in the day or at night.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Figure 3. Average 24{-}hour PM10 and PM2.5 at 8 monitoring stations over 2 months}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{pm-ratios}{%
+\subsubsection{PM ratios}\label{pm-ratios}}
+
+Given that the type of fuel and the type of burning (contained vs.~open)
+produce very different particulate ``fingerprints'', the ratio of PM10
+to PM2.5 values were examined to determine if a clear difference between
+locations, and therefore source, could be identified. The results are
+presented in Figure 4.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Figure 4. PM10:PM2.5 values for each location.  Each line shows the minimum and maximum ratio for the day based on ratios calculated hourly.  }
+\end{Highlighting}
+\end{Shaded}
+
+The results presented in Figure 5 illustrate both the presence of PM10
+relative to PM2.5 as well as the range of values over which that ratio
+is observed.
+
+Unlike the peak values presented in Table 2, the values presented in
+Figure 5 indicate the relative presence of larger (PM10) particles to
+smaller ones (PM2.5) regardless of their concentration: the ratio of two
+low concentrations can be the same as the ratio of two high
+concentrations and is therefore more indicative of source than of
+proximity.
+
+In general, the calculated ratios are concentrated between 1.25 with few
+values exceeding 7.5. Three from each the Guardian Shelter and the Ward
+6B exceeded 7.5 though the most values exceeding 5 were at the Guardian
+Shelter. It is interesting to note that despite the large daily
+variations at the Lighthouse Clinic, the parameter ratios there were
+fairly consistent with few spikes, and only 2 days with ratios above 5.
+
+An examination of the range of ratios is used to better understand the
+variability within a single day. For example, several of the calculated
+values at the Guardian Shelter have a range of more than 5, and one day
+had a range of close to 10 i.e.~the minimum ratio recorded for that day
+was close to 1 and the maximum value was over 10. The composition of
+emissions recorded at that location ranged significantly, and likely
+reflects a range of burning styles and/or fuel type.
+
+\hypertarget{exposure-data}{%
+\subsection{Exposure Data}\label{exposure-data}}
+
+The relative amount of time exposed to a given category of air quality
+is shown in Figure 5. Each square represents 1\% of the measured time.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Figure 5. Percentage of measured values categorised according to WHO 2021 targets}
+\end{Highlighting}
+\end{Shaded}
+
+Unsurprisingly, a majority of the air quality at the Lighthouse Clinic
+and the Guardian Shelter falls at, or below, interim target 1. What is
+less expected is that even though there are fewer days of hazardous or
+very bad air quality at the other locations, most locations do not have
+target 4 values for even half of the time. Mercy James, the Malaria
+Project and Lions Sight have the largest percentages of days with air
+quality meeting the interim target 4. Interim target 2 rated air
+occupies an unexpectedly large proportion of the days at all locations,
+indicating that although there are a few very toxic, and a few quite
+good days, the majority of the air is actually neither, but still cause
+for concern.
+
+\hypertarget{the-air-we-breathe-is-not-good-perspectives-from-qech}{%
+\subsection{``The Air We Breathe is Not Good'': Perspectives from
+QECH}\label{the-air-we-breathe-is-not-good-perspectives-from-qech}}
+
+As the previous section has described, the open incineration of waste at
+QECH has created hazardous conditions for those occupying the space.
+These risks were not lost on staff and caregivers, as interviews
+demonstrated broad and near universal awareness of the poor air quality
+within the hospital grounds. Overwhelmingly, within interviews, both
+caregivers and staff were quick to decry the poor quality of the air,
+generally without prompting. The few exceptions were those staff posted
+on the peripheries of the hospital grounds, at a distance from the
+spaces of incineration. However, even those who did not experience the
+impacts of the burning, were aware of it, and considered themselves
+fortunate to be posted in a section of the hospital where it was less of
+a problem. Furthermore, according to caregivers and staff who work night
+shifts, air quality can be particularly bad at night and in the early
+morning, despite the smoke being less visible, due to the habit of
+janitorial staff concentrating their burning during the late hours. The
+poor air quality on hospital grounds is also a frequent cause for
+complaint by patients and visitors, with nearly every staff member
+interviewed being able to recall having received a complaint, and in
+turn, complaining to administration. One of the staff members
+responsible for the burning said that he personally, had received
+hundreds of complaints, but was powerless to affect meaningful change,
+aside from burning at different hours, until the incinerator could be
+repaired.
+
+Nonetheless, despite this consensus that air quality caused by the
+burning of waste within the hospital was a problem, significant
+differences emerged between respondents over their understandings of
+potential impacts, the effectiveness of various coping mechanisms, and
+their problematisations linked to the burning of specific waste
+materials. In addition, interviews revealed that these understandings
+were informed by a significant amount of misinformation, even amongst
+trained medical staff, which may contribute to them being less able to
+mitigate potential risks for themselves and those who rely on their
+care.
+
+\hypertarget{impacts-problematisations-and-misinformation}{%
+\subsubsection{Impacts, Problematisations, and
+Misinformation}\label{impacts-problematisations-and-misinformation}}
+
+The poor air quality within QECH was responsible, according to
+respondents, for a wide array of health impacts. The most common ones
+cited included: coughing and sneezing, sore throat, stinging eyes,
+breathing difficulties, and persistent cold and flu. Nausea was also
+mentioned, but was not a commonly described impact. Only one respondent,
+of the 26 total interviewed, did not describe lingering health impacts
+which they could ascribe to the smoke, however, they also described
+having chronic eye irritation, but did not believe the smoke was a
+contributing factor.
+
+In addition to these impacts, which respondents bear on a daily basis,
+many also believed that the smoke could contribute to a number of more
+serious, long-term health complications. For instance, nearly a quarter
+of respondents raised concerns of the potential impact that the smoke
+could have on patients or staff with asthma. Others flagged poor air
+quality as a potential risk factor for certain cancers, lung disease, or
+heart problems. For a few, the smoke posed an unknown danger, they were
+not sure what types of impacts it could have, but they were sure it was
+harmful in some way.
+
+Also, understandably, given the large tuberculosis ward present within
+the hospital grounds, and the high prevalence of the disease within
+Malawi \citep{worldhealthorganization2019global} there was significant
+concern (more than half of respondents) about the impact the poor air
+quality could have on the infected. However, there also persisted a
+belief among several respondents, including several nurses, that the
+smoke could be a cause of the disease itself. As one staff member
+(20/11/2019) stated, ``I believe breathing this air for a long time can
+cause Tuberculosis.'' This, however, was only one of the few instances
+of misinformation which staff members held regarding air quality and
+health. Another example, voiced by several respondents, included a
+belief that some staff members were immune to ill-health impacts of the
+smoke, because they had received vaccinations from the hospital.
+Nonetheless, they were concerned about the impacts of the smoke on
+patients and visitors, as one staff member (18/11/2019) expressed:
+
+\begin{quote}
+Personally, I have never experienced {[}eye discomfort{]} because I get
+vaccinated and I am protected including other staff. However, we realize
+that the air can badly affect other people and patients who come to this
+place.
+\end{quote}
+
+Another interesting misconception that emerged, which may be tied partly
+to translation and transcription, was a different cultural understanding
+of smoke versus smells. More than half of the respondents appeared to
+conflate the two, with some expressing a belief that it was the odour of
+what was being burnt that was harmful, rather than the smoke being given
+off. This has led many to specifically problematise the burning of
+certain wastes, such as plastics, medicines, and other medical wastes,
+which give off distinctive or less pleasant odours, as opposed to the
+burning of other items, such as garden refuse, which may produce
+significant smoke, and contribute to higher recorded values of
+particulate matter, but produce a less pungent, or more normalised,
+odour.
+
+\hypertarget{coping-mechanisms}{%
+\subsubsection{Coping Mechanisms}\label{coping-mechanisms}}
+
+Finally, in order to manage the impacts of QECH's persistently poor air
+quality, staff and visitors reported having developed a number of coping
+mechanisms, designed to help them get through their daily routines.
+These included staying indoors, blocking doors and windows, and taking
+breaks away from hospital grounds in order to catch some breaths of
+fresher air. However, for janitorial staff, inside was not necessarily
+better, as several reported their indoor workspaces as being also
+insufferable for long periods of time from the smell of cleaning agents
+and other chemicals. Furthermore, amongst respondents there was a
+general disagreement over the effectiveness of personal protective
+equipment (PPE), such as face masks, towards mitigating the impacts of
+the smoke. A few staff described pleading to hospital administration for
+such equipment, to no avail. However, other staff members, who do have
+access to PPE, noted that even face masks do little to mitigate the
+impacts of the smoke, describing them as ineffective.
+
+Most staff, however, have been unable to find anyway to mitigate the
+impacts of the smoke, and only found relief once they reached home at
+the end of their shift, as one of the janitorial staff described, ``we
+only feel safe when we are home''. Of course, this relief is not
+available for the hundreds of patients and caregivers who sleep at QECH
+or are unable to leave. Ultimately, most place their hope in the
+construction of the new incinerator (which had not yet been activated at
+the time of the interviews), and biding their time as construction drags
+on; coping as best they can. As one staff member (25/11/2019)
+described,``we are just hoping we will start breathing good air soon,
+when the new incinerator is opened.''
 
@@
 etc.
-
-\hypertarget{template-instructions}{%
-\section{Template instructions}\label{template-instructions}}
-
-Please make sure that your manuscript follows the guidelines in the
-Guide for Authors of the relevant journal. It is not necessary to
-typeset your manuscript in exactly the same way as an article, unless
-you are submitting to a camera-ready copy (CRC) journal.
-
-For detailed instructions regarding the elsevier article class, see
-\url{https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions}
-
-\hypertarget{bibliography-styles}{%
-\section{Bibliography styles}\label{bibliography-styles}}
-
-Here are two sample references: \citet{Feynman1963118}
-\citet{Dirac1953888}.
-
-By default, natbib will be used with the \texttt{authoryear} style, set
-in \texttt{classoption} variable in YAML. You can sets extra options
-with \texttt{natbiboptions} variable in YAML header. Example
-
-\begin{verbatim}
-natbiboptions: longnamesfirst,angle,semicolon
-\end{verbatim}
-
-There are various more specific bibliography styles available at
-\url{https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files}.
-To use one of these, add it in the header using, for example,
-\texttt{biblio-style:\ model1-num-names}.
-
-\hypertarget{using-csl}{%
-\subsection{Using CSL}\label{using-csl}}
-
-If \texttt{cite-method} is set to \texttt{citeproc} in
-\texttt{elsevier\_article()}, then pandoc is used for citations instead
-of \texttt{natbib}. In this case, the \texttt{csl} option is used to
-format the references. By default, this template will provide an
-appropriate style, but alternative \texttt{csl} files are available from
-\url{https://www.zotero.org/styles?q=elsevier}. These can be downloaded
-and stored locally, or the url can be used as in the example header.
-
-\hypertarget{equations}{%
-\section{Equations}\label{equations}}
-
-Here is an equation: \[ 
-  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
-  \left(\frac{x}{\beta}\right)^{\alpha-1}
-  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
-  \alpha,\beta,x > 0 .
-\]
-
-Inline equations work as well: \(\sum_{i = 2}^\infty\{\alpha_i^\beta\}\)
 
@@
 \renewcommand\refname{References}
-  \bibliography{bibliography.bib,introduction.bib,man-qech-air-quality.bib}
+  \bibliography{man-qechairquality.bib}
 

### 2023-04-05 manuscript/manuscript.qmd
Align at 72 spaces

@@
 
-Built in 1964, Malawi's year of independence, Queen Elizabeth Hospital (QECH), or Queen's, as it's known locally, is the country's largest public hospital.
-Designed in the Tropical Modernism style of architecture popular within Africa in the late colonial period, Queen's open-plan form sprawls across a broad swath of central Blantyre, with dozens of wards, specialised facilities, and administrative buildings, linked together by a bewildering maze of covered walkways.
-A hive of around-the-clock activity, Queen's bustles from dawn to dusk with a constant stream of patients, drawn from across the country, and maintains a city-like feel even after dark as staff and stay-in family members cook and care for their in-patient dependants.
-Yet, for the first-time visitor, it's not the architecture or the crowds that are the most striking when entering the grounds, but the smell.
-Immediately noticeable, even from the road outside the hospital, Queen's is smothered by an acrid-smelling, white smoke, which hangs over the grounds day and night: the product of several constantly burning fires spread across the campus.
-Breathing this air, when a significant amount of burning is occurring, primarily at night when the smoke is less visible, can be incredibly uncomfortable.
-Moreover, the tropical architecture, designed to let air and light circulate, means that even indoors, within patient wards, surgery, and treatment areas, the air quality can also be unbearable.
-Poor air quality associated with burning, through released particulate matter (PM), has been linked to multiple negative health outcomes like asthma, heart disease, heart failure, stroke, and cancer among others [@anderson2012clearing; @bell2012assessment].
-In a space of healing and recovery, why is such burning occurring, and how can it continue to persist?
+Built in 1964, Malawi's year of independence, Queen Elizabeth Hospital
+(QECH), or Queen's, as it's known locally, is the country's largest
+public hospital. Designed in the Tropical Modernism style of
+architecture popular within Africa in the late colonial period, Queen's
+open-plan form sprawls across a broad swath of central Blantyre, with
+dozens of wards, specialised facilities, and administrative buildings,
+linked together by a bewildering maze of covered walkways. A hive of
+around-the-clock activity, Queen's bustles from dawn to dusk with a
+constant stream of patients, drawn from across the country, and
+maintains a city-like feel even after dark as staff and stay-in family
+members cook and care for their in-patient dependants. Yet, for the
+first-time visitor, it's not the architecture or the crowds that are the
+most striking when entering the grounds, but the smell. Immediately
+noticeable, even from the road outside the hospital, Queen's is
+smothered by an acrid-smelling, white smoke, which hangs over the
+grounds day and night: the product of several constantly burning fires
+spread across the campus. Breathing this air, when a significant amount
+of burning is occurring, primarily at night when the smoke is less
+visible, can be incredibly uncomfortable. Moreover, the tropical
+architecture, designed to let air and light circulate, means that even
+indoors, within patient wards, surgery, and treatment areas, the air
+quality can also be unbearable. Poor air quality associated with
+burning, through released particulate matter (PM), has been linked to
+multiple negative health outcomes like asthma, heart disease, heart
+failure, stroke, and cancer among others [@anderson2012clearing;
+@bell2012assessment]. In a space of healing and recovery, why is such
+burning occurring, and how can it continue to persist?
 
-Queen's, simply, produces an immense amount of waste that it is unable to efficiently manage.
-The dozens of wards, offices, and kitchens, serving the hundreds of patients and staff, generates considerable infectious medical waste, mixed domestic and office waste, and kitchen waste, which needs to be removed and disposed of daily.
-Some individual wards implement waste separation, binning medical wastes separately from general waste, however separation ends at the point of collection.
-Queen's is not served by regular municipal waste collection.
-What municipal collection that is done is ad-hoc, sporadic, and based on the hospital's available financial resources.
-Rather, most hospital waste (infectious and non-infectious) is gathered by grounds staff and burnt at the hospital's incinerator, located at a central point within the hospital campus [@kalina2020this].
-However, the vast majority of the deposited waste is not incinerated, as a new incinerator provided by the Ministry of Health in late 2019, is only able to handle a small percentage of the hospital's waste.
-Furthermore, the incinerator is frequently inoperable, under repair, or without power,.
-Rather, the waste is openly burned, in a constantly smouldering pile to the side of the incinerator building, sending up the aforementioned clouds of smoke, which choke the grounds (Figure 1).
-Innumerable other, smaller, fires, scattered across the QECH campus (Figure 2) contribute their part, as grounds and maintenance staff habitually burn leaves and other garden refuse.
-Furthermore, caregivers who reside separately on hospital grounds, cook for themselves throughout the day and night using biomass (charcoal or wood), and burn their own domestic waste.
-As a result, air quality at the hospital is a constant source of discussion, with patients, staff, and caregivers struggling to cope, when just breathing is, as the quote used in the title of this article, 'unbearable.'
+Queen's, simply, produces an immense amount of waste that it is unable
+to efficiently manage. The dozens of wards, offices, and kitchens,
+serving the hundreds of patients and staff, generates considerable
+infectious medical waste, mixed domestic and office waste, and kitchen
+waste, which needs to be removed and disposed of daily. Some individual
+wards implement waste separation, binning medical wastes separately from
+general waste, however separation ends at the point of collection.
+Queen's is not served by regular municipal waste collection. What
+municipal collection that is done is ad-hoc, sporadic, and based on the
+hospital's available financial resources. Rather, most hospital waste
+(infectious and non-infectious) is gathered by grounds staff and burnt
+at the hospital's incinerator, located at a central point within the
+hospital campus [@kalina2020this]. However, the vast majority of the
+deposited waste is not incinerated, as a new incinerator provided by the
+Ministry of Health in late 2019, is only able to handle a small
+percentage of the hospital's waste. Furthermore, the incinerator is
+frequently inoperable, under repair, or without power,. Rather, the
+waste is openly burned, in a constantly smouldering pile to the side of
+the incinerator building, sending up the aforementioned clouds of smoke,
+which choke the grounds (Figure 1). Innumerable other, smaller, fires,
+scattered across the QECH campus (Figure 2) contribute their part, as
+grounds and maintenance staff habitually burn leaves and other garden
+refuse. Furthermore, caregivers who reside separately on hospital
+grounds, cook for themselves throughout the day and night using biomass
+(charcoal or wood), and burn their own domestic waste. As a result, air
+quality at the hospital is a constant source of discussion, with
+patients, staff, and caregivers struggling to cope, when just breathing
+is, as the quote used in the title of this article, 'unbearable.'
 
-Challenges with solid waste management (SWM) are not unique to Queen's, and remain persistent globally, particularly within African nations, and the Global South more broadly [@bundhoo2018solid; @sergekubanza2022exploring; @muheirwe2022paradox; @srivastava2015urban].
-Explanations for these disparities have centred on the on waste generation outstripping waste management infrastructure [@muheirwe2022paradox], and the high costs of waste management systems, which often prove unaffordable for many low-and-middle-income countries (LMICs) [@guerrero2013solid; @sergekubanza2022exploring].
-Waste collection is an important step within SWM systems, and is a common barrier within LMIC contexts [@bundhoo2018solid].
-Waste inequalities between high-income countries (HICs) and low-income countries (LICs) is often most pernicious and visible at the point of collection, with some HICs achieving 100% collection rates, while some African nations have rates well below 50% [@shi2021unbalanced].
-Inequalities persist beyond collection, however.
-Even once waste is collected, municipalities in LMICs may burn or dump it in non-sanitary landfills due to the absence of further treatment or disposal options [@cook2021globala].
-According to @kaza2018what, 90% of waste in LIC is disposed of in unregulated dumps or openly burned, while the quantity of waste generation in these countries is expected to triple by 2050.
-As a result, open waste burning, in nations like Malawi, can be expected to become more prevalent.
+Challenges with solid waste management (SWM) are not unique to Queen's,
+and remain persistent globally, particularly within African nations, and
+the Global South more broadly [@bundhoo2018solid;
+@sergekubanza2022exploring; @muheirwe2022paradox; @srivastava2015urban].
+Explanations for these disparities have centred on the on waste
+generation outstripping waste management infrastructure
+[@muheirwe2022paradox], and the high costs of waste management systems,
+which often prove unaffordable for many low-and-middle-income countries
+(LMICs) [@guerrero2013solid; @sergekubanza2022exploring]. Waste
+collection is an important step within SWM systems, and is a common
+barrier within LMIC contexts [@bundhoo2018solid]. Waste inequalities
+between high-income countries (HICs) and low-income countries (LICs) is
+often most pernicious and visible at the point of collection, with some
+HICs achieving 100% collection rates, while some African nations have
+rates well below 50% [@shi2021unbalanced]. Inequalities persist beyond
+collection, however. Even once waste is collected, municipalities in
+LMICs may burn or dump it in non-sanitary landfills due to the absence
+of further treatment or disposal options [@cook2021globala]. According
+to @kaza2018what, 90% of waste in LIC is disposed of in unregulated
+dumps or openly burned, while the quantity of waste generation in these
+countries is expected to triple by 2050. As a result, open waste
+burning, in nations like Malawi, can be expected to become more
+prevalent.
 
@@
 
-Although hospital waste is among the common types of solid waste openly burned in LICs (in addition to municipal solid waste, sewage sludge, market or commercial waste, agricultural residues), the open burning of hospital waste has not garnered as much discussion as the burning of other waste fractions [@cook2021globala].
-The World Health Organisation (WHO) defines hospital, or health-care waste, as waste generated by health-care activities, of which approximately 85% is general, non-hazardous waste, but the remaining 15% can be considered hazardous, and may be toxic, infectious, or radioactive.
-These infectious fractions are commonly incinerated in both high and middle-income countries, however in the absence of capacity, or costly incineration or autoclave infrastructure, open burning is often a last resort option for disposal in LIC contexts [@who2018healthcare].
-The open burning of biomedical waste may reduce infection risk from potentially harmful pathogens, however, such waste may contain sharps, radioactive waste, mercury-containing instruments, and plastics, the open burning of which may result in the emission of dioxins, furans, and toxic particulate matter (PM) [@who2018healthcare; @cook2021globala].
+Although hospital waste is among the common types of solid waste openly
+burned in LICs (in addition to municipal solid waste, sewage sludge,
+market or commercial waste, agricultural residues), the open burning of
+hospital waste has not garnered as much discussion as the burning of
+other waste fractions [@cook2021globala]. The World Health Organisation
+(WHO) defines hospital, or health-care waste, as waste generated by
+health-care activities, of which approximately 85% is general,
+non-hazardous waste, but the remaining 15% can be considered hazardous,
+and may be toxic, infectious, or radioactive. These infectious fractions
+are commonly incinerated in both high and middle-income countries,
+however in the absence of capacity, or costly incineration or autoclave
+infrastructure, open burning is often a last resort option for disposal
+in LIC contexts [@who2018healthcare]. The open burning of biomedical
+waste may reduce infection risk from potentially harmful pathogens,
+however, such waste may contain sharps, radioactive waste,
+mercury-containing instruments, and plastics, the open burning of which
+may result in the emission of dioxins, furans, and toxic particulate
+matter (PM) [@who2018healthcare; @cook2021globala].
 
-Poor air quality in in African cities can be attributed to multiple factors, including vehicular emissions, dust, or the use of low-quality fuels for cooking, lighting, and heating, however, the open burning of waste has been identified as one of the key contributors to air pollution within urban areas [@amegah2018proliferation; @bulto2020impact].
-Although open trashing burning is a problem in LICs broadly, for instance, @wiedinmyer2014global estimate that open trash burning contributes to 29% of the total global PM2.5 anthropogenic emissions, it is of particular concern in Sub-Saharan African (SSA) cities, which are home to 19 of the world's 50 biggest dumpsites (United Nations Development Programme, 2021).
-Particulate matter, or PM, is not specific to trash burning, but rather the product of any type of incomplete combustion and is one of the most commonly measured indicators for air quality.
-PM is measured in terms of its particle size and generally classified into PM with particle sizes less than 10μm (PM10) and those with sizes less than 2.5μm (PM2.5), while one way of measuring the severity of air pollution is by expressing the mass of PM per volume of air (μg PMx/m3 air).
-PM has been linked to multiple negative health outcomes.
-Amongst particulate air pollution, PM10 and PM2.5 are of particular concern as they can penetrate deep into the lungs, and their exposure has been associated with asthma, heart disease, heart failure, stroke, and cancer [@mohapatra2014effect; @xing2016impact].
-As a result, the health impacts of ambient air particle pollution may be significant, especially for the most vulnerable.
-For instance, the @ihme2019ambient estimates PM related air pollution resulted in 4.14 million deaths worldwide (in 2019), of which, 394,000 occurred in Africa.
-Yet, despite its importance, the air quality and health picture in Africa is worryingly incomplete.
-For instance, the WHO only collects data from 10 African countries (covering 39 cities) [@worldhealthorganization2016ambient].
-Likewise, robust, epidemiological data from and for the African context is also lacking, though emerging data suggests strong correlations between PM and poor health outcomes [@coker2018narrative; @heft-neal2018robust].
-Moving forward, it is not just the air at Queen's which lacks clarity.
+Poor air quality in in African cities can be attributed to multiple
+factors, including vehicular emissions, dust, or the use of low-quality
+fuels for cooking, lighting, and heating, however, the open burning of
+waste has been identified as one of the key contributors to air
+pollution within urban areas [@amegah2018proliferation;
+@bulto2020impact]. Although open trashing burning is a problem in LICs
+broadly, for instance, @wiedinmyer2014global estimate that open trash
+burning contributes to 29% of the total global PM2.5 anthropogenic
+emissions, it is of particular concern in Sub-Saharan African (SSA)
+cities, which are home to 19 of the world's 50 biggest dumpsites (United
+Nations Development Programme, 2021). Particulate matter, or PM, is not
+specific to trash burning, but rather the product of any type of
+incomplete combustion and is one of the most commonly measured
+indicators for air quality. PM is measured in terms of its particle size
+and generally classified into PM with particle sizes less than 10μm
+(PM10) and those with sizes less than 2.5μm (PM2.5), while one way of
+measuring the severity of air pollution is by expressing the mass of PM
+per volume of air (μg PMx/m3 air). PM has been linked to multiple
+negative health outcomes. Amongst particulate air pollution, PM10 and
+PM2.5 are of particular concern as they can penetrate deep into the
+lungs, and their exposure has been associated with asthma, heart
+disease, heart failure, stroke, and cancer [@mohapatra2014effect;
+@xing2016impact]. As a result, the health impacts of ambient air
+particle pollution may be significant, especially for the most
+vulnerable. For instance, the @ihme2019ambient estimates PM related air
+pollution resulted in 4.14 million deaths worldwide (in 2019), of which,
+394,000 occurred in Africa. Yet, despite its importance, the air quality
+and health picture in Africa is worryingly incomplete. For instance, the
+WHO only collects data from 10 African countries (covering 39 cities)
+[@worldhealthorganization2016ambient]. Likewise, robust, epidemiological
+data from and for the African context is also lacking, though emerging
+data suggests strong correlations between PM and poor health outcomes
+[@coker2018narrative; @heft-neal2018robust]. Moving forward, it is not
+just the air at Queen's which lacks clarity.
 
-Utilising a mixed-methods approach, including a network of custom-designed air quality monitors and qualitative fieldwork with hospital staff, patients, and caregivers, the purpose of this investigation was to gain a multi-dimensional understanding of the impacts of the open burning of the various solid and organic waste fractions at QECH.
-Specifically, this work aimed to intensively and longitudinally measure the air quality, in particulate matter (PM), at multiple locations at and surrounding the central burning point.
-Furthermore, we also aimed to qualitatively understand how affected individuals perceive air quality at QECH and understand potential health impacts.
+Utilising a mixed-methods approach, including a network of
+custom-designed air quality monitors and qualitative fieldwork with
+hospital staff, patients, and caregivers, the purpose of this
+investigation was to gain a multi-dimensional understanding of the
+impacts of the open burning of the various solid and organic waste
+fractions at QECH. Specifically, this work aimed to intensively and
+longitudinally measure the air quality, in particulate matter (PM), at
+multiple locations at and surrounding the central burning point.
+Furthermore, we also aimed to qualitatively understand how affected
+individuals perceive air quality at QECH and understand potential health
+impacts.
 
-This research responds to a several specific gaps within the body of academic air quality literature: Although there has been ample discussion of air quality challenges in African cities [@mbow-diokhane2019air; @miralvarez2020scoping; @petkova2013particulate; @singh2021aira], there is a paucity of data on air quality impacts linked to trash burning.
-In particular, within African cities, there has been a total dearth of scholarship on the burning of medical waste, or on air quality within hospital contexts.
-Furthermore, while the voices of the citizens of Northern cities [@deguen2012new; @howel2002urban; @reames2019people] have been well documented on localised air quality issues, the voices of African urban dwellers have not been given the same consideration, and there remains a pressing need to understand how these populations experience the impacts of the open burning of trash within their communities and understand potential risks.
+This research responds to a several specific gaps within the body of
+academic air quality literature: Although there has been ample
+discussion of air quality challenges in African cities
+[@mbow-diokhane2019air; @miralvarez2020scoping; @petkova2013particulate;
+@singh2021aira], there is a paucity of data on air quality impacts
+linked to trash burning. In particular, within African cities, there has
+been a total dearth of scholarship on the burning of medical waste, or
+on air quality within hospital contexts. Furthermore, while the voices
+of the citizens of Northern cities [@deguen2012new; @howel2002urban;
+@reames2019people] have been well documented on localised air quality
+issues, the voices of African urban dwellers have not been given the
+same consideration, and there remains a pressing need to understand how
+these populations experience the impacts of the open burning of trash
+within their communities and understand potential risks.
 
-Findings suggest that particulate matter concentrations are routinely above the WHO guidelines and are especially worrisome at several locations.
-Over the course of the two-month study, the hazardous limits for both PM10 and PM2.5 were exceeded at all locations except for the Lions Sight hospital.
-The limits were exceeded fewer than 50 times at five of the locations, but the monitors at the Lighthouse Clinic and the Guardian Shelter recorded more than 50 instances above the hazardous limits.
-The extremely hazardous air quality at the Guardian shelter is mostly from cooking over open fires, while the PM at the Lighthouse Clinic is a function of being located directly adjacent to the smouldering waste pile.
+Findings suggest that particulate matter concentrations are routinely
+above the WHO guidelines and are especially worrisome at several
+locations. Over the course of the two-month study, the hazardous limits
+for both PM10 and PM2.5 were exceeded at all locations except for the
+Lions Sight hospital. The limits were exceeded fewer than 50 times at
+five of the locations, but the monitors at the Lighthouse Clinic and the
+Guardian Shelter recorded more than 50 instances above the hazardous
+limits. The extremely hazardous air quality at the Guardian shelter is
+mostly from cooking over open fires, while the PM at the Lighthouse
+Clinic is a function of being located directly adjacent to the
+smouldering waste pile.
 
-Interviews show that patients, staff, and caregivers alike, are keenly aware that the air quality at Queen's is poor, with most respondents reporting frequent respiratory-related illness.
-Many also linked the smoke to potential long-term health complications and expressed a belief that the pollution could be contributing to other, more potentially life-threatening diseases, such as asthma, cancer, and tuberculosis.
-Moreover, the tropical design of hospital buildings have rendered most coping mechanisms ineffective, with most staff only finding relief at home: relief that is not available to the hundreds of patients and caregivers who sleep at QECH or are unable to leave.
+Interviews show that patients, staff, and caregivers alike, are keenly
+aware that the air quality at Queen's is poor, with most respondents
+reporting frequent respiratory-related illness. Many also linked the
+smoke to potential long-term health complications and expressed a belief
+that the pollution could be contributing to other, more potentially
+life-threatening diseases, such as asthma, cancer, and tuberculosis.
+Moreover, the tropical design of hospital buildings have rendered most
+coping mechanisms ineffective, with most staff only finding relief at
+home: relief that is not available to the hundreds of patients and
+caregivers who sleep at QECH or are unable to leave.
 
-In many contexts, waste burning is a necessary, but sporadic event with limited exposure.
-However, in the context of a large hospital, serving the sickest and poorest, this study shows that the consistent, and highly toxic smoke produced is an unrelenting and unnecessary burden that must be addressed, lest the patients leave in worse condition than in which they arrived.
+In many contexts, waste burning is a necessary, but sporadic event with
+limited exposure. However, in the context of a large hospital, serving
+the sickest and poorest, this study shows that the consistent, and
+highly toxic smoke produced is an unrelenting and unnecessary burden
+that must be addressed, lest the patients leave in worse condition than
+in which they arrived.
 
@@
 
-This study utilised a mixed-methods approach to both quantitatively measure the air quality at eight locations around QECH, and to qualitatively investigate the perceived impacts amongst staff and caregivers.
-Although the work was conducted over one sustained period in late 2019, it must be contextualised within five years' experience of research and activism within QECH by the authors.
-All relevant permissions were obtained from QECH beforehand, through consultations with administration and staff, and the research was approved by the National Committee On Research in The Social Sciences And Humanities in Malawi, Protocol NO.
-P.03/19/356.
+This study utilised a mixed-methods approach to both quantitatively
+measure the air quality at eight locations around QECH, and to
+qualitatively investigate the perceived impacts amongst staff and
+caregivers. Although the work was conducted over one sustained period in
+late 2019, it must be contextualised within five years' experience of
+research and activism within QECH by the authors. All relevant
+permissions were obtained from QECH beforehand, through consultations
+with administration and staff, and the research was approved by the
+National Committee On Research in The Social Sciences And Humanities in
+Malawi, Protocol NO. P.03/19/356.
 
@@
 
-The sensors were located within the QECH campus in consultation with hospital management.
-The goal was to locate sensors across the campus with a range of distances and directions from the incinerator, though the final decision was based on accessibility, the permission of the unit's head, and convenience.
-Each of the locations is briefly described below and shown in Figure 2:
+The sensors were located within the QECH campus in consultation with
+hospital management. The goal was to locate sensors across the campus
+with a range of distances and directions from the incinerator, though
+the final decision was based on accessibility, the permission of the
+unit's head, and convenience. Each of the locations is briefly described
+below and shown in Figure 2:
 
--   **Lions Sight First Eye Hospital**, often shortened to **Lions Sight**, is the largest eye hospital in Malawi and serves as the main teaching eye hospital for the Kamuzu University of Health Sciences (formerly College of Medicine).
-It is staffed by 5 consultants and a team of clinical officers and nurses.
-It provides a mix of public (no cost) and private (at cost) services.
+-   **Lions Sight First Eye Hospital**, often shortened to **Lions
+    Sight**, is the largest eye hospital in Malawi and serves as the
+    main teaching eye hospital for the Kamuzu University of Health
+    Sciences (formerly College of Medicine). It is staffed by 5
+    consultants and a team of clinical officers and nurses. It provides
+    a mix of public (no cost) and private (at cost) services.
 
--   **The Blantyre Malaria Project (BMP)**, established by Professors Terrie Taylor and Malcolm Molyneux, has carried out clinical research and patient care in the area of paediatric malaria, specifically cerebral malaria, for more than 30 years.
-Its research infrastructure includes an administrative team, an inpatient research ward, an MRI centre, and an outpatient research clinic in Ndirande township and in other districts outside of Blantyre.
+-   **The Blantyre Malaria Project (BMP)**, established by Professors
+    Terrie Taylor and Malcolm Molyneux, has carried out clinical
+    research and patient care in the area of paediatric malaria,
+    specifically cerebral malaria, for more than 30 years. Its research
+    infrastructure includes an administrative team, an inpatient
+    research ward, an MRI centre, and an outpatient research clinic in
+    Ndirande township and in other districts outside of Blantyre.
 
--   **Ward 6B** is a male ward for trauma and orthopaedic patients.
-It was established as part of the original hospital design.
-It is a 60-bed ward with a dedicated nursing station and treatment rooms.
+-   **Ward 6B** is a male ward for trauma and orthopaedic patients. It
+    was established as part of the original hospital design. It is a
+    60-bed ward with a dedicated nursing station and treatment rooms.
 
--   **The Blantyre Lighthouse Trust Clinic**, also known as **Umodzi Family Centre** or simply, the **Lighthouse Clinic**, is one of four operating across Malawi.
-The clinics work in collaboration with the Ministry of Health (MOH) to provide integrated HIV testing, treatment and care for people living with HIV.
+-   **The Blantyre Lighthouse Trust Clinic**, also known as **Umodzi
+    Family Centre** or simply, the **Lighthouse Clinic**, is one of four
+    operating across Malawi. The clinics work in collaboration with the
+    Ministry of Health (MOH) to provide integrated HIV testing,
+    treatment and care for people living with HIV.
 
--   The **Moyo Nutritional Rehabilitation and Research Unit**, known on campus as **uMoyo**, is a 57-bed nutritional rehabilitation unit (NRU) for treating infants and children with severe malnutrition and acute illnesses.
-It is one of 104 operational NRUs in Malawi.
-It also has an outpatient therapeutic feeding program (OTP) for children with malnutrition who can be treated outside of the hospital.
+-   The **Moyo Nutritional Rehabilitation and Research Unit**, known on
+    campus as **uMoyo**, is a 57-bed nutritional rehabilitation unit
+    (NRU) for treating infants and children with severe malnutrition and
+    acute illnesses. It is one of 104 operational NRUs in Malawi. It
+    also has an outpatient therapeutic feeding program (OTP) for
+    children with malnutrition who can be treated outside of the
+    hospital.
 
--   **Mercy James Institute for Pediatric Surgery and Intensive Care (MJC)**, shortened on campus to **Mercy James**, was opened in 2017 by the NGO Raising Malawi, which was founded by Madonna.
-With 3 operating rooms and 50 beds, it is the first and only first dedicated paediatric surgery hospital in Malawi.
-Prior to its opening, QECH had fewer than 10 intensive care beds.
+-   **Mercy James Institute for Pediatric Surgery and Intensive Care
+    (MJC)**, shortened on campus to **Mercy James**, was opened in 2017
+    by the NGO Raising Malawi, which was founded by Madonna. With 3
+    operating rooms and 50 beds, it is the first and only first
+    dedicated paediatric surgery hospital in Malawi. Prior to its
+    opening, QECH had fewer than 10 intensive care beds.
 
--   The **Administration Building** is the operational hub of the campus, containing mostly offices and meeting rooms.
+-   The **Administration Building** is the operational hub of the
+    campus, containing mostly offices and meeting rooms.
 
--   When a patient arrives at QECH, she or he will usually arrive with a guardian: someone to cook for them, buy medicine, do their laundry, and help them bathe.
-The **Guardian Shelter** is a gender-separated concrete floor shelter for sleeping along with a cooking pavilion and toilet/shower block are maintained by a local NGO.
-Though simple, the facility is secure and offers one of the only free/accessible toilet facilities for visitors on the QECH campus.
+-   When a patient arrives at QECH, she or he will usually arrive with a
+    guardian: someone to cook for them, buy medicine, do their laundry,
+    and help them bathe. The **Guardian Shelter** is a gender-separated
+    concrete floor shelter for sleeping along with a cooking pavilion
+    and toilet/shower block are maintained by a local NGO. Though
+    simple, the facility is secure and offers one of the only
+    free/accessible toilet facilities for visitors on the QECH campus.
 
@@
 
-As the largest public hospital in the country, Queen's produces an immense amount of waste.
-Hospital staff, such as doctors and nurses, primarily generate and handle medical waste (dry, wet, sharps), which is disposed of in separate waste bins labelled accordingly within the ward.
-Hospital visitors (caretakers and patients) generate food waste and are advised by staff to discard such waste in waste bins outside, on hospital grounds.
-Cleaning staff are responsible for emptying bins and maintaining the cleanliness of indoor and outdoor spaces.
-Collected bags are temporarily stored in the sluice room, and cleaners during the night shift are responsible for transporting the waste on wheelchairs or trolleys to the incinerator for burning by waste disposers.
+As the largest public hospital in the country, Queen's produces an
+immense amount of waste. Hospital staff, such as doctors and nurses,
+primarily generate and handle medical waste (dry, wet, sharps), which is
+disposed of in separate waste bins labelled accordingly within the ward.
+Hospital visitors (caretakers and patients) generate food waste and are
+advised by staff to discard such waste in waste bins outside, on
+hospital grounds. Cleaning staff are responsible for emptying bins and
+maintaining the cleanliness of indoor and outdoor spaces. Collected bags
+are temporarily stored in the sluice room, and cleaners during the night
+shift are responsible for transporting the waste on wheelchairs or
+trolleys to the incinerator for burning by waste disposers.
 
-Yet, according to respondents and our own observations, waste management procedures within the hospital are fraught with challenges which impact the implementation of best practices.
-These include the lack of a coherent hospital-wide waste management policy, insufficient waste management training for staff, and a persistent lack of material and financial resources.
-For instance, although hospital administration was able to articulate a set of standard operating principles (SoPs) for waste management, below the administration, doctors and nurses were generally limited in their understanding about its contents.
-Furthermore, cleaners, who are responsible for day-to-day cleansing and waste disposal, yet rate at the bottom of the institutional hierarchy at Queens, were unable to access or analyse hospital policy, and generally tended to rely on "experience" and "human judgement" when handling potentially hazardous waste.
-Moreover, although nearly all doctors and nurses had prior education on waste management best practices, cleaners began their waste handling duties without any prior waste management experience, and although there is an orientation programme for new hires, there is no regular refresher training or capacity building regarding waste for existing staff, at any level.
+Yet, according to respondents and our own observations, waste management
+procedures within the hospital are fraught with challenges which impact
+the implementation of best practices. These include the lack of a
+coherent hospital-wide waste management policy, insufficient waste
+management training for staff, and a persistent lack of material and
+financial resources. For instance, although hospital administration was
+able to articulate a set of standard operating principles (SoPs) for
+waste management, below the administration, doctors and nurses were
+generally limited in their understanding about its contents.
+Furthermore, cleaners, who are responsible for day-to-day cleansing and
+waste disposal, yet rate at the bottom of the institutional hierarchy at
+Queens, were unable to access or analyse hospital policy, and generally
+tended to rely on "experience" and "human judgement" when handling
+potentially hazardous waste. Moreover, although nearly all doctors and
+nurses had prior education on waste management best practices, cleaners
+began their waste handling duties without any prior waste management
+experience, and although there is an orientation programme for new
+hires, there is no regular refresher training or capacity building
+regarding waste for existing staff, at any level.
 
-Despite its national prominence, and international reputation for academic medical research, Queen's is plagued by constant shortages of human and material resources.
-Yet, these shortages are not shared equally across the space, with a sharp division in cleanliness, resources, and quality of care between public wards, supported by the Malawi Ministry of Health, and the private wads with international funding, such as the aforementioned Mercy James Institute.
-For staff, these shortages have manifested in poor salaries, missed pay cheques, failing infrastructure and limited supplies [@kalina2020this].
-Regarding waste management, these shortages have manifested in a chronic shortage of black bags, waste containers and cleaning staff.
-As a result, due to a lack of bags and storage containers, there is minimal segregation of waste, and highly infectious items (bandages, gloves, syringes, and other sharps) are disposed of along with non-hazardous waste (food, newspapers, packaging).
-As a result, rather than facilitating different disposal pathways for hazardous and non-hazardous waste, all waste is mixed, necessitating more burning than would be necessary if non-hazardous waste was diverted, and contributing to workplace risk, with reports of accidental exposure to potentially infected waste, including accidental pokes from improperly disposed of sharps, common amongst hospital staff.
-Regardless, separation ends at the point of collection.
-The hospital is not serviced by regular municipal waste collection, and as a result, nearly all of the waste produced by the hospital, hazardous and non-hazardous, has to be disposed of on-site, either within the limited capacity of the hospital incinerator, or more commonly, through open burning.
-As a result, the constantly smouldering pile of waste puts forth a continuous cloud of grey smoke, which mingling with the dozens of other fires on the grounds, from cooking fires and burning garden waste, blankets Queen's in a permanent cloud of foul smelling haze.
+Despite its national prominence, and international reputation for
+academic medical research, Queen's is plagued by constant shortages of
+human and material resources. Yet, these shortages are not shared
+equally across the space, with a sharp division in cleanliness,
+resources, and quality of care between public wards, supported by the
+Malawi Ministry of Health, and the private wads with international
+funding, such as the aforementioned Mercy James Institute. For staff,
+these shortages have manifested in poor salaries, missed pay cheques,
+failing infrastructure and limited supplies [@kalina2020this]. Regarding
+waste management, these shortages have manifested in a chronic shortage
+of black bags, waste containers and cleaning staff. As a result, due to
+a lack of bags and storage containers, there is minimal segregation of
+waste, and highly infectious items (bandages, gloves, syringes, and
+other sharps) are disposed of along with non-hazardous waste (food,
+newspapers, packaging). As a result, rather than facilitating different
+disposal pathways for hazardous and non-hazardous waste, all waste is
+mixed, necessitating more burning than would be necessary if
+non-hazardous waste was diverted, and contributing to workplace risk,
+with reports of accidental exposure to potentially infected waste,
+including accidental pokes from improperly disposed of sharps, common
+amongst hospital staff. Regardless, separation ends at the point of
+collection. The hospital is not serviced by regular municipal waste
+collection, and as a result, nearly all of the waste produced by the
+hospital, hazardous and non-hazardous, has to be disposed of on-site,
+either within the limited capacity of the hospital incinerator, or more
+commonly, through open burning. As a result, the constantly smouldering
+pile of waste puts forth a continuous cloud of grey smoke, which
+mingling with the dozens of other fires on the grounds, from cooking
+fires and burning garden waste, blankets Queen's in a permanent cloud of
+foul smelling haze.
 
@@
 
-The PM monitoring device centres around a Raspberry Pi 3 Model B single-board computer complete with an operating system and storage space.
-The added advantage of storage space is that the data are not lost if the internet connection breaks down.
-Because the Pi can only read digital signals, we needed to include an analog-to-digital (ADC) converter.
-The amount of dust is measured by a Nova PM SDS011 High Precision Laser sensor.
-This sensor measures particles at 2.5 and 10 micrometres in diameter in mg/m3.
-A list of the specific hardware and software components can be found in Appendix TODO.
+The PM monitoring device centres around a Raspberry Pi 3 Model B
+single-board computer complete with an operating system and storage
+space. The added advantage of storage space is that the data are not
+lost if the internet connection breaks down. Because the Pi can only
+read digital signals, we needed to include an analog-to-digital (ADC)
+converter. The amount of dust is measured by a Nova PM SDS011 High
+Precision Laser sensor. This sensor measures particles at 2.5 and 10
+micrometres in diameter in mg/m3. A list of the specific hardware and
+software components can be found in Appendix TODO.
 
@@
 
-The choice of where the sensors were installed was done in collaboration with the QECH administration, but the position of the sensors at the individual buildings were decided by the research team.
-Four air quality sensors were installed on the outside of buildings (Guardian Shelter, Mercy James, Malaria Project, Lighthouse Clinic) while the remaining four sensors were installed inside (Ward 6B, Administration, Lions Sight, uMoyo).
-All of the sensors were mounted to a wall at a height of approximately two metres from the ground with the help of QECH maintenance department personnel.
-This height was chosen to prevent the public from tampering with the equipment when they visit the hospital.
-Additionally, we assured that each pipe that inhaled ambient air was freely protruding in the building (room) in order to capture the air quality.
+The choice of where the sensors were installed was done in collaboration
+with the QECH administration, but the position of the sensors at the
+individual buildings were decided by the research team. Four air quality
+sensors were installed on the outside of buildings (Guardian Shelter,
+Mercy James, Malaria Project, Lighthouse Clinic) while the remaining
+four sensors were installed inside (Ward 6B, Administration, Lions
+Sight, uMoyo). All of the sensors were mounted to a wall at a height of
+approximately two metres from the ground with the help of QECH
+maintenance department personnel. This height was chosen to prevent the
+public from tampering with the equipment when they visit the hospital.
+Additionally, we assured that each pipe that inhaled ambient air was
+freely protruding in the building (room) in order to capture the air
+quality.
 
-Because connecting each unit to a power supply was not possible, each unit was equipped with an external battery (5000mAh) which powered the sensor unit for three days continuously.
-To avoid any down time, we changed the batteries every two days.
-Data were also collected from the sensor at the time of battery replacement.
-The sensors were capable of being connected to WIFI, a feature which enabled us to download the data into our computers wirelessly.
+Because connecting each unit to a power supply was not possible, each
+unit was equipped with an external battery (5000mAh) which powered the
+sensor unit for three days continuously. To avoid any down time, we
+changed the batteries every two days. Data were also collected from the
+sensor at the time of battery replacement. The sensors were capable of
+being connected to WIFI, a feature which enabled us to download the data
+into our computers wirelessly.
 
@@
 
-Qualitative data collection consisted of 26 interviews with caregivers and hospital staff (including janitorial and maintenance staff, nurses, doctors, and administrators) conducted in October and November 2019, as well as an additional 31 interviews conducted with staff, patients, and caregivers in January and February 2021.
-Interviews were semi-structured, and included questions on hospital waste management practices, perceived best practices, and workplace health, safety, and risk.
-However, the semi-structured nature of the interviews allowed for the participants to raise topics of interest, and for unexpected themes to emerge.
-These interviews are supplemented by participatory observation recorded within detailed field notes, and supported by nearly a decade of sustained involvement and activism by the authors in the waste management practices of the hospital.
+Qualitative data collection consisted of 26 interviews with caregivers
+and hospital staff (including janitorial and maintenance staff, nurses,
+doctors, and administrators) conducted in October and November 2019, as
+well as an additional 31 interviews conducted with staff, patients, and
+caregivers in January and February 2021. Interviews were
+semi-structured, and included questions on hospital waste management
+practices, perceived best practices, and workplace health, safety, and
+risk. However, the semi-structured nature of the interviews allowed for
+the participants to raise topics of interest, and for unexpected themes
+to emerge. These interviews are supplemented by participatory
+observation recorded within detailed field notes, and supported by
+nearly a decade of sustained involvement and activism by the authors in
+the waste management practices of the hospital.
 
-Interview respondents were chosen using a purposive or judgement sampling regimen, i.e.
-a subjective sampling method in which respondents are selected based on their ability to effectively contribute to the study's research objectives [@kitchin2013conducting].
-All respondents were purposively selected based on availability, willingness to provide written, informed consent, and their individual insight into the questions posed within the study.
-Interviews were conducted in the local language (Chichewa), audio recorded, and transcribed into English.
+Interview respondents were chosen using a purposive or judgement
+sampling regimen, i.e. a subjective sampling method in which respondents
+are selected based on their ability to effectively contribute to the
+study's research objectives [@kitchin2013conducting]. All respondents
+were purposively selected based on availability, willingness to provide
+written, informed consent, and their individual insight into the
+questions posed within the study. Interviews were conducted in the local
+language (Chichewa), audio recorded, and transcribed into English.
 Participation was voluntary, and responses were recorded anonymously.
-Written, informed consent was obtained from each respondent prior to the interview.
-The study was approved by the National Committee on Research in the Social Sciences and Humanities (NCRSH) of Malawi; Protocol NO.
-P.03/19/356.
-Collected data were analysed thematically and was coded within the software programme Nvivo, which organises materials and assists with the coding process.
+Written, informed consent was obtained from each respondent prior to the
+interview. The study was approved by the National Committee on Research
+in the Social Sciences and Humanities (NCRSH) of Malawi; Protocol NO.
+P.03/19/356. Collected data were analysed thematically and was coded
+within the software programme Nvivo, which organises materials and
+assists with the coding process.
 
@@
 
-Though extensive and useful in understanding the conditions at QECH, the data set would benefit from a comparison with background values.
-As trash-burning is extensive throughout the city, there were no obvious locations that could be confidently used as representative background values.
-And although there were several possible areas upwind and outside of town, the logistics and safety involved in accessing them to change the batteries meant that the measured values at Queens could not be compared to a stable background concentration.
-Knowing the wind direction and velocity would have helped in better identifying the potential sources of the measured PM.
-However, the necessary equipment was not easily available in Blantyre at the time, and as an exploratory study, the immediate hazards for the staff and patients at QECH are relevant regardless of the burning source.
+Though extensive and useful in understanding the conditions at QECH, the
+data set would benefit from a comparison with background values. As
+trash-burning is extensive throughout the city, there were no obvious
+locations that could be confidently used as representative background
+values. And although there were several possible areas upwind and
+outside of town, the logistics and safety involved in accessing them to
+change the batteries meant that the measured values at Queens could not
+be compared to a stable background concentration. Knowing the wind
+direction and velocity would have helped in better identifying the
+potential sources of the measured PM. However, the necessary equipment
+was not easily available in Blantyre at the time, and as an exploratory
+study, the immediate hazards for the staff and patients at QECH are
+relevant regardless of the burning source.
 
-Knowing the exact burning locations and burning times, especially at the incinerator and the Guardian Shelter, would have helped in better identifying the sources and movements of the plumes.
-However, given the size of the campus and it's 24-hour schedule, a much larger research team would have been required to quantify all the burning.
-Finally, The humidity, and the potential for interference was not accounted for; the measurements were not adjusted for humidity which could affect the values, though not enough to significantly alter the key findings and need for immediate action.
+Knowing the exact burning locations and burning times, especially at the
+incinerator and the Guardian Shelter, would have helped in better
+identifying the sources and movements of the plumes. However, given the
+size of the campus and it's 24-hour schedule, a much larger research
+team would have been required to quantify all the burning. Finally, The
+humidity, and the potential for interference was not accounted for; the
+measurements were not adjusted for humidity which could affect the
+values, though not enough to significantly alter the key findings and
+need for immediate action.
 
@@
 
-The Malawi Bureau of Standards is the national agency responsible for setting and publishing all standards in the country.
-At the time of writing, the official website (www.mbsmw.org) was not available.
-However, published work that references Malawian standards [@kutlarjoss2017time; @mapoma2014air] indicate that the maximum 24-h PM10 value is 25 μg/m3 and the maximum annual PM2.5 value is 8 μg/m3.
-There is no daily maximum value for PM2.5.
-It should be noted that both of these values are at, or below, the level-4 interim WHO targets.
+The Malawi Bureau of Standards is the national agency responsible for
+setting and publishing all standards in the country. At the time of
+writing, the official website (www.mbsmw.org) was not available.
+However, published work that references Malawian standards
+[@kutlarjoss2017time; @mapoma2014air] indicate that the maximum 24-h
+PM10 value is 25 μg/m3 and the maximum annual PM2.5 value is 8 μg/m3.
+There is no daily maximum value for PM2.5. It should be noted that both
+of these values are at, or below, the level-4 interim WHO targets.
 
-Therefore, for international comparisons and due to the lack of comprehensive Malawian Standards, we use the WHO recommendations as a basis of comparison for the measured values [@worldhealthorganization2021who].
-The relevant particulate matter targets are presented in Table 1.
+Therefore, for international comparisons and due to the lack of
+comprehensive Malawian Standards, we use the WHO recommendations as a
+basis of comparison for the measured values
+[@worldhealthorganization2021who]. The relevant particulate matter
+targets are presented in Table 1.
 
@@
 
-Due to localised variation (air movement in the immediate area) and the density of the observations, the plotted 5-minute data obscure persistent trends.
-The full set of 5-minute data are plotted and presented in Appendix XA1 (TODO).
-However, the number of measured values that exceeded the hazardous limit for both parameters are summarised in Table 2.
+Due to localised variation (air movement in the immediate area) and the
+density of the observations, the plotted 5-minute data obscure
+persistent trends. The full set of 5-minute data are plotted and
+presented in Appendix XA1 (TODO). However, the number of measured values
+that exceeded the hazardous limit for both parameters are summarised in
+Table 2.
 
@@
 
-Over the course of the two-month study, the hazardous limits for both parameters were exceeded at all locations except for the Lions Sight hospital.
-The limits were exceeded fewer than 50 times at five of the locations, and only the monitors at the Lighthouse Clinic and the Guardian Shelter recorded more than 50 instances above the hazardous limits.
-At the Lighthouse Clinic, the PM10 limit was exceeded almost twice as often as the PM2.5 limit, while at the Guardian Shelter, the reverse is true.
-Because the Lighthouse Clinic is within 50 m of the incinerator, the number and predominance of PM10 peaks is characteristic of incomplete combustion and dust that is typical for the area and the open burning that occurs there.
-The residents at the Guardian Shelter however are mostly cooking within contained clay stoves that have been designed to improve combustion of the coal and wood that is burned within them.
-The fewer number of PM10 peaks is a testament to this intervention, though the frequency of PM2.5 peaks is still beyond acceptable.
+Over the course of the two-month study, the hazardous limits for both
+parameters were exceeded at all locations except for the Lions Sight
+hospital. The limits were exceeded fewer than 50 times at five of the
+locations, and only the monitors at the Lighthouse Clinic and the
+Guardian Shelter recorded more than 50 instances above the hazardous
+limits. At the Lighthouse Clinic, the PM10 limit was exceeded almost
+twice as often as the PM2.5 limit, while at the Guardian Shelter, the
+reverse is true. Because the Lighthouse Clinic is within 50 m of the
+incinerator, the number and predominance of PM10 peaks is characteristic
+of incomplete combustion and dust that is typical for the area and the
+open burning that occurs there. The residents at the Guardian Shelter
+however are mostly cooking within contained clay stoves that have been
+designed to improve combustion of the coal and wood that is burned
+within them. The fewer number of PM10 peaks is a testament to this
+intervention, though the frequency of PM2.5 peaks is still beyond
+acceptable.
 
@@
 
-The 24-hour averaged values (logged every 5 minutes) are presented for each location, for both PM10 and PM2.5 in Figure 3.
-As well as dampening local variation and peaks, the WHO air quality guidelines are also based on 24-hour averages which are the standard against which the health risks can be judged.
+The 24-hour averaged values (logged every 5 minutes) are presented for
+each location, for both PM10 and PM2.5 in Figure 3. As well as dampening
+local variation and peaks, the WHO air quality guidelines are also based
+on 24-hour averages which are the standard against which the health
+risks can be judged.
 
-Overall, PM2.5 values remained below 100 μg/m3 at 6 of 8 locations (Administration, Lions Sight, Malaria Project, uMoyo, Ward 6B, and Mercy James); PM10 values were consistently below 100μg/m3 at the same locations, but with several average values extending slightly above, and then infrequently.
+Overall, PM2.5 values remained below 100 μg/m3 at 6 of 8 locations
+(Administration, Lions Sight, Malaria Project, uMoyo, Ward 6B, and Mercy
+James); PM10 values were consistently below 100μg/m3 at the same
+locations, but with several average values extending slightly above, and
+then infrequently.
 
-The daily averages at both the Lighthouse Clinic and the Guardian Shelter are both consistently higher for both parameters and the two averages closely followed the same general trends.
-Though the peaks at the Lighthouse Clinic were higher than the Guardian Shelter, the low values were consistently lower, indicating more times of little or no burning, unlike the Guardian Shelter emissions which were relatively constant.
-However, further analysis of 12 hour averages (8:00-15:59 (working hours) and 16:00-7:59 (evening)) did not indicate clear differences between the time periods; stated differently, the data did not clearly point to more or less burning in the day or at night.
+The daily averages at both the Lighthouse Clinic and the Guardian
+Shelter are both consistently higher for both parameters and the two
+averages closely followed the same general trends. Though the peaks at
+the Lighthouse Clinic were higher than the Guardian Shelter, the low
+values were consistently lower, indicating more times of little or no
+burning, unlike the Guardian Shelter emissions which were relatively
+constant. However, further analysis of 12 hour averages (8:00-15:59
+(working hours) and 16:00-7:59 (evening)) did not indicate clear
+differences between the time periods; stated differently, the data did
+not clearly point to more or less burning in the day or at night.
 
@@
 
-Given that the type of fuel and the type of burning (contained vs.
-open) produce very different particulate "fingerprints", the ratio of PM10 to PM2.5 values were examined to determine if a clear difference between locations, and therefore source, could be identified.
-The results are presented in Figure 4.
+Given that the type of fuel and the type of burning (contained vs. open)
+produce very different particulate "fingerprints", the ratio of PM10 to
+PM2.5 values were examined to determine if a clear difference between
+locations, and therefore source, could be identified. The results are
+presented in Figure 4.
 
@@
 
-The results presented in Figure 5 illustrate both the presence of PM10 relative to PM2.5 as well as the range of values over which that ratio is observed.
+The results presented in Figure 5 illustrate both the presence of PM10
+relative to PM2.5 as well as the range of values over which that ratio
+is observed.
 
-Unlike the peak values presented in Table 2, the values presented in Figure 5 indicate the relative presence of larger (PM10) particles to smaller ones (PM2.5) regardless of their concentration: the ratio of two low concentrations can be the same as the ratio of two high concentrations and is therefore more indicative of source than of proximity.
+Unlike the peak values presented in Table 2, the values presented in
+Figure 5 indicate the relative presence of larger (PM10) particles to
+smaller ones (PM2.5) regardless of their concentration: the ratio of two
+low concentrations can be the same as the ratio of two high
+concentrations and is therefore more indicative of source than of
+proximity.
 
-In general, the calculated ratios are concentrated between 1.25 with few values exceeding 7.5.
-Three from each the Guardian Shelter and the Ward 6B exceeded 7.5 though the most values exceeding 5 were at the Guardian Shelter.
-It is interesting to note that despite the large daily variations at the Lighthouse Clinic, the parameter ratios there were fairly consistent with few spikes, and only 2 days with ratios above 5.
+In general, the calculated ratios are concentrated between 1.25 with few
+values exceeding 7.5. Three from each the Guardian Shelter and the Ward
+6B exceeded 7.5 though the most values exceeding 5 were at the Guardian
+Shelter. It is interesting to note that despite the large daily
+variations at the Lighthouse Clinic, the parameter ratios there were
+fairly consistent with few spikes, and only 2 days with ratios above 5.
 
-An examination of the range of ratios is used to better understand the variability within a single day.
-For example, several of the calculated values at the Guardian Shelter have a range of more than 5, and one day had a range of close to 10 i.e.
-the minimum ratio recorded for that day was close to 1 and the maximum value was over 10.
-The composition of emissions recorded at that location ranged significantly, and likely reflects a range of burning styles and/or fuel type.
+An examination of the range of ratios is used to better understand the
+variability within a single day. For example, several of the calculated
+values at the Guardian Shelter have a range of more than 5, and one day
+had a range of close to 10 i.e. the minimum ratio recorded for that day
+was close to 1 and the maximum value was over 10. The composition of
+emissions recorded at that location ranged significantly, and likely
+reflects a range of burning styles and/or fuel type.
 
@@
 
-The relative amount of time exposed to a given category of air quality is shown in Figure 5. Each square represents 1% of the measured time.
+The relative amount of time exposed to a given category of air quality
+is shown in Figure 5. Each square represents 1% of the measured time.
 
@@
 
-Unsurprisingly, a majority of the air quality at the Lighthouse Clinic and the Guardian Shelter falls at, or below, interim target 1.
-What is less expected is that even though there are fewer days of hazardous or very bad air quality at the other locations, most locations do not have target 4 values for even half of the time.
-Mercy James, the Malaria Project and Lions Sight have the largest percentages of days with air quality meeting the interim target 4.
-Interim target 2 rated air occupies an unexpectedly large proportion of the days at all locations, indicating that although there are a few very toxic, and a few quite good days, the majority of the air is actually neither, but still cause for concern.
+Unsurprisingly, a majority of the air quality at the Lighthouse Clinic
+and the Guardian Shelter falls at, or below, interim target 1. What is
+less expected is that even though there are fewer days of hazardous or
+very bad air quality at the other locations, most locations do not have
+target 4 values for even half of the time. Mercy James, the Malaria
+Project and Lions Sight have the largest percentages of days with air
+quality meeting the interim target 4. Interim target 2 rated air
+occupies an unexpectedly large proportion of the days at all locations,
+indicating that although there are a few very toxic, and a few quite
+good days, the majority of the air is actually neither, but still cause
+for concern.
 
@@
 
-As the previous section has described, the open incineration of waste at QECH has created hazardous conditions for those occupying the space.
-These risks were not lost on staff and caregivers, as interviews demonstrated broad and near universal awareness of the poor air quality within the hospital grounds.
-Overwhelmingly, within interviews, both caregivers and staff were quick to decry the poor quality of the air, generally without prompting.
-The few exceptions were those staff posted on the peripheries of the hospital grounds, at a distance from the spaces of incineration.
-However, even those who did not experience the impacts of the burning, were aware of it, and considered themselves fortunate to be posted in a section of the hospital where it was less of a problem.
-Furthermore, according to caregivers and staff who work night shifts, air quality can be particularly bad at night and in the early morning, despite the smoke being less visible, due to the habit of janitorial staff concentrating their burning during the late hours.
-The poor air quality on hospital grounds is also a frequent cause for complaint by patients and visitors, with nearly every staff member interviewed being able to recall having received a complaint, and in turn, complaining to administration.
-One of the staff members responsible for the burning said that he personally, had received hundreds of complaints, but was powerless to affect meaningful change, aside from burning at different hours, until the incinerator could be repaired.
+As the previous section has described, the open incineration of waste at
+QECH has created hazardous conditions for those occupying the space.
+These risks were not lost on staff and caregivers, as interviews
+demonstrated broad and near universal awareness of the poor air quality
+within the hospital grounds. Overwhelmingly, within interviews, both
+caregivers and staff were quick to decry the poor quality of the air,
+generally without prompting. The few exceptions were those staff posted
+on the peripheries of the hospital grounds, at a distance from the
+spaces of incineration. However, even those who did not experience the
+impacts of the burning, were aware of it, and considered themselves
+fortunate to be posted in a section of the hospital where it was less of
+a problem. Furthermore, according to caregivers and staff who work night
+shifts, air quality can be particularly bad at night and in the early
+morning, despite the smoke being less visible, due to the habit of
+janitorial staff concentrating their burning during the late hours. The
+poor air quality on hospital grounds is also a frequent cause for
+complaint by patients and visitors, with nearly every staff member
+interviewed being able to recall having received a complaint, and in
+turn, complaining to administration. One of the staff members
+responsible for the burning said that he personally, had received
+hundreds of complaints, but was powerless to affect meaningful change,
+aside from burning at different hours, until the incinerator could be
+repaired.
 
-Nonetheless, despite this consensus that air quality caused by the burning of waste within the hospital was a problem, significant differences emerged between respondents over their understandings of potential impacts, the effectiveness of various coping mechanisms, and their problematisations linked to the burning of specific waste materials.
-In addition, interviews revealed that these understandings were informed by a significant amount of misinformation, even amongst trained medical staff, which may contribute to them being less able to mitigate potential risks for themselves and those who rely on their care.
+Nonetheless, despite this consensus that air quality caused by the
+burning of waste within the hospital was a problem, significant
+differences emerged between respondents over their understandings of
+potential impacts, the effectiveness of various coping mechanisms, and
+their problematisations linked to the burning of specific waste
+materials. In addition, interviews revealed that these understandings
+were informed by a significant amount of misinformation, even amongst
+trained medical staff, which may contribute to them being less able to
+mitigate potential risks for themselves and those who rely on their
+care.
 
@@
 
-The poor air quality within QECH was responsible, according to respondents, for a wide array of health impacts.
-The most common ones cited included: coughing and sneezing, sore throat, stinging eyes, breathing difficulties, and persistent cold and flu.
-Nausea was also mentioned, but was not a commonly described impact.
-Only one respondent, of the 26 total interviewed, did not describe lingering health impacts which they could ascribe to the smoke, however, they also described having chronic eye irritation, but did not believe the smoke was a contributing factor.
+The poor air quality within QECH was responsible, according to
+respondents, for a wide array of health impacts. The most common ones
+cited included: coughing and sneezing, sore throat, stinging eyes,
+breathing difficulties, and persistent cold and flu. Nausea was also
+mentioned, but was not a commonly described impact. Only one respondent,
+of the 26 total interviewed, did not describe lingering health impacts
+which they could ascribe to the smoke, however, they also described
+having chronic eye irritation, but did not believe the smoke was a
+contributing factor.
 
-In addition to these impacts, which respondents bear on a daily basis, many also believed that the smoke could contribute to a number of more serious, long-term health complications.
-For instance, nearly a quarter of respondents raised concerns of the potential impact that the smoke could have on patients or staff with asthma.
-Others flagged poor air quality as a potential risk factor for certain cancers, lung disease, or heart problems.
-For a few, the smoke posed an unknown danger, they were not sure what types of impacts it could have, but they were sure it was harmful in some way.
+In addition to these impacts, which respondents bear on a daily basis,
+many also believed that the smoke could contribute to a number of more
+serious, long-term health complications. For instance, nearly a quarter
+of respondents raised concerns of the potential impact that the smoke
+could have on patients or staff with asthma. Others flagged poor air
+quality as a potential risk factor for certain cancers, lung disease, or
+heart problems. For a few, the smoke posed an unknown danger, they were
+not sure what types of impacts it could have, but they were sure it was
+harmful in some way.
 
-Also, understandably, given the large tuberculosis ward present within the hospital grounds, and the high prevalence of the disease within Malawi [@worldhealthorganization2019global] there was significant concern (more than half of respondents) about the impact the poor air quality could have on the infected.
-However, there also persisted a belief among several respondents, including several nurses, that the smoke could be a cause of the disease itself.
-As one staff member (20/11/2019) stated, "I believe breathing this air for a long time can cause Tuberculosis." This, however, was only one of the few instances of misinformation which staff members held regarding air quality and health.
-Another example, voiced by several respondents, included a belief that some staff members were immune to ill-health impacts of the smoke, because they had received vaccinations from the hospital.
-Nonetheless, they were concerned about the impacts of the smoke on patients and visitors, as one staff member (18/11/2019) expressed:
+Also, understandably, given the large tuberculosis ward present within
+the hospital grounds, and the high prevalence of the disease within
+Malawi [@worldhealthorganization2019global] there was significant
+concern (more than half of respondents) about the impact the poor air
+quality could have on the infected. However, there also persisted a
+belief among several respondents, including several nurses, that the
+smoke could be a cause of the disease itself. As one staff member
+(20/11/2019) stated, "I believe breathing this air for a long time can
+cause Tuberculosis." This, however, was only one of the few instances of
+misinformation which staff members held regarding air quality and
+health. Another example, voiced by several respondents, included a
+belief that some staff members were immune to ill-health impacts of the
+smoke, because they had received vaccinations from the hospital.
+Nonetheless, they were concerned about the impacts of the smoke on
+patients and visitors, as one staff member (18/11/2019) expressed:
 
-> Personally, I have never experienced \[eye discomfort\] because I get vaccinated and I am protected including other staff.
-However, we realize that the air can badly affect other people and patients who come to this place.
+> Personally, I have never experienced \[eye discomfort\] because I get
+> vaccinated and I am protected including other staff. However, we
+> realize that the air can badly affect other people and patients who
+> come to this place.
 
-Another interesting misconception that emerged, which may be tied partly to translation and transcription, was a different cultural understanding of smoke versus smells.
-More than half of the respondents appeared to conflate the two, with some expressing a belief that it was the odour of what was being burnt that was harmful, rather than the smoke being given off.
-This has led many to specifically problematise the burning of certain wastes, such as plastics, medicines, and other medical wastes, which give off distinctive or less pleasant odours, as opposed to the burning of other items, such as garden refuse, which may produce significant smoke, and contribute to higher recorded values of particulate matter, but produce a less pungent, or more normalised, odour.
+Another interesting misconception that emerged, which may be tied partly
+to translation and transcription, was a different cultural understanding
+of smoke versus smells. More than half of the respondents appeared to
+conflate the two, with some expressing a belief that it was the odour of
+what was being burnt that was harmful, rather than the smoke being given
+off. This has led many to specifically problematise the burning of
+certain wastes, such as plastics, medicines, and other medical wastes,
+which give off distinctive or less pleasant odours, as opposed to the
+burning of other items, such as garden refuse, which may produce
+significant smoke, and contribute to higher recorded values of
+particulate matter, but produce a less pungent, or more normalised,
+odour.
 
@@
 
-Finally, in order to manage the impacts of QECH's persistently poor air quality, staff and visitors reported having developed a number of coping mechanisms, designed to help them get through their daily routines.
-These included staying indoors, blocking doors and windows, and taking breaks away from hospital grounds in order to catch some breaths of fresher air.
-However, for janitorial staff, inside was not necessarily better, as several reported their indoor workspaces as being also insufferable for long periods of time from the smell of cleaning agents and other chemicals.
-Furthermore, amongst respondents there was a general disagreement over the effectiveness of personal protective equipment (PPE), such as face masks, towards mitigating the impacts of the smoke.
-A few staff described pleading to hospital administration for such equipment, to no avail.
-However, other staff members, who do have access to PPE, noted that even face masks do little to mitigate the impacts of the smoke, describing them as ineffective.
+Finally, in order to manage the impacts of QECH's persistently poor air
+quality, staff and visitors reported having developed a number of coping
+mechanisms, designed to help them get through their daily routines.
+These included staying indoors, blocking doors and windows, and taking
+breaks away from hospital grounds in order to catch some breaths of
+fresher air. However, for janitorial staff, inside was not necessarily
+better, as several reported their indoor workspaces as being also
+insufferable for long periods of time from the smell of cleaning agents
+and other chemicals. Furthermore, amongst respondents there was a
+general disagreement over the effectiveness of personal protective
+equipment (PPE), such as face masks, towards mitigating the impacts of
+the smoke. A few staff described pleading to hospital administration for
+such equipment, to no avail. However, other staff members, who do have
+access to PPE, noted that even face masks do little to mitigate the
+impacts of the smoke, describing them as ineffective.
 
-Most staff, however, have been unable to find anyway to mitigate the impacts of the smoke, and only found relief once they reached home at the end of their shift, as one of the janitorial staff described, "we only feel safe when we are home".
-Of course, this relief is not available for the hundreds of patients and caregivers who sleep at QECH or are unable to leave.
-Ultimately, most place their hope in the construction of the new incinerator (which had not yet been activated at the time of the interviews), and biding their time as construction drags on; coping as best they can.
-As one staff member (25/11/2019) described,"we are just hoping we will start breathing good air soon, when the new incinerator is opened."
+Most staff, however, have been unable to find anyway to mitigate the
+impacts of the smoke, and only found relief once they reached home at
+the end of their shift, as one of the janitorial staff described, "we
+only feel safe when we are home". Of course, this relief is not
+available for the hundreds of patients and caregivers who sleep at QECH
+or are unable to leave. Ultimately, most place their hope in the
+construction of the new incinerator (which had not yet been activated at
+the time of the interviews), and biding their time as construction drags
+on; coping as best they can. As one staff member (25/11/2019)
+described,"we are just hoping we will start breathing good air soon,
+when the new incinerator is opened."
 

### 2023-04-05 manuscript/manuscript.qmd
Add appendices

@@
 at the hospital's incinerator, located at a central point within the
-hospital campus [@kalina2020this]. However, the vast majority of the
+hospital campus[^1] [@kalina2020this]. However, the vast majority of the
 deposited waste is not incinerated, as a new incinerator provided by the
@@
 percentage of the hospital's waste. Furthermore, the incinerator is
-frequently inoperable, under repair, or without power,. Rather, the
+frequently inoperable, under repair, or without power[^2]. Rather, the
 waste is openly burned, in a constantly smouldering pile to the side of
@@
 grounds and maintenance staff habitually burn leaves and other garden
-refuse. Furthermore, caregivers who reside separately on hospital
+refuse. Furthermore, caregivers[^3] who reside separately on hospital
 grounds, cook for themselves throughout the day and night using biomass
@@
 is, as the quote used in the title of this article, 'unbearable.'
+
+[^1]: Sharps are handled and disposed of separately.
+
+[^2]: There was an older incinerator on the grounds which failed in
+    2017. Between 2017 and 2021, when the new incinerator was
+    commissioned, nothing was incinerated at all.
+
+[^3]: Patients are required to bring a caregiver for in-patient
+    hospitals stays, usually a family member, who assist with feeding,
+    bathing, and all other non-medical services.
 
@@
     research and patient care in the area of paediatric malaria,
-    specifically cerebral malaria, for more than 30 years. Its research
-    infrastructure includes an administrative team, an inpatient
-    research ward, an MRI centre, and an outpatient research clinic in
-    Ndirande township and in other districts outside of Blantyre.
+    specifically cerebral malaria, for more than 30 years[^4]. Its
+    research infrastructure includes an administrative team, an
+    inpatient research ward, an MRI centre, and an outpatient research
+    clinic in Ndirande township and in other districts outside of
+    Blantyre.
 
@@
     concrete floor shelter for sleeping along with a cooking pavilion
-    and toilet/shower block are maintained by a local NGO. Though
-    simple, the facility is secure and offers one of the only
-    free/accessible toilet facilities for visitors on the QECH campus.
+    and toilet/shower block are maintained by a the local NGO Chira
+    fund. Though simple, the facility is secure and offers one of the
+    only free/accessible toilet facilities for visitors on the QECH
+    campus.
+
+[^4]: BMP is credited with the development of the Blantyre Coma Score,
+    which is used widely.
 
@@
 micrometres in diameter in mg/m3. A list of the specific hardware and
-software components can be found in Appendix TODO.
+software components can be found in Appendix B.
 
@@
 persistent trends. The full set of 5-minute data are plotted and
-presented in Appendix XA1 (TODO). However, the number of measured values
+presented in Appendix A Figure A.1 (TODO). However, the number of measured values
 that exceeded the hazardous limit for both parameters are summarised in
@@
 
-# Appendix
+# Appendix {.unnumbered}
 
@@
 etc.
+
+## Appendix A {.unnumbered}
+
+```{r}
+
+# Figure XA1. PM2.5 and PM10 values collected every 5 minutes over 10 months at 8 locations. 
+# TODO: ##Lars can you please fix the x-axis labels so they aren’t squished
+
+```
+
+## Appendix B {.unnumbered}
+
+### Hardware Components {.unnumbered}
+
+1.  Raspberry Pi 3 Model B single-board computer technical
+    specifications:
+    -   Broadcom BCM2837 64bit Quad Core Processor powered Single Board
+        Computer running at 1.2GHz 1GB RAM
+    -   BCM43438 WiFi on board
+    -   Bluetooth Low Energy (BLE) on board
+    -   40pin extended GPIO
+    -   4 x USB 2 ports 4 pole Stereo output and Composite video port
+    -   Full size HDMI CSI camera port for connecting the Raspberry Pi
+        camera
+    -   DSI display port for connecting the Raspberry Pi touch screen
+        display
+    -   Micro SD port for loading your operating system and storing data
+    -   Upgraded switched Micro USB power source (now supports up to 2.4
+        Amps)
+    -   Same form factor as the Pi 2 Model B, however the LEDs have
+        changed position
+    -   RPI3-MODB-16GB-NOOBS
+2.  Nova PM SDS011 Sensor
+    -   See Data Sheet attachment for technical specifications
+3.  MCP3008 microchip
+    -   analog-to-digital converter (ADC)
+4.  Romoss Sense8+ 30000mAh QC Type-C
+    -   Power bank battery
+5.  Pipe
+    -   clear rubber pipe attainable from any local hardware store
+6.  Box
+    -   any box that can fit all the components comfortbaly can be used.
+        The box will need to have a whole drilled in it for the pipe.
+
+### Software ‘components’ and data management {.unnumbered}
+
+1. The Raspberry Pi operating system runs off of Rasbian software on a NOOBS scandisk. 
+2.
+Air quality monitoring code is written in Python using the [aqi library](https://pypi.org/project/python-aqi/) (A library of algorithms to convert between AQI value and pollutant concentration).
+The code is stored on the [Open Data Durban GitHub account](https://github.com/opendatadurban/hospital_stations) and is free for all to access.
+To be able to collect data, the device needs to be connected to a reliable wifi network, the SSID and password need to be coded onto the device during step 2 in the building method.
+3. Once the whole device has been put together and placed on site, to access that data we use a combination of propgrams:
+    - [Putty](https://www.putty.org/) - this is an SSH client that we use to gain access to the Raspberry Pi from another computer.
+To be able to access the Raspberry Pi in question, the computer in use and the air quality monitoring device need to be on the same network for the connection to be successful.
+    - [Winscp](https://winscp.net/eng/download.php) - this is an FTP client that we use, in conjunction with Putty, to be able to transfer data files from the air quality monitoring device to one’s computer.
+    - [Remote.it](https://app.remote.it/) - this is an application that supports remote SSH, so this is used when wanting to access data from a device that is not connected to the same network as the computer in use.
+Remote.it is used together with Putty and Winscp to access the air quality data remotely.
+
+### Connection/building method {.unnumbered}
+
+To build the sensor, we followed these steps:
+
+1. Set up the Raspberry Pi with the Rasbian software using the NOOBS scandisk. (insert the scandisk into the Rasberry Pi) A comprehensive start up guide can be found here. 
+2.
+Next we load the code onto the Raspberry Pi.
+The code can be found on the Open Data Durban Hospital Stations Repository Github.
+To use it you can follow the steps in the file called Initiate_WS.txt.
+Once all the steps have been completed, the Raspberry Pi is now a device that can be connected to a particulate matter monitoring sensor.
+3. Connect the hardware components
+    - Connect the Raspberry Pi to the Nova PM Sensor using the break out board that comes with the Sensor. 
+    - Connect the battery to the Raspberry Pi via the micro port
+    - Attach the rubber pipe to the sensor 
+    - Assemble all components inside the box, with the pipe coming out of the drilled hole
+    - To prevent dirt entering the system, close off all openings between the pipe and the box with silicone putty
+    - Secure box in an elevated position (not too close to the ground) and with the pipe facing the direction of the inflow of the air under investigation.
 

### 2023-04-05 R/01-data-preparation.R
Add data preparation script

@@ -1,0 +1,173 @@
+# description -------------------------------------------------------------
+
+# setup -------------------------------------------------------------------
+
+# R packages on CRAN
+
+library(readr)
+library(dplyr)
+library(forcats)
+library(lubridate)
+library(ggplot2)
+library(janitor)
+library(tidyr)
+library(leaflet)
+library(mapview)
+
+# R packages not on CRAN
+
+# install.packages("devtools")
+# devtools::install_github("hrbrmstr/waffle")
+library(waffle)
+
+# read data ---------------------------------------------------------------
+
+# devtools::install_github("Global-Health-Engineering/qechairquality")
+library(qechairquality)
+
+# data preparation --------------------------------------------------------
+
+## prepare a vector of names for sensor locations
+
+location_vec <- c("Ward 6B",
+                  "Guardian Shelter",
+                  "Lighthouse Clinic",
+                  "Lions Sight",
+                  "Malaria Project",
+                  "Mercy James",
+                  "uMoyo")
+
+## read data from package and manipulate variables
+
+dat_in <- qechairquality |>
+  # move to data cleaning process
+  mutate(date = as_date(date_time)) |>
+  mutate(hour = hour(date_time)) |>
+  # remove values above 1000, as they could not plausibly be this high
+  filter(value <= 1000) |>
+  # not all sensors were installed on the same day or stopped
+  # collecting data on the same day.
+  # define a daterange for entire dataset
+  filter(date_time >= "2019-10-05 00:00:00",
+         date_time <= "2019-11-25 00:00:00") |>
+  ## rename locations to meaningful names
+  mutate(location = case_when(
+    location == "6B" ~ location_vec[[1]],
+    location == "guardian" ~ location_vec[[2]],
+    location == "Lhouse" ~ location_vec[[3]],
+    location == "Lions" ~ location_vec[[4]],
+    location == "MJC" ~ location_vec[[6]],
+    location == "Moyo" ~ location_vec[[7]],
+    TRUE ~ location
+  ))
+
+# Prepare hospital location vector
+
+vec_hospital_name <- dat_in |>
+  count(location) |>
+  pull(location)
+
+# define US EPA levels for air quality
+usepa_level_fct <- c("Good",
+                     "Moderate",
[…]

### 2023-04-05 manuscript/manuscript.qmd
Figures and tables added

@@
         - Highlight 3
+execute: 
+  echo: false
+  warning: false
 editor: 
@@
     wrap: 72
+editor_options: 
+  chunk_output_type: console
 ---
+
+```{r setup}
+
+# Source data preparation script
+source(here::here("R/01-data-preparation.R"))
+
+```
 
@@
 the incinerator building, sending up the aforementioned clouds of smoke,
-which choke the grounds (Figure 1). Innumerable other, smaller, fires,
-scattered across the QECH campus (Figure 2) contribute their part, as
+which choke the grounds (@fig-open-burning). Innumerable other, smaller, fires,
+scattered across the QECH campus (@fig-map-sensors) contribute their part, as
 grounds and maintenance staff habitually burn leaves and other garden
@@
 ```{r}
-# Figure 1: Open burning at QECH, with the old incinerator building in the background (Authors).
-knitr::include_graphics(here::here("manuscript/img/image-replace-tree.png"))
+#| label: fig-open-burning
+#| fig-cap: Open burning at QECH, with the old incinerator building in the background (Authors).
+#| fig-align: center
+#| out-width: 80%
+
+knitr::include_graphics(here::here("manuscript/img/open-waste-burning-incinerator-qech.jpg"))
 ```
@@
 ```{r}
-# map
+#| label: prepare-sensor-map
+
+# Icon from: https://github.com/apancik/public-domain-icons/blob/master/dist/symbol%20fire.svg
+
+fire_icon <- makeIcon(iconUrl = here::here("manuscript/img/symbol-fire.png"),
+                      iconWidth = 50,
+                      iconHeight = 50)
+
+# Create a palette that maps factor levels to colors
+pal <- colorFactor(c("#EE4B2B", "navy"), domain = c("sensor", "fire"))
+
+fire_loc <- locations  |>
+  filter(type == "fire")
+
+sensor_loc <- locations  |>
+  filter(type == "sensor")
+
+qech_sensor_map <- leaflet(sensor_loc)  |>
+  setView(lng = 35.021071999805755,
+          lat = -15.803269709021687,
+          zoom = 18)  |>
+  #addProviderTiles(providers$Stamen)  |>
+  addTiles()  |>
+  addMarkers(
+    lng = fire_loc$long,
+    lat = fire_loc$lat,
+    label = fire_loc$location,
+    icon = ~fire_icon,
+    labelOptions = labelOptions(noHide = T,
+                                textsize = "15px",
+                                direction = "bottom",
+                                offset = c(0, 20),
+                                style = list(
+                                  "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
+                                  "border-color" = "rgba(0,0,0,0.5)"
+                                  #"font-style" = "italic"
+                                ))
+  )  |>
+  addCircleMarkers(
+    lng = ~long,
+    lat = ~lat,
+    label = ~location,
+    radius = 16,
+    stroke = FALSE,
+    fillOpacity = 1,
+    color = ~pal(type),
+    #icon = ~fire_icon,
+    labelOptions = labelOptions(noHide = T,
+                                textsize = "15px",
+                                direction = "bottom",
+                                offset = c(0, 10),
+                                style = list(
+                                  "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
+                                  "border-color" = "rgba(0,0,0,0.5)"
+                                  #"font-style" = "italic"
+                                )))
+
+qech_sensor_map |> 
+  mapshot(file = here::here("manuscript/figs/map-sensors-qech-blantyre.png"))
+```
+
+```{r}
+#| label: fig-map-sensors
+#| fig-cap: Sensor locations within QECH
+#| fig-align: center
+#| out-width: 80%
+
 # Figure 2: Sensor locations within QECH
+knitr::include_graphics(here::here("manuscript/figs/map-sensores-qech-blantyre.png"))
 ```
@@
 [@worldhealthorganization2021who]. The relevant particulate matter
-targets are presented in Table 1.
+targets are presented in @tbl-who-targets.
+
+ (TODO: REPLACE screenshot)
 
 ```{r}
-# Table 1. Particulate matter targets from @worldhealthorganization2021who
+#| label: tbl-who-targets
+#| tbl-cap: Particulate matter targets from [@worldhealthorganization2021who]
+
 knitr::include_graphics(here::here("manuscript/img/image-replace-table.png"))
@@
 persistent trends. The full set of 5-minute data are plotted and
-presented in Appendix A Figure A.1 (TODO). However, the number of measured values
+presented in Appendix A Figure A.1. However, the number of measured values
 that exceeded the hazardous limit for both parameters are summarised in
-Table 2.
+@tbl-peaks.
 
 ```{r}
-# Table 2.
-Peaks for both PM2.5 and PM10 are the number of data points above the WHO interim target 1 (annual) which is the least stringent;  the number of observations recorded (count) is provided for reference.
+#| label: tbl-peaks
+#| tbl-cap: Peaks for both PM~2.5~ and PM~10~ are the number of data points above the WHO interim target 1 (annual) which is the least stringent; the number of observations recorded (count) is provided for reference.
 
+observations_location <- dat_in |> 
+  select(date_time, location) |> 
+  unique() |> 
+  count(location)
+
+no_peaks_who <- dat_in |> 
+  mutate(peaks = case_when(
+    indicator == "pm2.5" & value >  35 ~ TRUE,
+    indicator == "pm10" & value >  70 ~ TRUE,
+    TRUE ~ FALSE
+  )) |> 
+  mutate(date = as_date(date_time)) |> 
+  group_by(indicator, location) |> 
+  summarise(
+    peaks = sum(peaks)
+  ) |> 
+  ungroup() |> 
+  pivot_wider(names_from = indicator, values_from = peaks)
+
+observations_location |> 
+  left_join(no_peaks_who) |> 
+  knitr::kable()
 ```
@@
 The 24-hour averaged values (logged every 5 minutes) are presented for
-each location, for both PM10 and PM2.5 in Figure 3. As well as dampening
+each location, for both PM10 and PM2.5 in fig-24-hour-average. As well as dampening
 local variation and peaks, the WHO air quality guidelines are also based
@@
 ```{r}
-# Figure 3. Average 24-hour PM10 and PM2.5 at 8 monitoring stations over 2 months
+#| label: fig-24-hour-average
+#| fig-cap: Average 24-hour PM10 and PM2.5 at 8 monitoring stations over 2 months
+#| fig-align: center
+
+dat_in_sum_day |> 
+  ggplot(aes(x = date, y = mean, group = indicator, color = indicator)) +
+  geom_point(size = 0.8) +
+  geom_line() +
+  labs(x = NULL) +
+  ylab(bquote(uq/m^3)) +
+  facet_wrap(~location, ncol = 4) + 
+  scale_y_continuous(limits = c(0, 400)) +
+  scale_color_brewer(type = "qual", palette = 2) + 
+  theme_bw(base_size = 14) + 
+  theme(panel.grid.minor = element_blank(),
+        panel.grid.major.x = element_blank())
 
@@
 locations, and therefore source, could be identified. The results are
-presented in Figure 4.
+presented in @fig-pm-ratios.
 
 ```{r}
-# Figure 4. PM10:PM2.5 values for each location.  Each line shows the minimum and maximum ratio for the day based on ratios calculated hourly.  
+#| label: fig-pm-ratios
+#| fig-cap: PM10:PM2.5 values for each location.
+Each line shows the minimum and maximum ratio for the day based on ratios calculated hourly.
+Does not include 14th to 16th October at location uMoyo due to extreme outliers.
+#| fig-align: center
+
+dat_in_ratios <- dat_in |> 
+  mutate(date = as_date(date_time)) |> 
+  mutate(hour = hour(date_time)) |> 
+  group_by(date, hour, location, indicator) |> 
+  summarise(mean = mean(value)) |> 
+  pivot_wider(names_from = indicator, 
+              values_from = mean) |> 
+  mutate(p10_pm2.5_ratio = pm10/pm2.5)  |>
+
+  filter((date != "2019-10-14" & date != "2019-10-15" & date != "2019-10-16")
+         | location != "uMoyo") |> 
+  #filter(p10_pm2.5_ratio > 30)
+
+  group_by(date, location) |> 
+  summarise(min = min(p10_pm2.5_ratio),
+            max = max(p10_pm2.5_ratio))
+
+# Plot
+
+dat_in_ratios |> 
+  ggplot(aes(x = date)) +
+  geom_linerange(aes(ymin = min, ymax = max)) +
+  labs(x = "hour") +
+  facet_wrap(~location, ncol = 2)
 ```
@@
 The relative amount of time exposed to a given category of air quality
-is shown in Figure 5. Each square represents 1% of the measured time.
+is shown in @fig-percent-exposure. Each square represents 1% of the measured time.
 
 ```{r}
-# Figure 5. Percentage of measured values categorised according to WHO 2021 targets
+#| label: fig-percent-exposure
+#| fig-cap: Percentage of measured values categorised according to WHO 2021 targets
+#| fig-align: center
+
+dat_air_quality_sum_who_2021 |> 
+  #filter(indicator == "pm2.5") |> 
+  group_by(location, indicator) |> 
+  mutate(exposure_percent = round(exposure_hrs / sum(exposure_hrs) * 100, 0)) |>
+
+  # manipulate data for purpose of plot
+
+  mutate(exposure_percent = case_when(
+    location == "Administration" 
+    & indicator == "pm10"
+    & air_quality_who_annual == "Interim target 2" ~ exposure_percent - 1,
+
+    location == "Administration" 
+    & indicator == "pm2.5"
+    & air_quality_who_annual == "Interim target 3" ~ exposure_percent + 1,
+
+    location == "uMoyo" 
+    & indicator == "pm2.5"
+    & air_quality_who_annual == "Interim target 4" ~ exposure_percent - 1,
+
+    location == "Mercy James" 
+    & indicator == "pm10"
+    & air_quality_who_annual == "Interim target 4" ~ exposure_percent + 1,
+
+    TRUE ~ exposure_percent
+  )) |>
+
+  #filter(is.na(exposure_percent))
+
+ggplot(aes(fill = air_quality_who_annual, values = exposure_percent)) +
+  geom_waffle(color ="white", size = 0.5, n_rows = 5, na.rm = TRUE) +
+  coord_equal() +
+  labs(fill = "WHO 2021 targets (annual)") +
+  scale_fill_brewer(type = "div", palette = "RdYlBu", direction = -1) +
+  scale_x_discrete(expand=c(0,0)) +
+  scale_y_discrete(expand=c(0,0)) +
+  facet_grid(location ~ indicator, switch = "y") +
+
+  #facet_wrap(~location, ncol = 2) +
+  theme_minimal(base_size = 14) +
+  # help from: https://stackoverflow.com/questions/34749636/left-justify-text-from-multi-line-facet-labels
+  theme(strip.text.x = element_text(hjust = 0),
+        strip.text.y = element_text(vjust = 0, size = 4))
 ```
@@
 ```{r}
+#| label: fig-5-minute-intervals
+#| fig-cap: PM2.5 and PM10 values collected every 5 minutes over 10 months at 8 locations.
+#| fig-align: center
 
-# Figure XA1. PM2.5 and PM10 values collected every 5 minutes over 10 months at 8 locations. 
-# TODO: ##Lars can you please fix the x-axis labels so they aren’t squished
+dat_in |> 
+  ggplot(aes(x = date_time, y = value)) +
+  geom_line(alpha = 0.5) +
+  labs(title = "Hospital Air Quality",
+       subtitle = "PM2.5 and PM10 measurements in 5-min intervals at eight locations") +
+  ylab(bquote(uq/m^3)) +
+  facet_grid(indicator~location) +
+  theme_bw(base_size = 14) + 
+  theme(panel.grid.minor = element_blank(),
+        panel.grid.major.x = element_blank(),
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
 
@@
 
-# Figures and tables
-
-@fig-meaningless is generated using an R chunk.
-
-```{r}
-#| label: fig-meaningless
-#| fig-cap: A meaningless scatterplot
-#| fig-width: 5
-#| fig-height: 5
-#| fig-align: center
-#| out-width: 50%
-#| echo: false
-plot(runif(25), runif(25))
-```
-
-# Tables coming from R
-
-Tables can also be generated using R chunks, as shown in @tbl-simple
-example.
-
-```{r}
-#| label: tbl-simple
-#| tbl-cap: Caption centered above table
-#| echo: true
-knitr::kable(head(mtcars)[,1:4])
-```
-
 # References {.unnumbered}

### 2023-04-05 manuscript/manuscript.tex
Figures and tables added

@@
 \fi
-
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{241,243,245}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
-\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
 
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, boxrule=0pt, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, breakable]}{\end{tcolorbox}}\fi
 
@@
 at the hospital's incinerator, located at a central point within the
-hospital campus \citep{kalina2020this}. However, the vast majority of
-the deposited waste is not incinerated, as a new incinerator provided by
-the Ministry of Health in late 2019, is only able to handle a small
-percentage of the hospital's waste. Furthermore, the incinerator is
-frequently inoperable, under repair, or without power,. Rather, the
-waste is openly burned, in a constantly smouldering pile to the side of
-the incinerator building, sending up the aforementioned clouds of smoke,
-which choke the grounds (Figure 1). Innumerable other, smaller, fires,
-scattered across the QECH campus (Figure 2) contribute their part, as
-grounds and maintenance staff habitually burn leaves and other garden
-refuse. Furthermore, caregivers who reside separately on hospital
+hospital campus\footnote{Sharps are handled and disposed of separately.}
+\citep{kalina2020this}. However, the vast majority of the deposited
+waste is not incinerated, as a new incinerator provided by the Ministry
+of Health in late 2019, is only able to handle a small percentage of the
+hospital's waste. Furthermore, the incinerator is frequently inoperable,
+under repair, or without power\footnote{There was an older incinerator
+  on the grounds which failed in 2017. Between 2017 and 2021, when the
+  new incinerator was commissioned, nothing was incinerated at all.}.
+Rather, the waste is openly burned, in a constantly smouldering pile to
+the side of the incinerator building, sending up the aforementioned
+clouds of smoke, which choke the grounds
+(Figure~\ref{fig-open-burning}). Innumerable other, smaller, fires,
+scattered across the QECH campus (Figure~\ref{fig-map-sensors})
+contribute their part, as grounds and maintenance staff habitually burn
+leaves and other garden refuse. Furthermore, caregivers\footnote{Patients
+  are required to bring a caregiver for in-patient hospitals stays,
+  usually a family member, who assist with feeding, bathing, and all
+  other non-medical services.} who reside separately on hospital
 grounds, cook for themselves throughout the day and night using biomass
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Figure 1: Open burning at QECH, with the old incinerator building in the background (Authors).}
-\NormalTok{knitr}\SpecialCharTok{::}\FunctionTok{include\_graphics}\NormalTok{(here}\SpecialCharTok{::}\FunctionTok{here}\NormalTok{(}\StringTok{"manuscript/img/image{-}replace{-}tree.png"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{img/image-replace-tree.png}
+\begin{figure}
+
+{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{img/open-waste-burning-incinerator-qech.jpg}
 
 }
+
+\caption{\label{fig-open-burning}Open burning at QECH, with the old
+incinerator building in the background (Authors).}
 
@@
   and patient care in the area of paediatric malaria, specifically
-  cerebral malaria, for more than 30 years. Its research infrastructure
-  includes an administrative team, an inpatient research ward, an MRI
-  centre, and an outpatient research clinic in Ndirande township and in
-  other districts outside of Blantyre.
+  cerebral malaria, for more than 30 years\footnote{BMP is credited with
+    the development of the Blantyre Coma Score, which is used widely.}.
+  Its research infrastructure includes an administrative team, an
+  inpatient research ward, an MRI centre, and an outpatient research
+  clinic in Ndirande township and in other districts outside of
+  Blantyre.
 \item
@@
   gender-separated concrete floor shelter for sleeping along with a
-  cooking pavilion and toilet/shower block are maintained by a local
-  NGO. Though simple, the facility is secure and offers one of the only
-  free/accessible toilet facilities for visitors on the QECH campus.
+  cooking pavilion and toilet/shower block are maintained by a the local
+  NGO Chira fund. Though simple, the facility is secure and offers one
+  of the only free/accessible toilet facilities for visitors on the QECH
+  campus.
 \end{itemize}
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# map}
-\CommentTok{\# Figure 2: Sensor locations within QECH}
-\end{Highlighting}
-\end{Shaded}
+\begin{figure}
+
+{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{figs/map-sensores-qech-blantyre.png}
+
+}
+
+\caption{\label{fig-map-sensors}Sensor locations within QECH}
+
+\end{figure}
 
@@
 \citep{worldhealthorganization2021who}. The relevant particulate matter
-targets are presented in Table 1.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Table 1. Particulate matter targets from @worldhealthorganization2021who}
-\NormalTok{knitr}\SpecialCharTok{::}\FunctionTok{include\_graphics}\NormalTok{(here}\SpecialCharTok{::}\FunctionTok{here}\NormalTok{(}\StringTok{"manuscript/img/image{-}replace{-}table.png"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{img/image-replace-table.png}
+targets are presented in \textbf{?@tbl-who-targets}.
+
+(TODO: REPLACE screenshot)
+
+\begin{table}
+
+\caption{\textbf{?(caption)}}\begin{minipage}[t]{\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics[width=2.04in,height=\textheight]{img/image-replace-table.png}
 
@@
 
-\end{figure}
+}
+
+\end{minipage}%
+
+\end{table}
 
@@
 persistent trends. The full set of 5-minute data are plotted and
-presented in Appendix XA1 (TODO). However, the number of measured values
-that exceeded the hazardous limit for both parameters are summarised in
-Table 2.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Table 2.
-Peaks for both PM2.5 and PM10 are the number of data points above the WHO interim target 1 (annual) which is the least stringent;  the number of observations recorded (count) is provided for reference.
-}
-\end{Highlighting}
-\end{Shaded}
+presented in Appendix A Figure A.1. However, the number of measured
+values that exceeded the hazardous limit for both parameters are
+summarised in Table~\ref{tbl-peaks}.
+
+\hypertarget{tbl-peaks}{}
+\begin{longtable}[]{@{}lrrr@{}}
+\caption{\label{tbl-peaks}Peaks for both PM\textsubscript{2.5} and
+PM\textsubscript{10} are the number of data points above the WHO interim
+target 1 (annual) which is the least stringent; the number of
+observations recorded (count) is provided for reference.}\tabularnewline
+\toprule()
+location & n & pm10 & pm2.5 \\
+\midrule()
+\endfirsthead
+\toprule()
+location & n & pm10 & pm2.5 \\
+\midrule()
+\endhead
+Administration & 10900 & 171 & 94 \\
+Guardian Shelter & 11032 & 4529 & 5314 \\
+Lighthouse Clinic & 11472 & 4365 & 3891 \\
+Lions Sight & 11749 & 91 & 14 \\
+Malaria & 12031 & 617 & 578 \\
+Mercy James & 12193 & 736 & 749 \\
+Ward 6B & 11861 & 1113 & 1236 \\
+uMoyo & 10637 & 703 & 399 \\
+\bottomrule()
+\end{longtable}
 
@@
 The 24-hour averaged values (logged every 5 minutes) are presented for
-each location, for both PM10 and PM2.5 in Figure 3. As well as dampening
-local variation and peaks, the WHO air quality guidelines are also based
-on 24-hour averages which are the standard against which the health
-risks can be judged.
+each location, for both PM10 and PM2.5 in fig-24-hour-average. As well
+as dampening local variation and peaks, the WHO air quality guidelines
+are also based on 24-hour averages which are the standard against which
+the health risks can be judged.
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Figure 3. Average 24{-}hour PM10 and PM2.5 at 8 monitoring stations over 2 months}
-\end{Highlighting}
-\end{Shaded}
+\begin{figure}
+
+{\centering \includegraphics{manuscript_files/figure-pdf/fig-24-hour-average-1.pdf}
+
+}
+
+\caption{\label{fig-24-hour-average}Average 24-hour PM10 and PM2.5 at 8
+monitoring stations over 2 months}
+
+\end{figure}
 
@@
 locations, and therefore source, could be identified. The results are
-presented in Figure 4.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Figure 4. PM10:PM2.5 values for each location.  Each line shows the minimum and maximum ratio for the day based on ratios calculated hourly.  }
-\end{Highlighting}
-\end{Shaded}
+presented in Figure~\ref{fig-pm-ratios}.
+
+\begin{figure}
+
+{\centering \includegraphics{manuscript_files/figure-pdf/fig-pm-ratios-1.pdf}
+
+}
+
+\caption{\label{fig-pm-ratios}PM10:PM2.5 values for each location. Each
+line shows the minimum and maximum ratio for the day based on ratios
+calculated hourly. Does not include 14th to 16th October at location
+uMoyo due to extreme outliers.}
+
+\end{figure}
 
@@
 The relative amount of time exposed to a given category of air quality
-is shown in Figure 5. Each square represents 1\% of the measured time.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Figure 5. Percentage of measured values categorised according to WHO 2021 targets}
-\end{Highlighting}
-\end{Shaded}
+is shown in Figure~\ref{fig-percent-exposure}. Each square represents
+1\% of the measured time.
+
+\begin{figure}
+
+{\centering \includegraphics{manuscript_files/figure-pdf/fig-percent-exposure-1.pdf}
+
+}
+
+\caption{\label{fig-percent-exposure}Percentage of measured values
+categorised according to WHO 2021 targets}
+
+\end{figure}
 
@@
 \hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
+\section*{Appendix}\label{appendix}}
+\addcontentsline{toc}{section}{Appendix}
 
@@
 
-\hypertarget{figures-and-tables}{%
-\section{Figures and tables}\label{figures-and-tables}}
-
-Figure~\ref{fig-meaningless} is generated using an R chunk.
+\hypertarget{appendix-a}{%
+\subsection*{Appendix A}\label{appendix-a}}
+\addcontentsline{toc}{subsection}{Appendix A}
 
@@
 
-{\centering \includegraphics[width=0.5\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-meaningless-1.pdf}
+{\centering \includegraphics{manuscript_files/figure-pdf/fig-5-minute-intervals-1.pdf}
 
@@
 
-\caption{\label{fig-meaningless}A meaningless scatterplot}
+\caption{\label{fig-5-minute-intervals}PM2.5 and PM10 values collected
+every 5 minutes over 10 months at 8 locations.}
 
@@
 
-\hypertarget{tables-coming-from-r}{%
-\section{Tables coming from R}\label{tables-coming-from-r}}
-
-Tables can also be generated using R chunks, as shown in
-Table~\ref{tbl-simple} example.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{knitr}\SpecialCharTok{::}\FunctionTok{kable}\NormalTok{(}\FunctionTok{head}\NormalTok{(mtcars)[,}\DecValTok{1}\SpecialCharTok{:}\DecValTok{4}\NormalTok{])}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{tbl-simple}{}
-\begin{longtable}[]{@{}lrrrr@{}}
-\caption{\label{tbl-simple}Caption centered above table}\tabularnewline
-\toprule()
-& mpg & cyl & disp & hp \\
-\midrule()
-\endfirsthead
-\toprule()
-& mpg & cyl & disp & hp \\
-\midrule()
-\endhead
-Mazda RX4 & 21.0 & 6 & 160 & 110 \\
-Mazda RX4 Wag & 21.0 & 6 & 160 & 110 \\
-Datsun 710 & 22.8 & 4 & 108 & 93 \\
-Hornet 4 Drive & 21.4 & 6 & 258 & 110 \\
-Hornet Sportabout & 18.7 & 8 & 360 & 175 \\
-Valiant & 18.1 & 6 & 225 & 105 \\
-\bottomrule()
-\end{longtable}
+\hypertarget{appendix-b}{%
+\subsection*{Appendix B}\label{appendix-b}}
+\addcontentsline{toc}{subsection}{Appendix B}
+
+\hypertarget{hardware-components}{%
+\subsubsection*{Hardware Components}\label{hardware-components}}
+\addcontentsline{toc}{subsubsection}{Hardware Components}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Raspberry Pi 3 Model B single-board computer technical specifications:
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  Broadcom BCM2837 64bit Quad Core Processor powered Single Board
+  Computer running at 1.2GHz 1GB RAM
+\item
+  BCM43438 WiFi on board
+\item
+  Bluetooth Low Energy (BLE) on board
+\item
+  40pin extended GPIO
+\item
+  4 x USB 2 ports 4 pole Stereo output and Composite video port
+\item
+  Full size HDMI CSI camera port for connecting the Raspberry Pi camera
+\item
+  DSI display port for connecting the Raspberry Pi touch screen display
+\item
+  Micro SD port for loading your operating system and storing data
+\item
+  Upgraded switched Micro USB power source (now supports up to 2.4 Amps)
+\item
+  Same form factor as the Pi 2 Model B, however the LEDs have changed
+  position
+\item
+  RPI3-MODB-16GB-NOOBS
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  Nova PM SDS011 Sensor
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  See Data Sheet attachment for technical specifications
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{2}
+\tightlist
+\item
+  MCP3008 microchip
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  analog-to-digital converter (ADC)
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
+\tightlist
+\item
+  Romoss Sense8+ 30000mAh QC Type-C
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  Power bank battery
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{4}
+\tightlist
+\item
+  Pipe
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  clear rubber pipe attainable from any local hardware store
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{5}
+\tightlist
+\item
+  Box
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  any box that can fit all the components comfortbaly can be used. The
+  box will need to have a whole drilled in it for the pipe.
+\end{itemize}
+
+\hypertarget{software-components-and-data-management}{%
+\subsubsection*{Software `components' and data
+management}\label{software-components-and-data-management}}
+\addcontentsline{toc}{subsubsection}{Software `components' and data
+management}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  The Raspberry Pi operating system runs off of Rasbian software on a
+  NOOBS scandisk.
+\item
+  Air quality monitoring code is written in Python using the
+  \href{https://pypi.org/project/python-aqi/}{aqi library} (A library of
+  algorithms to convert between AQI value and pollutant concentration).
+  The code is stored on the
+  \href{https://github.com/opendatadurban/hospital_stations}{Open Data
+  Durban GitHub account} and is free for all to access. To be able to
+  collect data, the device needs to be connected to a reliable wifi
+  network, the SSID and password need to be coded onto the device during
+  step 2 in the building method.
+\item
+  Once the whole device has been put together and placed on site, to
+  access that data we use a combination of propgrams:
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  \href{https://www.putty.org/}{Putty} - this is an SSH client that we
+  use to gain access to the Raspberry Pi from another computer. To be
+  able to access the Raspberry Pi in question, the computer in use and
+  the air quality monitoring device need to be on the same network for
+  the connection to be successful.
+\item
+  \href{https://winscp.net/eng/download.php}{Winscp} - this is an FTP
+  client that we use, in conjunction with Putty, to be able to transfer
+  data files from the air quality monitoring device to one's computer.
+\item
+  \href{https://app.remote.it/}{Remote.it} - this is an application that
+  supports remote SSH, so this is used when wanting to access data from
+  a device that is not connected to the same network as the computer in
+  use. Remote.it is used together with Putty and Winscp to access the
+  air quality data remotely.
+\end{itemize}
+
+\hypertarget{connectionbuilding-method}{%
+\subsubsection*{Connection/building
+method}\label{connectionbuilding-method}}
+\addcontentsline{toc}{subsubsection}{Connection/building method}
+
+To build the sensor, we followed these steps:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Set up the Raspberry Pi with the Rasbian software using the NOOBS
+  scandisk. (insert the scandisk into the Rasberry Pi) A comprehensive
+  start up guide can be found here.
+\item
+  Next we load the code onto the Raspberry Pi. The code can be found on
+  the Open Data Durban Hospital Stations Repository Github. To use it
+  you can follow the steps in the file called Initiate\_WS.txt. Once all
+  the steps have been completed, the Raspberry Pi is now a device that
+  can be connected to a particulate matter monitoring sensor.\\
+\item
+  Connect the hardware components
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  Connect the Raspberry Pi to the Nova PM Sensor using the break out
+  board that comes with the Sensor.
+\item
+  Connect the battery to the Raspberry Pi via the micro port
+\item
+  Attach the rubber pipe to the sensor
+\item
+  Assemble all components inside the box, with the pipe coming out of
+  the drilled hole
+\item
+  To prevent dirt entering the system, close off all openings between
+  the pipe and the box with silicone putty
+\item
+  Secure box in an elevated position (not too close to the ground) and
+  with the pipe facing the direction of the inflow of the air under
+  investigation.
+\end{itemize}
 

### 2023-04-05 manuscript/manuscript.qmd
Abstract, Highlights, Authors

@@
 ---
-title: Short Paper
-subtitle: A Short Subtitle
+title: The invisible title
 author:
-  - name: Alice Anonymous
-    email: alice@example.com
+  - name: Elizabeth Tilley
+    orcid: 0000-0002-2095-9724
+    email: tilleye@ethz.ch
     affiliations: 
-        - id: some-tech
-          name: Some Institute of Technology
-          department: Department Name
-          address: Street Address
-          city: City
-          state: State
-          postal-code: Postal Code
+        - id: ethz
+          name: ETH Zurich, Global Health Engineering
+          department: Department of Mechanical and Process Engineering
+          address: Clausiusstrasse 37
+          city: Zurich
+          postal-code: 8092
     attributes:
         corresponding: true
-    note: This is the first author footnote.
-  - name: Bob Security
-    email: bob@example.com
+  - name: Marc Kalina
+    orcid: 0000-0002-6335-3845
+    email: mkalina@ethz.ch
     affiliations:
-        - id: another-u
-          name: Another University
-          department: Department Name
-          address: Street Address
-          city: City
-          state: State
-          postal-code: Postal Code
-    note: |
-      Another author footnote, this is a very long footnote and it should be a really long footnote. But this footnote is not yet sufficiently long enough to make two lines of footnote text.
-  - name: Cat Memes
-    email: cat@example.com
-    affiliations:
-        - ref: another-u
-    note: Yet another author footnote.
-  - name: Derek Zoolander
-    email: derek@example.com
+        - ref: ethz
+  - name: Jonathan Kwangulero
+    orcid: 0000-0002-4050-5121
+    email: kwangerlo@gmail.comm
     affilations:
-        - ref: some-tech
+        - id: polytech
+          name: The Polytechnic, University of Malawi
+          department: Department of Environmental Health
+  - name: Hope Chilunga
+    orcid: 0000-0003-4216-6319
+    email: hchilunga@poly.ac.mw
+    affilations:
+        - ref: polytech
 abstract: |
-  This is the abstract.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit.
-Nam placerat turpis purus, eu tristique ex tincidunt et.
-Mauris sed augue eget turpis ultrices tincidunt.
-Sed et mi in leo porta egestas.
-Aliquam non laoreet velit.
-Nunc quis ex vitae eros aliquet auctor nec ac libero.
-Duis laoreet sapien eu mi luctus, in bibendum leo molestie.
-Sed hendrerit diam diam, ac dapibus nisl volutpat vitae.
-Aliquam bibendum varius libero, eu efficitur justo rutrum at.
-Sed at tempus elit.
+  Built in 1964, Malawi’s year of independence, Queen Elizabeth Hospital (QECH), or Queen’s, as it’s known locally, is the country’s largest public hospital.
+However, Queen’s is not served by regular municipal waste collection.
+What municipal collection that is done is ad-hoc, sporadic, and based on the hospital’s available financial resources.
+Rather, most hospital waste (infectious and non-infectious) is gathered by grounds staff and openly burned, in several constantly smouldering piles, sending up clouds of smoke.
+This study utilised a mixed-methods approach to both quantitatively measure the air quality around QECH, and to qualitatively investigate the perceived impacts amongst staff and caregivers.
+Low-cost sensors measuring particulate matter (PM) with particle sizes less than 10μm (PM10) and less than 2.5μm (PM2.5), expressed as the mass of PM per volume of air (μg PMx/m3 air) were recorded every 5 minutes at 8 locations across the QECH for 2 months.
+Qualitative data collection consisted of 57 interviews with patients, caregivers and hospital staff (including janitorial and maintenance staff, nurses, doctors, and administrators).
+Our results show that a) safe air quality thresholds are consistently exceeded across space and time; and b) that the most problematic air quality surrounds the shelter for caregivers and those receiving treatment for HIV/AIDs.
+Nearly every staff member interviewed recalled having received a complaint about the smoke, but those in charge of waste management felt powerless to make changes or deal with complaints.
+Waste management interventions are desperately needed lest the patients who arrive at Queen’s leave with more health issues than the ones with which they arrived.
 keywords: 
-  - keyword1
-  - keyword2
+  - waste management
+  - healthcare waste
+  - trash burning
+  - air quality
+  - Malawi
+  - urbanisation
 date: last-modified
@@
       highlights:
-        - Highlight 1
-        - Highlight 2
-        - Highlight 3
+        - Low-cost monitors were used to assess air quality at Malawi’s largest hospital
+        - Open waste burning is physically extensive and temporally frequent across the hospital campus
+        - Air quality measured as PM2.5 and PM10 consistently exceeded even the most hazardous levels 
+        - The most exposed locations were where guardians of patients stay and the HIV clinic
 execute: 

### 2023-04-05 manuscript/manuscript.tex
Abstract, Highlights, Authors

@@
 \hypersetup{
-  pdftitle={Short Paper},
-  pdfauthor={Alice Anonymous; Bob Security; Cat Memes; Derek Zoolander},
-  pdfkeywords={keyword1, keyword2},
+  pdftitle={The invisible title},
+  pdfauthor={Elizabeth Tilley; Marc Kalina; Jonathan Kwangulero; Hope Chilunga},
+  pdfkeywords={waste management, healthcare waste, trash burning, air
+quality, Malawi, urbanisation},
   colorlinks=true,
@@
 \begin{frontmatter}
-\title{Short Paper \\\large{A Short Subtitle} }
-\author[1]{Alice Anonymous%
+\title{The invisible title}
+\author[1]{Elizabeth Tilley%
 \corref{cor1}%
-\fnref{fn1}}
- \ead{alice@example.com} 
-\author[2]{Bob Security%
-%
-\fnref{fn2}}
- \ead{bob@example.com} 
-\author[2]{Cat Memes%
-%
-\fnref{fn3}}
- \ead{cat@example.com} 
-\author[]{Derek Zoolander%
+}
+ \ead{tilleye@ethz.ch} 
+\author[1]{Marc Kalina%
 %
 }
- \ead{derek@example.com}
-
-\affiliation[1]{organization={Some Institute of Technology, Department
-Name},addressline={Street Address},city={City},postcode={Postal
-Code},postcodesep={}}
-\affiliation[2]{organization={Another University, Department
-Name},addressline={Street Address},city={City},postcode={Postal
-Code},postcodesep={}}
+ \ead{mkalina@ethz.ch} 
+\author[]{Jonathan Kwangulero%
+%
+}
+ \ead{kwangerlo@gmail.comm} 
+\author[]{Hope Chilunga%
+%
+}
+ \ead{hchilunga@poly.ac.mw}
+
+\affiliation[1]{organization={ETH Zurich, Global Health
+Engineering, Department of Mechanical and Process
+Engineering},addressline={Clausiusstrasse
+37},city={Zurich},postcode={8092},postcodesep={}}
 
 \cortext[cor1]{Corresponding author}
-\fntext[fn1]{This is the first author footnote.}
-\fntext[fn2]{Another author footnote, this is a very long footnote and
-it should be a really long footnote. But this footnote is not yet
-sufficiently long enough to make two lines of footnote text.}
-\fntext[fn3]{Yet another author footnote.}
 
 \begin{abstract}
-This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing
-elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget
-velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris
-sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta
-egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet
-auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo
-molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae.
-Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at
-tempus elit.
+Built in 1964, Malawi's year of independence, Queen Elizabeth Hospital
+(QECH), or Queen's, as it's known locally, is the country's largest
+public hospital. However, Queen's is not served by regular municipal
+waste collection. What municipal collection that is done is ad-hoc,
+sporadic, and based on the hospital's available financial resources.
+Rather, most hospital waste (infectious and non-infectious) is gathered
+by grounds staff and openly burned, in several constantly smouldering
+piles, sending up clouds of smoke. This study utilised a mixed-methods
+approach to both quantitatively measure the air quality around QECH, and
+to qualitatively investigate the perceived impacts amongst staff and
+caregivers. Low-cost sensors measuring particulate matter (PM) with
+particle sizes less than 10μm (PM10) and less than 2.5μm (PM2.5),
+expressed as the mass of PM per volume of air (μg PMx/m3 air) were
+recorded every 5 minutes at 8 locations across the QECH for 2 months.
+Qualitative data collection consisted of 57 interviews with patients,
+caregivers and hospital staff (including janitorial and maintenance
+staff, nurses, doctors, and administrators). Our results show that a)
+safe air quality thresholds are consistently exceeded across space and
+time; and b) that the most problematic air quality surrounds the shelter
+for caregivers and those receiving treatment for HIV/AIDs. Nearly every
+staff member interviewed recalled having received a complaint about the
+smoke, but those in charge of waste management felt powerless to make
+changes or deal with complaints. Waste management interventions are
+desperately needed lest the patients who arrive at Queen's leave with
+more health issues than the ones with which they arrived.
 \end{abstract}
@@
 \begin{highlights}
-\item Highlight 1\item Highlight 2\item Highlight 3
+\item Low-cost monitors were used to assess air quality at Malawi's
+largest hospital\item Open waste burning is physically extensive and
+temporally frequent across the hospital campus\item Air quality measured
+as PM2.5 and PM10 consistently exceeded even the most hazardous
+levels\item The most exposed locations were where guardians of patients
+stay and the HIV clinic
 \end{highlights}
@@
 \begin{keyword}
-    keyword1 \sep 
-    keyword2
+    waste management \sep healthcare waste \sep trash burning \sep air
+quality \sep Malawi \sep 
+    urbanisation
 \end{keyword}
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, breakable]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable]}{\end{tcolorbox}}\fi
 

### 2023-04-06 manuscript/manuscript.qmd
Add authors, conclusions, acknowledgements

@@
 ---
-title: The invisible title
+title: "'It Is Unbearable To Breath Here': Air Quality, Open Incineration, And Misinformation In Blantyre, Malawi"
 author:
@@
         - id: ethz
-          name: ETH Zurich, Global Health Engineering
+          name: ETH Zurich
           department: Department of Mechanical and Process Engineering
-          address: Clausiusstrasse 37
           city: Zurich
-          postal-code: 8092
+          country: Switzerland
     attributes:
         corresponding: true
+  - name: Hope Chilunga
+    orcid: 0000-0003-4216-6319
+    email: hchilunga@poly.ac.mw
+    affiliations: 
+        - id: mubas
+          name: Malawi University of Business and Applied Sciences
+          department: Department of Environmental Health
+          city: Blantyre
+          country: Malawi
+  - name: Jonathan Kwangulero
+    orcid: 0000-0002-4050-5121
+    email: kwangerlo@gmail.comm
+    affilations:
+        - ref: mubas
+  - name: Lars Schöbitz
+    orcid: 0000-0003-2196-5015
+    email: lschoebitz@ethz.ch
+    affiliations:
+        - ref: ethz
+  - name: Heiko Heilgendorff
+    email: me@me.com
+    affiliations: 
+        - id: ukzn
+          name: University of KwaZulu-Natal
+          department: School of Mathematics, Statistics and Computer Science
+          city: Durban
+          country: South Africa
   - name: Marc Kalina
@@
         - ref: ethz
-  - name: Jonathan Kwangulero
-    orcid: 0000-0002-4050-5121
-    email: kwangerlo@gmail.comm
-    affilations:
-        - id: polytech
-          name: The Polytechnic, University of Malawi
-          department: Department of Environmental Health
-  - name: Hope Chilunga
-    orcid: 0000-0003-4216-6319
-    email: hchilunga@poly.ac.mw
-    affilations:
-        - ref: polytech
 abstract: |
-  Built in 1964, Malawi’s year of independence, Queen Elizabeth Hospital (QECH), or Queen’s, as it’s known locally, is the country’s largest public hospital.
+  Blantyre, Malawi’s Queen Elizabeth Central Hospital (QECH), or Queen’s, as it’s known locally, is the country’s largest public hospital.
 However, Queen’s is not served by regular municipal waste collection.
@@
 Rather, most hospital waste (infectious and non-infectious) is gathered by grounds staff and openly burned, in several constantly smouldering piles, sending up clouds of smoke.
-This study utilised a mixed-methods approach to both quantitatively measure the air quality around QECH, and to qualitatively investigate the perceived impacts amongst staff and caregivers.
+Speaking directly to an identified knowledge gap on air quality impacts linked to trash burning and the paucity of African urban dwellers’ voices on air quality issues, this study employed a mixed-methods approach to both quantitatively measure the air quality around QECH, and to qualitatively investigate the perceived impacts amongst staff and caregivers.
 Low-cost sensors measuring particulate matter (PM) with particle sizes less than 10μm (PM10) and less than 2.5μm (PM2.5), expressed as the mass of PM per volume of air (μg PMx/m3 air) were recorded every 5 minutes at 8 locations across the QECH for 2 months.
 Qualitative data collection consisted of 57 interviews with patients, caregivers and hospital staff (including janitorial and maintenance staff, nurses, doctors, and administrators).
-Our results show that a) safe air quality thresholds are consistently exceeded across space and time; and b) that the most problematic air quality surrounds the shelter for caregivers and those receiving treatment for HIV/AIDs.
-Nearly every staff member interviewed recalled having received a complaint about the smoke, but those in charge of waste management felt powerless to make changes or deal with complaints.
+Our results show that safe air quality thresholds are consistently exceeded across space and time and that the most problematic air quality surrounds the shelter for caregivers and those receiving treatment for HIV/AIDs.
+Moreover, staff and visitors are severely impacted by the poor air quality within the space, but  feel powerless to make changes or address complaints.
 Waste management interventions are desperately needed lest the patients who arrive at Queen’s leave with more health issues than the ones with which they arrived.
@@
 
-Built in 1964, Malawi's year of independence, Queen Elizabeth Hospital
+Built in 1964, Malawi's year of independence, Queen Elizabeth Central Hospital
 (QECH), or Queen's, as it's known locally, is the country's largest
@@
 
+The air quality, as a result of open waste burning at Queen’s is poor and not suitable for a city, let alone a hospital.
+Over the course of this two-month study, the hazardous limits for both PM parameters were exceeded at all locations except for the Lions Sight hospital.
+The WHO limits were exceeded fewer than 50 times at five of the locations, and only the monitors at the Lighthouse Clinic and the Guardian Shelter recorded more than 50 instances above the most hazardous limit.
+The composition of emissions recorded ranged significantly, analysed as the ratio PM10:PM2.5 likely reflects a range of burning styles and/or fuel (waste) type, and no discernable trend was observed..
+Among interview respondents, there was a general  consensus that air quality caused by the burning of waste within the hospital was a problem, although there were significant differences between respondents over their understandings of potential impacts, the effectiveness of various coping mechanisms, and their problematisations linked to the burning of specific waste materials.
+For most, going home or leaving the Queen’s campus was the only pathway to relief, though for patients who can not leave, endurance was the only option.
+This work was originally planned as a before-and-after study:  we envisioned reporting these baseline results compared to follow-up measurements once the long-promised incinerator was commissioned.
+However, nearly 4 years later, the incinerator is rarely functional (partly for technical reasons, partly for financial ones) and the smoke hovering over Queen’s is still thick.
+Waste management has long been known as an environmental health issue, but in the case of open trash burning at a medical facility, it becomes more:  for some of the most vulnerable, especially those with TB, asthma or other chronic disease, it could be a matter of life or death.
+
+# Acknowledgements
+
+The authors are grateful to the staff, patients and caregivers at QECH for their assistance, willingness to share their views, and importantly, perseverance in delivering the best care possible under difficult circumstances.
+Moreover, we would like to acknowledge the assistance and effort of Open Data Durban, now the Open Cities Lab, for their assistance in constructing and programming the sensors.
+
 # Appendix {.unnumbered}

### 2023-04-06 manuscript/manuscript.tex
Add authors, conclusions, acknowledgements

@@
 \hypersetup{
-  pdftitle={The invisible title},
-  pdfauthor={Elizabeth Tilley; Marc Kalina; Jonathan Kwangulero; Hope Chilunga},
+  pdftitle={`It Is Unbearable To Breath Here': Air Quality, Open Incineration, And Misinformation In Blantyre, Malawi},
+  pdfauthor={Elizabeth Tilley; Hope Chilunga; Jonathan Kwangulero; Lars Schöbitz; Heiko Heilgendorff; Marc Kalina},
   pdfkeywords={waste management, healthcare waste, trash burning, air
@@
 \begin{frontmatter}
-\title{The invisible title}
+\title{`It Is Unbearable To Breath Here': Air Quality, Open
+Incineration, And Misinformation In Blantyre, Malawi}
 \author[1]{Elizabeth Tilley%
@@
  \ead{tilleye@ethz.ch} 
+\author[2]{Hope Chilunga%
+%
+}
+ \ead{hchilunga@poly.ac.mw} 
+\author[]{Jonathan Kwangulero%
+%
+}
+ \ead{kwangerlo@gmail.comm} 
+\author[1]{Lars Schöbitz%
+%
+}
+ \ead{lschoebitz@ethz.ch} 
+\author[3]{Heiko Heilgendorff%
+%
+}
+ \ead{me@me.com} 
 \author[1]{Marc Kalina%
@@
  \ead{mkalina@ethz.ch}
-\author[]{Jonathan Kwangulero%
-%
-}
- \ead{kwangerlo@gmail.comm} 
-\author[]{Hope Chilunga%
-%
-}
- \ead{hchilunga@poly.ac.mw}
-
-\affiliation[1]{organization={ETH Zurich, Global Health
-Engineering, Department of Mechanical and Process
-Engineering},addressline={Clausiusstrasse
-37},city={Zurich},postcode={8092},postcodesep={}}
+
+\affiliation[1]{organization={ETH Zurich, Department of Mechanical and
+Process
+Engineering},city={Zurich},country={Switzerland},countrysep={,},postcodesep={}}
+\affiliation[2]{organization={Malawi University of Business and Applied
+Sciences, Department of Environmental
+Health},city={Blantyre},country={Malawi},countrysep={,},postcodesep={}}
+\affiliation[3]{organization={University of KwaZulu-Natal, School of
+Mathematics, Statistics and Computer
+Science},city={Durban},country={South
+Africa},countrysep={,},postcodesep={}}
 
@@
 \begin{abstract}
-Built in 1964, Malawi's year of independence, Queen Elizabeth Hospital
-(QECH), or Queen's, as it's known locally, is the country's largest
-public hospital. However, Queen's is not served by regular municipal
-waste collection. What municipal collection that is done is ad-hoc,
-sporadic, and based on the hospital's available financial resources.
-Rather, most hospital waste (infectious and non-infectious) is gathered
-by grounds staff and openly burned, in several constantly smouldering
-piles, sending up clouds of smoke. This study utilised a mixed-methods
-approach to both quantitatively measure the air quality around QECH, and
-to qualitatively investigate the perceived impacts amongst staff and
-caregivers. Low-cost sensors measuring particulate matter (PM) with
-particle sizes less than 10μm (PM10) and less than 2.5μm (PM2.5),
-expressed as the mass of PM per volume of air (μg PMx/m3 air) were
-recorded every 5 minutes at 8 locations across the QECH for 2 months.
-Qualitative data collection consisted of 57 interviews with patients,
-caregivers and hospital staff (including janitorial and maintenance
-staff, nurses, doctors, and administrators). Our results show that a)
-safe air quality thresholds are consistently exceeded across space and
-time; and b) that the most problematic air quality surrounds the shelter
-for caregivers and those receiving treatment for HIV/AIDs. Nearly every
-staff member interviewed recalled having received a complaint about the
-smoke, but those in charge of waste management felt powerless to make
-changes or deal with complaints. Waste management interventions are
-desperately needed lest the patients who arrive at Queen's leave with
-more health issues than the ones with which they arrived.
+Blantyre, Malawi's Queen Elizabeth Central Hospital (QECH), or Queen's,
+as it's known locally, is the country's largest public hospital.
+However, Queen's is not served by regular municipal waste collection.
+What municipal collection that is done is ad-hoc, sporadic, and based on
+the hospital's available financial resources. Rather, most hospital
+waste (infectious and non-infectious) is gathered by grounds staff and
+openly burned, in several constantly smouldering piles, sending up
+clouds of smoke. Speaking directly to an identified knowledge gap on air
+quality impacts linked to trash burning and the paucity of African urban
+dwellers' voices on air quality issues, this study employed a
+mixed-methods approach to both quantitatively measure the air quality
+around QECH, and to qualitatively investigate the perceived impacts
+amongst staff and caregivers. Low-cost sensors measuring particulate
+matter (PM) with particle sizes less than 10μm (PM10) and less than
+2.5μm (PM2.5), expressed as the mass of PM per volume of air (μg PMx/m3
+air) were recorded every 5 minutes at 8 locations across the QECH for 2
+months. Qualitative data collection consisted of 57 interviews with
+patients, caregivers and hospital staff (including janitorial and
+maintenance staff, nurses, doctors, and administrators). Our results
+show that safe air quality thresholds are consistently exceeded across
+space and time and that the most problematic air quality surrounds the
+shelter for caregivers and those receiving treatment for HIV/AIDs.
+Moreover, staff and visitors are severely impacted by the poor air
+quality within the space, but feel powerless to make changes or address
+complaints. Waste management interventions are desperately needed lest
+the patients who arrive at Queen's leave with more health issues than
+the ones with which they arrived.
 \end{abstract}
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 
+The air quality, as a result of open waste burning at Queen's is poor
+and not suitable for a city, let alone a hospital. Over the course of
+this two-month study, the hazardous limits for both PM parameters were
+exceeded at all locations except for the Lions Sight hospital. The WHO
+limits were exceeded fewer than 50 times at five of the locations, and
+only the monitors at the Lighthouse Clinic and the Guardian Shelter
+recorded more than 50 instances above the most hazardous limit. The
+composition of emissions recorded ranged significantly, analysed as the
+ratio PM10:PM2.5 likely reflects a range of burning styles and/or fuel
+(waste) type, and no discernable trend was observed.. Among interview
+respondents, there was a general consensus that air quality caused by
+the burning of waste within the hospital was a problem, although there
+were significant differences between respondents over their
+understandings of potential impacts, the effectiveness of various coping
+mechanisms, and their problematisations linked to the burning of
+specific waste materials. For most, going home or leaving the Queen's
+campus was the only pathway to relief, though for patients who can not
+leave, endurance was the only option. This work was originally planned
+as a before-and-after study: we envisioned reporting these baseline
+results compared to follow-up measurements once the long-promised
+incinerator was commissioned. However, nearly 4 years later, the
+incinerator is rarely functional (partly for technical reasons, partly
+for financial ones) and the smoke hovering over Queen's is still thick.
+Waste management has long been known as an environmental health issue,
+but in the case of open trash burning at a medical facility, it becomes
+more: for some of the most vulnerable, especially those with TB, asthma
+or other chronic disease, it could be a matter of life or death.
+
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+The authors are grateful to the staff, patients and caregivers at QECH
+for their assistance, willingness to share their views, and importantly,
+perseverance in delivering the best care possible under difficult
+circumstances. Moreover, we would like to acknowledge the assistance and
+effort of Open Data Durban, now the Open Cities Lab, for their
+assistance in constructing and programming the sensors.
+
 \hypertarget{appendix}{%

### 2023-04-06 manuscript/manuscript.qmd
Fix units

@@
           department: Department of Mechanical and Process Engineering
+          address: Global Health Engineering, Clausiusstrasse 37
           city: Zurich
           country: Switzerland
+          postal-code: 8092
     attributes:
@@
 Speaking directly to an identified knowledge gap on air quality impacts linked to trash burning and the paucity of African urban dwellers’ voices on air quality issues, this study employed a mixed-methods approach to both quantitatively measure the air quality around QECH, and to qualitatively investigate the perceived impacts amongst staff and caregivers.
-Low-cost sensors measuring particulate matter (PM) with particle sizes less than 10μm (PM10) and less than 2.5μm (PM2.5), expressed as the mass of PM per volume of air (μg PMx/m3 air) were recorded every 5 minutes at 8 locations across the QECH for 2 months.
+Low-cost sensors measuring particulate matter (PM) with particle sizes less than 10 µm (PM~10~) and less than 2.5 µm (PM~2.5~), expressed as the mass of PM per volume of air (µg PMx/m^3^ air) were recorded every 5 minutes at 8 locations across the QECH for 2 months.
 Qualitative data collection consisted of 57 interviews with patients, caregivers and hospital staff (including janitorial and maintenance staff, nurses, doctors, and administrators).
@@
         - Open waste burning is physically extensive and temporally frequent across the hospital campus
-        - Air quality measured as PM2.5 and PM10 consistently exceeded even the most hazardous levels 
+        - Air quality measured as PM~2.5~ and PM~10~ consistently exceeded even the most hazardous levels 
         - The most exposed locations were where guardians of patients stay and the HIV clinic
@@
 broadly, for instance, @wiedinmyer2014global estimate that open trash
-burning contributes to 29% of the total global PM2.5 anthropogenic
+burning contributes to 29% of the total global PM~2.5~ anthropogenic
 emissions, it is of particular concern in Sub-Saharan African (SSA)
@@
 indicators for air quality. PM is measured in terms of its particle size
-and generally classified into PM with particle sizes less than 10μm
-(PM10) and those with sizes less than 2.5μm (PM2.5), while one way of
+and generally classified into PM with particle sizes less than 10 µm
+(PM~10~) and those with sizes less than 2.5 µm (PM~2.5~), while one way of
 measuring the severity of air pollution is by expressing the mass of PM
-per volume of air (μg PMx/m3 air). PM has been linked to multiple
-negative health outcomes. Amongst particulate air pollution, PM10 and
-PM2.5 are of particular concern as they can penetrate deep into the
+per volume of air (µg PMx/m^3^ air). PM has been linked to multiple
+negative health outcomes. Amongst particulate air pollution, PM~10~ and
+PM~2.5~ are of particular concern as they can penetrate deep into the
 lungs, and their exposure has been associated with asthma, heart
@@
 locations. Over the course of the two-month study, the hazardous limits
-for both PM10 and PM2.5 were exceeded at all locations except for the
+for both PM~10~ and PM~2.5~ were exceeded at all locations except for the
 Lions Sight hospital. The limits were exceeded fewer than 50 times at
@@
 Precision Laser sensor. This sensor measures particles at 2.5 and 10
-micrometres in diameter in mg/m3. A list of the specific hardware and
+µm in diameter in mg/m^3^. A list of the specific hardware and
 software components can be found in Appendix B.
@@
 [@kutlarjoss2017time; @mapoma2014air] indicate that the maximum 24-h
-PM10 value is 25 μg/m3 and the maximum annual PM2.5 value is 8 μg/m3.
-There is no daily maximum value for PM2.5. It should be noted that both
+PM~10~ value is 25 µg/m^3^ and the maximum annual PM~2.5~ value is 8 µg/m^3^.
+There is no daily maximum value for PM~2.5~. It should be noted that both
 of these values are at, or below, the level-4 interim WHO targets.
@@
 persistent trends. The full set of 5-minute data are plotted and
-presented in Appendix A Figure A.1. However, the number of measured values
+presented in Appendix A @fig-five-minute-intervals. However, the number of measured values
 that exceeded the hazardous limit for both parameters are summarised in
@@
   mutate(peaks = case_when(
-    indicator == "pm2.5" & value >  35 ~ TRUE,
+    indicator == "pm25" & value >  35 ~ TRUE,
     indicator == "pm10" & value >  70 ~ TRUE,
@@
 Guardian Shelter recorded more than 50 instances above the hazardous
-limits. At the Lighthouse Clinic, the PM10 limit was exceeded almost
-twice as often as the PM2.5 limit, while at the Guardian Shelter, the
+limits. At the Lighthouse Clinic, the PM~10~ limit was exceeded almost
+twice as often as the PM~2.5~ limit, while at the Guardian Shelter, the
 reverse is true. Because the Lighthouse Clinic is within 50 m of the
-incinerator, the number and predominance of PM10 peaks is characteristic
+incinerator, the number and predominance of PM~10~ peaks is characteristic
 of incomplete combustion and dust that is typical for the area and the
@@
 designed to improve combustion of the coal and wood that is burned
-within them. The fewer number of PM10 peaks is a testament to this
-intervention, though the frequency of PM2.5 peaks is still beyond
+within them. The fewer number of PM~10~ peaks is a testament to this
+intervention, though the frequency of PM~2.5~ peaks is still beyond
 acceptable.
@@
 The 24-hour averaged values (logged every 5 minutes) are presented for
-each location, for both PM10 and PM2.5 in fig-24-hour-average. As well as dampening
+each location, for both PM~10~ and PM~2.5~ in fig-24-hour-average. As well as dampening
 local variation and peaks, the WHO air quality guidelines are also based
@@
 
-Overall, PM2.5 values remained below 100 μg/m3 at 6 of 8 locations
+Overall, PM~2.5~ values remained below 100 µg/m^3^ at 6 of 8 locations
 (Administration, Lions Sight, Malaria Project, uMoyo, Ward 6B, and Mercy
-James); PM10 values were consistently below 100μg/m3 at the same
+James); PM~10~ values were consistently below 100µg/m^3^ at the same
 locations, but with several average values extending slightly above, and
@@
 #| label: fig-24-hour-average
-#| fig-cap: Average 24-hour PM10 and PM2.5 at 8 monitoring stations over 2 months
+#| fig-cap: Average 24-hour PM~10~ and PM~2.5~ at 8 monitoring stations over 2 months
 #| fig-align: center
@@
 Given that the type of fuel and the type of burning (contained vs. open)
-produce very different particulate "fingerprints", the ratio of PM10 to
-PM2.5 values were examined to determine if a clear difference between
+produce very different particulate "fingerprints", the ratio of PM~10~ to
+PM~2.5~ values were examined to determine if a clear difference between
 locations, and therefore source, could be identified. The results are
@@
 #| label: fig-pm-ratios
-#| fig-cap: PM10:PM2.5 values for each location.
+#| fig-cap: PM~10~:PM~2.5~ values for each location.
 Each line shows the minimum and maximum ratio for the day based on ratios calculated hourly.
@@
               values_from = mean) |> 
-  mutate(p10_pm2.5_ratio = pm10/pm2.5)  |>
+  mutate(p10_pm25_ratio = pm10/pm25)  |>
 
@@
          | location != "uMoyo") |> 
-  #filter(p10_pm2.5_ratio > 30)
+  #filter(p10_PM~2.5~_ratio > 30)
 
   group_by(date, location) |> 
-  summarise(min = min(p10_pm2.5_ratio),
-            max = max(p10_pm2.5_ratio))
+  summarise(min = min(p10_pm25_ratio),
+            max = max(p10_pm25_ratio))
 
@@
 
-The results presented in Figure 5 illustrate both the presence of PM10
-relative to PM2.5 as well as the range of values over which that ratio
+The results presented in Figure 5 illustrate both the presence of PM~10~
+relative to PM~2.5~ as well as the range of values over which that ratio
 is observed.
@@
 Unlike the peak values presented in Table 2, the values presented in
-Figure 5 indicate the relative presence of larger (PM10) particles to
-smaller ones (PM2.5) regardless of their concentration: the ratio of two
+Figure 5 indicate the relative presence of larger (PM~10~) particles to
+smaller ones (PM~2.5~) regardless of their concentration: the ratio of two
 low concentrations can be the same as the ratio of two high
@@
 dat_air_quality_sum_who_2021 |> 
-  #filter(indicator == "pm2.5") |> 
+  #filter(indicator == "PM~2.5~") |> 
   group_by(location, indicator) |> 
@@
     location == "Administration" 
-    & indicator == "pm2.5"
+    & indicator == "pm25"
     & air_quality_who_annual == "Interim target 3" ~ exposure_percent + 1,
@@
     location == "uMoyo" 
-    & indicator == "pm2.5"
+    & indicator == "pm25"
     & air_quality_who_annual == "Interim target 4" ~ exposure_percent - 1,
@@
 The WHO limits were exceeded fewer than 50 times at five of the locations, and only the monitors at the Lighthouse Clinic and the Guardian Shelter recorded more than 50 instances above the most hazardous limit.
-The composition of emissions recorded ranged significantly, analysed as the ratio PM10:PM2.5 likely reflects a range of burning styles and/or fuel (waste) type, and no discernable trend was observed..
+The composition of emissions recorded ranged significantly, analysed as the ratio PM~10~:PM~2.5~ likely reflects a range of burning styles and/or fuel (waste) type, and no discernable trend was observed..
 Among interview respondents, there was a general  consensus that air quality caused by the burning of waste within the hospital was a problem, although there were significant differences between respondents over their understandings of potential impacts, the effectiveness of various coping mechanisms, and their problematisations linked to the burning of specific waste materials.
@@
 ```{r}
-#| label: fig-5-minute-intervals
-#| fig-cap: PM2.5 and PM10 values collected every 5 minutes over 10 months at 8 locations.
+#| label: fig-five-minute-intervals
+#| fig-cap: PM~2.5~ and PM~10~ values collected every 5 minutes over 10 months at 8 locations.
 #| fig-align: center
@@
   labs(title = "Hospital Air Quality",
-       subtitle = "PM2.5 and PM10 measurements in 5-min intervals at eight locations") +
+       subtitle = "PM~2.5~ and PM~10~ measurements in 5-min intervals at eight locations") +
   ylab(bquote(uq/m^3)) +

### 2023-04-06 manuscript/manuscript.qmd
Spell-check completed

@@
     email: lschoebitz@ethz.ch
+    affiliations:
+        - ref: ethz
+  - name: Saloni Vijay
+    orcid: 0000-0001-5883-5308
+    email: svijay@ethz.ch
     affiliations:
@@
 financial resources. For instance, although hospital administration was
-able to articulate a set of standard operating principles (SoPs) for
+able to articulate a set of standard operating principles (SOPs) for
 waste management, below the administration, doctors and nurses were
@@
 lost if the internet connection breaks down. Because the Pi can only
-read digital signals, we needed to include an analog-to-digital (ADC)
+read digital signals, we needed to include an analogue-to-digital (ADC)
 converter. The amount of dust is measured by a Nova PM SDS011 High
@@
 Because connecting each unit to a power supply was not possible, each
-unit was equipped with an external battery (5000mAh) which powered the
+unit was equipped with an external battery (5000 mAh) which powered the
 sensor unit for three days continuously. To avoid any down time, we
@@
 sensor at the time of battery replacement. The sensors were capable of
-being connected to WIFI, a feature which enabled us to download the data
-into our computers wirelessly.
+being connected to Wi-Fi, a feature which enabled wireless data access.
 
@@
 The WHO limits were exceeded fewer than 50 times at five of the locations, and only the monitors at the Lighthouse Clinic and the Guardian Shelter recorded more than 50 instances above the most hazardous limit.
-The composition of emissions recorded ranged significantly, analysed as the ratio PM~10~:PM~2.5~ likely reflects a range of burning styles and/or fuel (waste) type, and no discernable trend was observed..
+The composition of emissions recorded ranged significantly, analysed as the ratio PM~10~:PM~2.5~ likely reflects a range of burning styles and/or fuel (waste) type, and no discernible trend was observed..
 Among interview respondents, there was a general  consensus that air quality caused by the burning of waste within the hospital was a problem, although there were significant differences between respondents over their understandings of potential impacts, the effectiveness of various coping mechanisms, and their problematisations linked to the burning of specific waste materials.
@@
 3.  MCP3008 microchip
--   analog-to-digital converter (ADC)
+-   analogue-to-digital converter (ADC)
 4.  Romoss Sense8+ 30000mAh QC Type-C
@@
 6.  Box
--   any box that can fit all the components comfortbaly can be used.
+-   any box that can fit all the components comfortably can be used.
 The box will need to have a whole drilled in it for the pipe.
@@
 To be able to collect data, the device needs to be connected to a reliable wifi network, the SSID and password need to be coded onto the device during step 2 in the building method.
-3. Once the whole device has been put together and placed on site, to access that data we use a combination of propgrams:
+3. Once the whole device has been put together and placed on site, to access that data we use a combination of programs:
 - [Putty](https://www.putty.org/) - this is an SSH client that we use to gain access to the Raspberry Pi from another computer.
@@
 
-1. Set up the Raspberry Pi with the Rasbian software using the NOOBS scandisk. (insert the scandisk into the Rasberry Pi) A comprehensive start up guide can be found here. 
+1. Set up the Raspberry Pi with the Rasbian software using the NOOBS scandisk. (insert the scandisk into the Raspberry Pi) A comprehensive start up guide can be found here. 
 2.

### 2023-04-06 manuscript/manuscript.tex
Render

@@
 \affiliation[1]{organization={ETH Zurich, Department of Mechanical and
-Process
-Engineering},city={Zurich},country={Switzerland},countrysep={,},postcodesep={}}
+Process Engineering},addressline={Global Health Engineering,
+Clausiusstrasse
+37},city={Zurich},country={Switzerland},countrysep={,},postcode={8092},postcodesep={}}
 \affiliation[2]{organization={Malawi University of Business and Applied
@@
 amongst staff and caregivers. Low-cost sensors measuring particulate
-matter (PM) with particle sizes less than 10μm (PM10) and less than
-2.5μm (PM2.5), expressed as the mass of PM per volume of air (μg PMx/m3
-air) were recorded every 5 minutes at 8 locations across the QECH for 2
-months. Qualitative data collection consisted of 57 interviews with
-patients, caregivers and hospital staff (including janitorial and
-maintenance staff, nurses, doctors, and administrators). Our results
-show that safe air quality thresholds are consistently exceeded across
-space and time and that the most problematic air quality surrounds the
-shelter for caregivers and those receiving treatment for HIV/AIDs.
-Moreover, staff and visitors are severely impacted by the poor air
-quality within the space, but feel powerless to make changes or address
-complaints. Waste management interventions are desperately needed lest
-the patients who arrive at Queen's leave with more health issues than
-the ones with which they arrived.
+matter (PM) with particle sizes less than 10 µm (PM\textsubscript{10})
+and less than 2.5 µm (PM2.5), expressed as the mass of PM per volume of
+air (μg PMx/m3 air) were recorded every 5 minutes at 8 locations across
+the QECH for 2 months. Qualitative data collection consisted of 57
+interviews with patients, caregivers and hospital staff (including
+janitorial and maintenance staff, nurses, doctors, and administrators).
+Our results show that safe air quality thresholds are consistently
+exceeded across space and time and that the most problematic air quality
+surrounds the shelter for caregivers and those receiving treatment for
+HIV/AIDs. Moreover, staff and visitors are severely impacted by the poor
+air quality within the space, but feel powerless to make changes or
+address complaints. Waste management interventions are desperately
+needed lest the patients who arrive at Queen's leave with more health
+issues than the ones with which they arrived.
 \end{abstract}
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, frame hidden]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, enhanced, frame hidden, breakable]}{\end{tcolorbox}}\fi
 
@@
 
-Built in 1964, Malawi's year of independence, Queen Elizabeth Hospital
-(QECH), or Queen's, as it's known locally, is the country's largest
-public hospital. Designed in the Tropical Modernism style of
+Built in 1964, Malawi's year of independence, Queen Elizabeth Central
+Hospital (QECH), or Queen's, as it's known locally, is the country's
+largest public hospital. Designed in the Tropical Modernism style of
 architecture popular within Africa in the late colonial period, Queen's
@@
 quality. PM is measured in terms of its particle size and generally
-classified into PM with particle sizes less than 10μm (PM10) and those
-with sizes less than 2.5μm (PM2.5), while one way of measuring the
+classified into PM with particle sizes less than 10 µm (PM10) and those
+with sizes less than 2.5 µm (PM2.5), while one way of measuring the
 severity of air pollution is by expressing the mass of PM per volume of
@@
 persistent trends. The full set of 5-minute data are plotted and
-presented in Appendix A Figure A.1. However, the number of measured
-values that exceeded the hazardous limit for both parameters are
-summarised in Table~\ref{tbl-peaks}.
+presented in Appendix A Figure~\ref{fig-five-minute-intervals}. However,
+the number of measured values that exceeded the hazardous limit for both
+parameters are summarised in Table~\ref{tbl-peaks}.
 
@@
 
-{\centering \includegraphics{manuscript_files/figure-pdf/fig-5-minute-intervals-1.pdf}
-
-}
-
-\caption{\label{fig-5-minute-intervals}PM2.5 and PM10 values collected
-every 5 minutes over 10 months at 8 locations.}
+{\centering \includegraphics{manuscript_files/figure-pdf/fig-five-minute-intervals-1.pdf}
+
+}
+
+\caption{\label{fig-five-minute-intervals}PM2.5 and PM10 values
+collected every 5 minutes over 10 months at 8 locations.}
 

### 2023-04-06 manuscript/manuscript.qmd
Add reproducibility and data statement

@@
 date: last-modified
-bibliography: man-qechairquality.bib
+bibliography: [man-qechairquality.bib, packages.bib]
 format:
@@
 source(here::here("R/01-data-preparation.R"))
+
+knitr::write_bib(file = here::here("manuscript/packages.bib"))
 
@@
 
+### Computational reproducibility and data availability
+
+R Statistical Software version 4.2.1, RStudio IDE version 2023.3.0.386, and Quarto scientific publishing system version were used for quantitative data analysis and writing of the manuscript [@R-base; @allaire2022quartoa; @positteam2023rstudio].
+A set of additional R packages were used for data wrangling, analysis, and visualisation [@ggplot22016; @lubridate2011;  @R-dplyr; @R-forcats; @R-ggplot2; @R-janitor; @R-leaflet; @R-lubridate; @R-mapview; @R-readr; @R-tidyr; @R-waffle].
+
+Raw data and analysis-ready processed data is available as an R package at @R-qechairquality.
+The data underlying the tables and figures of this manuscript are contained in a repository alongside a reproducible document that contains the analysis code and the written narrative of the manuscript [TODO: manuscript software citation].
+
 # Theory/calculation (not applicable)
@@
               values_from = mean) |> 
-  mutate(p10_pm25_ratio = pm10/pm25)  |>
+  mutate(p10_pm25_ratio = pm10/pm2.5)  |>
 

### 2023-04-06 manuscript/manuscript.tex
Render

@@
   pdftitle={`It Is Unbearable To Breath Here': Air Quality, Open Incineration, And Misinformation In Blantyre, Malawi},
-  pdfauthor={Elizabeth Tilley; Hope Chilunga; Jonathan Kwangulero; Lars Schöbitz; Heiko Heilgendorff; Marc Kalina},
+  pdfauthor={Elizabeth Tilley; Hope Chilunga; Jonathan Kwangulero; Lars Schöbitz; Saloni Vijay; Heiko Heilgendorff; Marc Kalina},
   pdfkeywords={waste management, healthcare waste, trash burning, air
@@
  \ead{lschoebitz@ethz.ch} 
+\author[1]{Saloni Vijay%
+%
+}
+ \ead{svijay@ethz.ch} 
 \author[3]{Heiko Heilgendorff%
@@
 matter (PM) with particle sizes less than 10 µm (PM\textsubscript{10})
-and less than 2.5 µm (PM2.5), expressed as the mass of PM per volume of
-air (μg PMx/m3 air) were recorded every 5 minutes at 8 locations across
-the QECH for 2 months. Qualitative data collection consisted of 57
-interviews with patients, caregivers and hospital staff (including
-janitorial and maintenance staff, nurses, doctors, and administrators).
-Our results show that safe air quality thresholds are consistently
-exceeded across space and time and that the most problematic air quality
-surrounds the shelter for caregivers and those receiving treatment for
-HIV/AIDs. Moreover, staff and visitors are severely impacted by the poor
-air quality within the space, but feel powerless to make changes or
-address complaints. Waste management interventions are desperately
-needed lest the patients who arrive at Queen's leave with more health
-issues than the ones with which they arrived.
+and less than 2.5 µm (PM\textsubscript{2.5}), expressed as the mass of
+PM per volume of air (µg PMx/m\textsuperscript{3} air) were recorded
+every 5 minutes at 8 locations across the QECH for 2 months. Qualitative
+data collection consisted of 57 interviews with patients, caregivers and
+hospital staff (including janitorial and maintenance staff, nurses,
+doctors, and administrators). Our results show that safe air quality
+thresholds are consistently exceeded across space and time and that the
+most problematic air quality surrounds the shelter for caregivers and
+those receiving treatment for HIV/AIDs. Moreover, staff and visitors are
+severely impacted by the poor air quality within the space, but feel
+powerless to make changes or address complaints. Waste management
+interventions are desperately needed lest the patients who arrive at
+Queen's leave with more health issues than the ones with which they
+arrived.
 \end{abstract}
@@
 temporally frequent across the hospital campus\item Air quality measured
-as PM2.5 and PM10 consistently exceeded even the most hazardous
-levels\item The most exposed locations were where guardians of patients
-stay and the HIV clinic
+as PM\textsubscript{2.5} and PM\textsubscript{10} consistently exceeded
+even the most hazardous levels\item The most exposed locations were
+where guardians of patients stay and the HIV clinic
 \end{highlights}
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, enhanced, frame hidden, breakable]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, frame hidden, boxrule=0pt, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 \citet{wiedinmyer2014global} estimate that open trash burning
-contributes to 29\% of the total global PM2.5 anthropogenic emissions,
-it is of particular concern in Sub-Saharan African (SSA) cities, which
-are home to 19 of the world's 50 biggest dumpsites (United Nations
-Development Programme, 2021). Particulate matter, or PM, is not specific
-to trash burning, but rather the product of any type of incomplete
-combustion and is one of the most commonly measured indicators for air
-quality. PM is measured in terms of its particle size and generally
-classified into PM with particle sizes less than 10 µm (PM10) and those
-with sizes less than 2.5 µm (PM2.5), while one way of measuring the
-severity of air pollution is by expressing the mass of PM per volume of
-air (μg PMx/m3 air). PM has been linked to multiple negative health
-outcomes. Amongst particulate air pollution, PM10 and PM2.5 are of
-particular concern as they can penetrate deep into the lungs, and their
-exposure has been associated with asthma, heart disease, heart failure,
-stroke, and cancer \citep{mohapatra2014effect, xing2016impact}. As a
-result, the health impacts of ambient air particle pollution may be
-significant, especially for the most vulnerable. For instance, the
-\citet{ihme2019ambient} estimates PM related air pollution resulted in
-4.14 million deaths worldwide (in 2019), of which, 394,000 occurred in
-Africa. Yet, despite its importance, the air quality and health picture
-in Africa is worryingly incomplete. For instance, the WHO only collects
-data from 10 African countries (covering 39 cities)
+contributes to 29\% of the total global PM\textsubscript{2.5}
+anthropogenic emissions, it is of particular concern in Sub-Saharan
+African (SSA) cities, which are home to 19 of the world's 50 biggest
+dumpsites (United Nations Development Programme, 2021). Particulate
+matter, or PM, is not specific to trash burning, but rather the product
+of any type of incomplete combustion and is one of the most commonly
+measured indicators for air quality. PM is measured in terms of its
+particle size and generally classified into PM with particle sizes less
+than 10 µm (PM\textsubscript{10}) and those with sizes less than 2.5 µm
+(PM\textsubscript{2.5}), while one way of measuring the severity of air
+pollution is by expressing the mass of PM per volume of air (µg
+PMx/m\textsuperscript{3} air). PM has been linked to multiple negative
+health outcomes. Amongst particulate air pollution, PM\textsubscript{10}
+and PM\textsubscript{2.5} are of particular concern as they can
+penetrate deep into the lungs, and their exposure has been associated
+with asthma, heart disease, heart failure, stroke, and cancer
+\citep{mohapatra2014effect, xing2016impact}. As a result, the health
+impacts of ambient air particle pollution may be significant, especially
+for the most vulnerable. For instance, the \citet{ihme2019ambient}
+estimates PM related air pollution resulted in 4.14 million deaths
+worldwide (in 2019), of which, 394,000 occurred in Africa. Yet, despite
+its importance, the air quality and health picture in Africa is
+worryingly incomplete. For instance, the WHO only collects data from 10
+African countries (covering 39 cities)
 \citep{worldhealthorganization2016ambient}. Likewise, robust,
@@
 locations. Over the course of the two-month study, the hazardous limits
-for both PM10 and PM2.5 were exceeded at all locations except for the
-Lions Sight hospital. The limits were exceeded fewer than 50 times at
-five of the locations, but the monitors at the Lighthouse Clinic and the
-Guardian Shelter recorded more than 50 instances above the hazardous
-limits. The extremely hazardous air quality at the Guardian shelter is
-mostly from cooking over open fires, while the PM at the Lighthouse
-Clinic is a function of being located directly adjacent to the
-smouldering waste pile.
+for both PM\textsubscript{10} and PM\textsubscript{2.5} were exceeded at
+all locations except for the Lions Sight hospital. The limits were
+exceeded fewer than 50 times at five of the locations, but the monitors
+at the Lighthouse Clinic and the Guardian Shelter recorded more than 50
+instances above the hazardous limits. The extremely hazardous air
+quality at the Guardian shelter is mostly from cooking over open fires,
+while the PM at the Lighthouse Clinic is a function of being located
+directly adjacent to the smouldering waste pile.
 
@@
 lost if the internet connection breaks down. Because the Pi can only
-read digital signals, we needed to include an analog-to-digital (ADC)
+read digital signals, we needed to include an analogue-to-digital (ADC)
 converter. The amount of dust is measured by a Nova PM SDS011 High
-Precision Laser sensor. This sensor measures particles at 2.5 and 10
-micrometres in diameter in mg/m3. A list of the specific hardware and
-software components can be found in Appendix B.
+Precision Laser sensor. This sensor measures particles at 2.5 and 10 µm
+in diameter in mg/m\textsuperscript{3}. A list of the specific hardware
+and software components can be found in Appendix B.
 
@@
 
+\hypertarget{computational-reproducibility-and-data-availability}{%
+\subsubsection{Computational reproducibility and data
+availability}\label{computational-reproducibility-and-data-availability}}
+
+R Statistical Software version 4.2.1, RStudio IDE version 2023.3.0.386,
+and Quarto scientific publishing system version were used for
+quantitative data analysis and writing of the manuscript
+\citep{R-base, allaire2022quartoa, positteam2023rstudio}. A set of
+additional R packages were used for data wrangling, analysis, and
+visualisation
+\citep{ggplot22016, lubridate2011, R-dplyr, R-forcats, R-ggplot2, R-janitor, R-leaflet, R-lubridate, R-mapview, R-readr, R-tidyr, R-waffle}.
+
+Raw data and analysis-ready processed data is available as an R package
+at \citet{R-qechairquality}. The data underlying the tables and figures
+of this manuscript are contained in a repository alongside a
+reproducible document that contains the analysis code and the written
+narrative of the manuscript {[}TODO: manuscript software citation{]}.
+
 \hypertarget{theorycalculation-not-applicable}{%
@@
 \citep{kutlarjoss2017time, mapoma2014air} indicate that the maximum 24-h
-PM10 value is 25 μg/m3 and the maximum annual PM2.5 value is 8 μg/m3.
-There is no daily maximum value for PM2.5. It should be noted that both
-of these values are at, or below, the level-4 interim WHO targets.
+PM\textsubscript{10} value is 25 µg/m\textsuperscript{3} and the maximum
+annual PM\textsubscript{2.5} value is 8 µg/m\textsuperscript{3}. There
+is no daily maximum value for PM\textsubscript{2.5}. It should be noted
+that both of these values are at, or below, the level-4 interim WHO
+targets.
 
@@
 \endhead
-Administration & 10900 & 171 & 94 \\
-Guardian Shelter & 11032 & 4529 & 5314 \\
-Lighthouse Clinic & 11472 & 4365 & 3891 \\
-Lions Sight & 11749 & 91 & 14 \\
-Malaria & 12031 & 617 & 578 \\
-Mercy James & 12193 & 736 & 749 \\
-Ward 6B & 11861 & 1113 & 1236 \\
-uMoyo & 10637 & 703 & 399 \\
+Administration & 10900 & 171 & 0 \\
+Guardian Shelter & 11032 & 4529 & 0 \\
+Lighthouse Clinic & 11472 & 4365 & 0 \\
+Lions Sight & 11749 & 91 & 0 \\
+Malaria & 12031 & 617 & 0 \\
+Mercy James & 12193 & 736 & 0 \\
+Ward 6B & 11861 & 1113 & 0 \\
+uMoyo & 10637 & 703 & 0 \\
 \bottomrule()
@@
 Guardian Shelter recorded more than 50 instances above the hazardous
-limits. At the Lighthouse Clinic, the PM10 limit was exceeded almost
-twice as often as the PM2.5 limit, while at the Guardian Shelter, the
-reverse is true. Because the Lighthouse Clinic is within 50 m of the
-incinerator, the number and predominance of PM10 peaks is characteristic
-of incomplete combustion and dust that is typical for the area and the
-open burning that occurs there. The residents at the Guardian Shelter
-however are mostly cooking within contained clay stoves that have been
-designed to improve combustion of the coal and wood that is burned
-within them. The fewer number of PM10 peaks is a testament to this
-intervention, though the frequency of PM2.5 peaks is still beyond
-acceptable.
+limits. At the Lighthouse Clinic, the PM\textsubscript{10} limit was
+exceeded almost twice as often as the PM\textsubscript{2.5} limit, while
+at the Guardian Shelter, the reverse is true. Because the Lighthouse
+Clinic is within 50 m of the incinerator, the number and predominance of
+PM\textsubscript{10} peaks is characteristic of incomplete combustion
+and dust that is typical for the area and the open burning that occurs
+there. The residents at the Guardian Shelter however are mostly cooking
+within contained clay stoves that have been designed to improve
+combustion of the coal and wood that is burned within them. The fewer
+number of PM\textsubscript{10} peaks is a testament to this
+intervention, though the frequency of PM\textsubscript{2.5} peaks is
+still beyond acceptable.
 
@@
 The 24-hour averaged values (logged every 5 minutes) are presented for
-each location, for both PM10 and PM2.5 in fig-24-hour-average. As well
-as dampening local variation and peaks, the WHO air quality guidelines
-are also based on 24-hour averages which are the standard against which
-the health risks can be judged.
-
-Overall, PM2.5 values remained below 100 μg/m3 at 6 of 8 locations
-(Administration, Lions Sight, Malaria Project, uMoyo, Ward 6B, and Mercy
-James); PM10 values were consistently below 100μg/m3 at the same
-locations, but with several average values extending slightly above, and
-then infrequently.
+each location, for both PM\textsubscript{10} and PM\textsubscript{2.5}
+in fig-24-hour-average. As well as dampening local variation and peaks,
+the WHO air quality guidelines are also based on 24-hour averages which
+are the standard against which the health risks can be judged.
+
+Overall, PM\textsubscript{2.5} values remained below 100
+µg/m\textsuperscript{3} at 6 of 8 locations (Administration, Lions
+Sight, Malaria Project, uMoyo, Ward 6B, and Mercy James);
+PM\textsubscript{10} values were consistently below
+100µg/m\textsuperscript{3} at the same locations, but with several
+average values extending slightly above, and then infrequently.
 
@@
 
-\caption{\label{fig-24-hour-average}Average 24-hour PM10 and PM2.5 at 8
-monitoring stations over 2 months}
+\caption{\label{fig-24-hour-average}Average 24-hour PM\textsubscript{10}
+and PM\textsubscript{2.5} at 8 monitoring stations over 2 months}
 
@@
 Given that the type of fuel and the type of burning (contained vs.~open)
-produce very different particulate ``fingerprints'', the ratio of PM10
-to PM2.5 values were examined to determine if a clear difference between
-locations, and therefore source, could be identified. The results are
-presented in Figure~\ref{fig-pm-ratios}.
+produce very different particulate ``fingerprints'', the ratio of
+PM\textsubscript{10} to PM\textsubscript{2.5} values were examined to
+determine if a clear difference between locations, and therefore source,
+could be identified. The results are presented in
+Figure~\ref{fig-pm-ratios}.
 
@@
 
-\caption{\label{fig-pm-ratios}PM10:PM2.5 values for each location. Each
-line shows the minimum and maximum ratio for the day based on ratios
-calculated hourly. Does not include 14th to 16th October at location
-uMoyo due to extreme outliers.}
+\caption{\label{fig-pm-ratios}PM\textsubscript{10}:PM\textsubscript{2.5}
+values for each location. Each line shows the minimum and maximum ratio
+for the day based on ratios calculated hourly. Does not include 14th to
+16th October at location uMoyo due to extreme outliers.}
 
@@
 
-The results presented in Figure 5 illustrate both the presence of PM10
-relative to PM2.5 as well as the range of values over which that ratio
-is observed.
+The results presented in Figure 5 illustrate both the presence of
+PM\textsubscript{10} relative to PM\textsubscript{2.5} as well as the
+range of values over which that ratio is observed.
 
 Unlike the peak values presented in Table 2, the values presented in
-Figure 5 indicate the relative presence of larger (PM10) particles to
-smaller ones (PM2.5) regardless of their concentration: the ratio of two
-low concentrations can be the same as the ratio of two high
-concentrations and is therefore more indicative of source than of
-proximity.
+Figure 5 indicate the relative presence of larger (PM\textsubscript{10})
+particles to smaller ones (PM\textsubscript{2.5}) regardless of their
+concentration: the ratio of two low concentrations can be the same as
+the ratio of two high concentrations and is therefore more indicative of
+source than of proximity.
 
@@
 composition of emissions recorded ranged significantly, analysed as the
-ratio PM10:PM2.5 likely reflects a range of burning styles and/or fuel
-(waste) type, and no discernable trend was observed.. Among interview
-respondents, there was a general consensus that air quality caused by
-the burning of waste within the hospital was a problem, although there
-were significant differences between respondents over their
-understandings of potential impacts, the effectiveness of various coping
-mechanisms, and their problematisations linked to the burning of
-specific waste materials. For most, going home or leaving the Queen's
-campus was the only pathway to relief, though for patients who can not
-leave, endurance was the only option. This work was originally planned
-as a before-and-after study: we envisioned reporting these baseline
-results compared to follow-up measurements once the long-promised
-incinerator was commissioned. However, nearly 4 years later, the
-incinerator is rarely functional (partly for technical reasons, partly
-for financial ones) and the smoke hovering over Queen's is still thick.
-Waste management has long been known as an environmental health issue,
-but in the case of open trash burning at a medical facility, it becomes
-more: for some of the most vulnerable, especially those with TB, asthma
-or other chronic disease, it could be a matter of life or death.
+ratio PM\textsubscript{10}:PM\textsubscript{2.5} likely reflects a range
+of burning styles and/or fuel (waste) type, and no discernible trend was
+observed.. Among interview respondents, there was a general consensus
+that air quality caused by the burning of waste within the hospital was
+a problem, although there were significant differences between
+respondents over their understandings of potential impacts, the
+effectiveness of various coping mechanisms, and their problematisations
+linked to the burning of specific waste materials. For most, going home
+or leaving the Queen's campus was the only pathway to relief, though for
+patients who can not leave, endurance was the only option. This work was
+originally planned as a before-and-after study: we envisioned reporting
+these baseline results compared to follow-up measurements once the
+long-promised incinerator was commissioned. However, nearly 4 years
+later, the incinerator is rarely functional (partly for technical
+reasons, partly for financial ones) and the smoke hovering over Queen's
+is still thick. Waste management has long been known as an environmental
+health issue, but in the case of open trash burning at a medical
+facility, it becomes more: for some of the most vulnerable, especially
+those with TB, asthma or other chronic disease, it could be a matter of
+life or death.
 
@@
 
-\caption{\label{fig-five-minute-intervals}PM2.5 and PM10 values
-collected every 5 minutes over 10 months at 8 locations.}
+\caption{\label{fig-five-minute-intervals}PM\textsubscript{2.5} and
+PM\textsubscript{10} values collected every 5 minutes over 10 months at
+8 locations.}
 
@@
 \item
-  analog-to-digital converter (ADC)
+  analogue-to-digital converter (ADC)
 \end{itemize}
@@
 \item
-  any box that can fit all the components comfortbaly can be used. The
+  any box that can fit all the components comfortably can be used. The
   box will need to have a whole drilled in it for the pipe.
@@
   Once the whole device has been put together and placed on site, to
-  access that data we use a combination of propgrams:
+  access that data we use a combination of programs:
 \end{enumerate}
@@
   Set up the Raspberry Pi with the Rasbian software using the NOOBS
-  scandisk. (insert the scandisk into the Rasberry Pi) A comprehensive
+  scandisk. (insert the scandisk into the Raspberry Pi) A comprehensive
   start up guide can be found here.
@@
 \renewcommand\refname{References}
-  \bibliography{man-qechairquality.bib}
+  \bibliography{man-qechairquality.bib,packages.bib}
 

### 2023-04-11 manuscript/manuscript.qmd
Sync current google version

@@
 Low-cost sensors measuring particulate matter (PM) with particle sizes less than 10 µm (PM~10~) and less than 2.5 µm (PM~2.5~), expressed as the mass of PM per volume of air (µg PMx/m^3^ air) were recorded every 5 minutes at 8 locations across the QECH for 2 months.
-Qualitative data collection consisted of 57 interviews with patients, caregivers and hospital staff (including janitorial and maintenance staff, nurses, doctors, and administrators).
+Qualitative data collection consisted of 56 interviews with patients, caregivers and hospital staff (including janitorial and maintenance staff, nurses, doctors, and administrators).
 Our results show that safe air quality thresholds are consistently exceeded across space and time and that the most problematic air quality surrounds the shelter for caregivers and those receiving treatment for HIV/AIDs.
@@
 
-Qualitative data collection consisted of 26 interviews with caregivers
+Qualitative data collection consisted of 25 interviews with caregivers
 and hospital staff (including janitorial and maintenance staff, nurses,
@@
 well as an additional 31 interviews conducted with staff, patients, and
-caregivers in January and February 2021. Interviews were
+caregivers in January and February 2021, spread around the hospital’s various wards and departments. Interviews were
 semi-structured, and included questions on hospital waste management
 practices, perceived best practices, and workplace health, safety, and
-risk. However, the semi-structured nature of the interviews allowed for
+risk.
+The first round of interviews included specific questions on air quality, while the second round focused on waste management more broadly.
+However, the semi-structured nature of the interviews allowed for
 the participants to raise topics of interest, and for unexpected themes
@@
 
-Most staff, however, have been unable to find anyway to mitigate the
+Most staff, however, have been unable to find any way to mitigate the
 impacts of the smoke, and only found relief once they reached home at

### 2023-04-12 R/00-setup.R
Prepare ggplot theme

@@ -1,0 +1,63 @@
+# description -------------------------------------------------------------
+
+
+
+# packages ---------------------------------------------------------------
+
+library(ggplot2)
+library(scales)
+
+# theme adapted from:
+# https://github.com/teunbrand/ggplot_tricks
+# https://github.com/hrbrmstr/hrbrthemes/blob/master/R/theme-ipsum.r
+
+axis_title_just = "rt"
+xj <- switch(tolower(substr(axis_title_just, 1, 1)), b=0, l=0, m=0.5, c=0.5, r=1, t=1)
+yj <- switch(tolower(substr(axis_title_just, 2, 2)), b=0, l=0, m=0.5, c=0.5, r=1, t=1)
+
+base_size = 7
+axis_title_size = 6
+
+theme_set(
+  # Pick a starting theme
+  theme_gray(base_family = "Arial",
+             base_size = base_size) +
+    # Add your favourite elements
+    theme(
+      axis.line = element_line(),
+      axis.line.x = element_blank(),
+      axis.line.y = element_blank(),
+      axis.ticks = element_blank(),
+      axis.title=element_text(size=axis_title_size),
+      #axis.title.x=element_text(hjust=xj),
+      #axis.title.y=element_text(hjust=yj),
+      panel.background = element_rect(fill = "white"),
+      panel.grid.major = element_line("#cccccc", linewidth = 0.25),
+      legend.key = element_rect(fill = NA),
+      plot.margin = margin(2, 2, 2, 2)
+    )
+)
+
+# paramaters for ggsave
+
+param_dpi = 1000
+param_width_single = 90
+param_width_double = 190
+
+# test
+
+source("R/01-data-preparation.R")
+
+dat_in_sum_day |>
+  ggplot(aes(x = date, y = mean, group = indicator, color = indicator)) +
+  geom_point(size = 0.8) +
+  geom_line() +
+  labs(x = NULL) +
+  ylab(bquote(uq/m^3)) +
+  facet_wrap(~location, ncol = 4) +
+  scale_y_continuous(limits = c(0, 400)) +
+  scale_color_brewer(type = "qual", palette = 2)
+
+
+
+

### 2023-04-12 manuscript/manuscript.qmd
WIP

@@
 # Source data preparation script
+
+source(here::here("R/00-setup.R"))
 source(here::here("R/01-data-preparation.R"))
@@
 
-dat_in_sum_day |> 
+fig03 <- dat_in_sum_day |> 
   ggplot(aes(x = date, y = mean, group = indicator, color = indicator)) +
   geom_point(size = 0.8) +
-  geom_line() +
+  geom_line(alpha = 0.5) +
   labs(x = NULL) +
@@
   scale_y_continuous(limits = c(0, 400)) +
-  scale_color_brewer(type = "qual", palette = 2) + 
-  theme_bw(base_size = 14) + 
-  theme(panel.grid.minor = element_blank(),
-        panel.grid.major.x = element_blank())
+  scale_color_brewer(type = "qual", palette = 2)
+
+ggsave(filename = "manuscript/figs/fig-03-double-column.tiff",
+       plot = fig03,
+       dpi = param_dpi,
+       width = param_width_double,
+       height = param_width_double * width_height_ratio,
+       units = "mm")
 
@@
 
-dat_in_ratios |> 
+library(ggplot2)
+
+# create sample data
+df <- data.frame(x = 1:10, y = 1:10)
+
+# create plot
+ggplot(df, aes(x = x, y = y)) +
+  geom_point() +
+  xlab(expression(c('PM'[10], 'PM'[2.5])))
+
+fig04 <- dat_in_ratios |> 
   ggplot(aes(x = date)) +
   geom_linerange(aes(ymin = min, ymax = max)) +
-  labs(x = "hour") +
-  facet_wrap(~location, ncol = 2)
+  labs(x = "day",
+       y = "ratio") +
+  facet_wrap(~location, ncol = 2) +
+  scale_y_continuous(limits = c(0, 10))
+
+fig04
+
+ggsave(filename = "manuscript/figs/fig-04-double-column.tiff",
+       plot = fig04,
+       dpi = param_dpi,
+       width = param_width_double,
+       height = param_width_double * width_height_ratio,
+       units = "mm")
+
 ```
@@
 
-dat_air_quality_sum_who_2021 |> 
+percent_exposure <- dat_air_quality_sum_who_2021 |> 
   #filter(indicator == "PM~2.5~") |> 
@@
     TRUE ~ exposure_percent
-  )) |>
+  ))
 
@@
 
-ggplot(aes(fill = air_quality_who_annual, values = exposure_percent)) +
+fig05 <- ggplot(data = percent_exposure,
+                aes(fill = air_quality_who_annual, values = exposure_percent)) +
   geom_waffle(color ="white", size = 0.5, n_rows = 5, na.rm = TRUE) +
@@
         strip.text.y = element_text(vjust = 0, size = 4))
+
+fig05
+
 ```

### 2023-04-18 manuscript/manuscript.qmd
Add author
Finalise figures for TIFF export

@@
   - name: Heiko Heilgendorff
-    email: me@me.com
+    orcid: 0009-0004-1389-5541
+    email: heiko.heilgendorff@gmail.com
     affiliations: 
@@
   warning: false
+  freeze: true
 editor: 
@@
 
-Built in 1964, Malawi's year of independence, Queen Elizabeth Central Hospital
-(QECH), or Queen's, as it's known locally, is the country's largest
-public hospital. Designed in the Tropical Modernism style of
+Built in 1964, Malawi's year of independence, Queen Elizabeth Central
+Hospital (QECH), or Queen's, as it's known locally, is the country's
+largest public hospital. Designed in the Tropical Modernism style of
 architecture popular within Africa in the late colonial period, Queen's
@@
 the incinerator building, sending up the aforementioned clouds of smoke,
-which choke the grounds (@fig-open-burning). Innumerable other, smaller, fires,
-scattered across the QECH campus (@fig-map-sensors) contribute their part, as
-grounds and maintenance staff habitually burn leaves and other garden
-refuse. Furthermore, caregivers[^3] who reside separately on hospital
-grounds, cook for themselves throughout the day and night using biomass
-(charcoal or wood), and burn their own domestic waste. As a result, air
-quality at the hospital is a constant source of discussion, with
-patients, staff, and caregivers struggling to cope, when just breathing
-is, as the quote used in the title of this article, 'unbearable.'
+which choke the grounds (@fig-open-burning). Innumerable other, smaller,
+fires, scattered across the QECH campus (@fig-map-sensors) contribute
+their part, as grounds and maintenance staff habitually burn leaves and
+other garden refuse. Furthermore, caregivers[^3] who reside separately
+on hospital grounds, cook for themselves throughout the day and night
+using biomass (charcoal or wood), and burn their own domestic waste. As
+a result, air quality at the hospital is a constant source of
+discussion, with patients, staff, and caregivers struggling to cope,
+when just breathing is, as the quote used in the title of this article,
+'unbearable.'
 
@@
 and generally classified into PM with particle sizes less than 10 µm
-(PM~10~) and those with sizes less than 2.5 µm (PM~2.5~), while one way of
-measuring the severity of air pollution is by expressing the mass of PM
-per volume of air (µg PMx/m^3^ air). PM has been linked to multiple
+(PM~10~) and those with sizes less than 2.5 µm (PM~2.5~), while one way
+of measuring the severity of air pollution is by expressing the mass of
+PM per volume of air (µg PMx/m^3^ air). PM has been linked to multiple
 negative health outcomes. Amongst particulate air pollution, PM~10~ and
@@
 locations. Over the course of the two-month study, the hazardous limits
-for both PM~10~ and PM~2.5~ were exceeded at all locations except for the
-Lions Sight hospital. The limits were exceeded fewer than 50 times at
-five of the locations, but the monitors at the Lighthouse Clinic and the
-Guardian Shelter recorded more than 50 instances above the hazardous
+for both PM~10~ and PM~2.5~ were exceeded at all locations except for
+the Lions Sight hospital. The limits were exceeded fewer than 50 times
+at five of the locations, but the monitors at the Lighthouse Clinic and
+the Guardian Shelter recorded more than 50 instances above the hazardous
 limits. The extremely hazardous air quality at the Guardian shelter is
@@
 converter. The amount of dust is measured by a Nova PM SDS011 High
-Precision Laser sensor. This sensor measures particles at 2.5 and 10
-µm in diameter in mg/m^3^. A list of the specific hardware and
-software components can be found in Appendix B.
+Precision Laser sensor. This sensor measures particles at 2.5 and 10 µm
+in diameter in mg/m^3^. A list of the specific hardware and software
+components can be found in Appendix B.
 
@@
 well as an additional 31 interviews conducted with staff, patients, and
-caregivers in January and February 2021, spread around the hospital’s various wards and departments. Interviews were
-semi-structured, and included questions on hospital waste management
-practices, perceived best practices, and workplace health, safety, and
-risk.
-The first round of interviews included specific questions on air quality, while the second round focused on waste management more broadly.
-However, the semi-structured nature of the interviews allowed for
-the participants to raise topics of interest, and for unexpected themes
-to emerge. These interviews are supplemented by participatory
-observation recorded within detailed field notes, and supported by
-nearly a decade of sustained involvement and activism by the authors in
-the waste management practices of the hospital.
+caregivers in January and February 2021, spread around the hospital's
+various wards and departments. Interviews were semi-structured, and
+included questions on hospital waste management practices, perceived
+best practices, and workplace health, safety, and risk. The first round
+of interviews included specific questions on air quality, while the
+second round focused on waste management more broadly. However, the
+semi-structured nature of the interviews allowed for the participants to
+raise topics of interest, and for unexpected themes to emerge. These
+interviews are supplemented by participatory observation recorded within
+detailed field notes, and supported by nearly a decade of sustained
+involvement and activism by the authors in the waste management
+practices of the hospital.
 
@@
 
-R Statistical Software version 4.2.1, RStudio IDE version 2023.3.0.386, and Quarto scientific publishing system version were used for quantitative data analysis and writing of the manuscript [@R-base; @allaire2022quartoa; @positteam2023rstudio].
-A set of additional R packages were used for data wrangling, analysis, and visualisation [@ggplot22016; @lubridate2011;  @R-dplyr; @R-forcats; @R-ggplot2; @R-janitor; @R-leaflet; @R-lubridate; @R-mapview; @R-readr; @R-tidyr; @R-waffle].
-
-Raw data and analysis-ready processed data is available as an R package at @R-qechairquality.
-The data underlying the tables and figures of this manuscript are contained in a repository alongside a reproducible document that contains the analysis code and the written narrative of the manuscript [TODO: manuscript software citation].
+R Statistical Software version 4.2.1, RStudio IDE version 2023.3.0.386,
+and Quarto scientific publishing system version were used for
+quantitative data analysis and writing of the manuscript [@R-base;
+@allaire2022quartoa; @positteam2023rstudio]. A set of additional R
+packages were used for data wrangling, analysis, and visualisation
+[@ggplot22016; @lubridate2011; @R-dplyr; @R-forcats; @R-ggplot2;
+@R-janitor; @R-leaflet; @R-lubridate; @R-mapview; @R-readr; @R-tidyr;
+@R-waffle].
+
+Raw data and analysis-ready processed data is available as an R package
+at @R-qechairquality. The data underlying the tables and figures of this
+manuscript are contained in a repository alongside a reproducible
+document that contains the analysis code and the written narrative of
+the manuscript \[TODO: manuscript software citation\].
 
@@
 [@kutlarjoss2017time; @mapoma2014air] indicate that the maximum 24-h
-PM~10~ value is 25 µg/m^3^ and the maximum annual PM~2.5~ value is 8 µg/m^3^.
-There is no daily maximum value for PM~2.5~. It should be noted that both
-of these values are at, or below, the level-4 interim WHO targets.
+PM~10~ value is 25 µg/m^3^ and the maximum annual PM~2.5~ value is 8
+µg/m^3^. There is no daily maximum value for PM~2.5~. It should be noted
+that both of these values are at, or below, the level-4 interim WHO
+targets.
 
@@
 persistent trends. The full set of 5-minute data are plotted and
-presented in Appendix A @fig-five-minute-intervals. However, the number of measured values
-that exceeded the hazardous limit for both parameters are summarised in
-@tbl-peaks.
+presented in Appendix A @fig-five-minute-intervals. However, the number
+of measured values that exceeded the hazardous limit for both parameters
+are summarised in @tbl-peaks.
 
@@
 reverse is true. Because the Lighthouse Clinic is within 50 m of the
-incinerator, the number and predominance of PM~10~ peaks is characteristic
-of incomplete combustion and dust that is typical for the area and the
-open burning that occurs there. The residents at the Guardian Shelter
-however are mostly cooking within contained clay stoves that have been
-designed to improve combustion of the coal and wood that is burned
-within them. The fewer number of PM~10~ peaks is a testament to this
-intervention, though the frequency of PM~2.5~ peaks is still beyond
-acceptable.
+incinerator, the number and predominance of PM~10~ peaks is
+characteristic of incomplete combustion and dust that is typical for the
+area and the open burning that occurs there. The residents at the
+Guardian Shelter however are mostly cooking within contained clay stoves
+that have been designed to improve combustion of the coal and wood that
+is burned within them. The fewer number of PM~10~ peaks is a testament
+to this intervention, though the frequency of PM~2.5~ peaks is still
+beyond acceptable.
 
@@
 The 24-hour averaged values (logged every 5 minutes) are presented for
-each location, for both PM~10~ and PM~2.5~ in fig-24-hour-average. As well as dampening
-local variation and peaks, the WHO air quality guidelines are also based
-on 24-hour averages which are the standard against which the health
-risks can be judged.
+each location, for both PM~10~ and PM~2.5~ in fig-24-hour-average. As
+well as dampening local variation and peaks, the WHO air quality
+guidelines are also based on 24-hour averages which are the standard
+against which the health risks can be judged.
 
@@
 
-ggsave(filename = "manuscript/figs/fig-03-double-column.tiff",
+fig03
+
+ggsave(filename = here::here("manuscript/figs/fig-03-double-column.tiff"),
        plot = fig03,
@@
 Given that the type of fuel and the type of burning (contained vs. open)
-produce very different particulate "fingerprints", the ratio of PM~10~ to
-PM~2.5~ values were examined to determine if a clear difference between
-locations, and therefore source, could be identified. The results are
-presented in @fig-pm-ratios.
+produce very different particulate "fingerprints", the ratio of PM~10~
+to PM~2.5~ values were examined to determine if a clear difference
+between locations, and therefore source, could be identified. The
+results are presented in @fig-pm-ratios.
 
@@
 
-library(ggplot2)
-
-# create sample data
-df <- data.frame(x = 1:10, y = 1:10)
-
-# create plot
-ggplot(df, aes(x = x, y = y)) +
-  geom_point() +
-  xlab(expression(c('PM'[10], 'PM'[2.5])))
-
 fig04 <- dat_in_ratios |> 
@@
 
-ggsave(filename = "manuscript/figs/fig-04-double-column.tiff",
+ggsave(filename = here::here("manuscript/figs/fig-04-double-column.tiff"),
        plot = fig04,
@@
 Figure 5 indicate the relative presence of larger (PM~10~) particles to
-smaller ones (PM~2.5~) regardless of their concentration: the ratio of two
-low concentrations can be the same as the ratio of two high
+smaller ones (PM~2.5~) regardless of their concentration: the ratio of
+two low concentrations can be the same as the ratio of two high
 concentrations and is therefore more indicative of source than of
@@
 The relative amount of time exposed to a given category of air quality
-is shown in @fig-percent-exposure. Each square represents 1% of the measured time.
+is shown in @fig-percent-exposure. Each square represents 1% of the
+measured time.
 
@@
     location == "Administration" 
-    & indicator == "pm25"
+    & indicator == "pm2.5"
     & air_quality_who_annual == "Interim target 3" ~ exposure_percent + 1,
 
+    location == "Guardian Shelter" 
+    & indicator == "pm10"
+    & air_quality_who_annual == "Interim target 4" ~ exposure_percent + 1,
+
     location == "uMoyo" 
-    & indicator == "pm25"
+    & indicator == "pm2.5"
     & air_quality_who_annual == "Interim target 4" ~ exposure_percent - 1,
@@
 
-fig05 <- ggplot(data = percent_exposure,
+fig05 <- 
+ggplot(data = percent_exposure,
        aes(fill = air_quality_who_annual, values = exposure_percent)) +
@@
   #facet_wrap(~location, ncol = 2) +
-  theme_minimal(base_size = 14) +
+  #theme_minimal(base_size = 14) +
   # help from: https://stackoverflow.com/questions/34749636/left-justify-text-from-multi-line-facet-labels
   theme(strip.text.x = element_text(hjust = 0),
-        strip.text.y = element_text(vjust = 0, size = 4))
+        strip.text.y = element_text(vjust = 0, size = 4),
+        legend.position = "bottom",
+        legend.title = element_blank(),
+        legend.text = element_text(size = 4)) #+
 
 fig05
+
+ggsave(filename = here::here("manuscript/figs/fig-05-double-column.tiff"),
+       plot = fig05,
+       dpi = param_dpi,
+       width = param_width_double,
+       height = param_width_double * width_height_ratio,
+       units = "mm")
 
@@
 > vaccinated and I am protected including other staff. However, we
-> realize that the air can badly affect other people and patients who
+> realise that the air can badly affect other people and patients who
 > come to this place.
@@
 
-The air quality, as a result of open waste burning at Queen’s is poor and not suitable for a city, let alone a hospital.
-Over the course of this two-month study, the hazardous limits for both PM parameters were exceeded at all locations except for the Lions Sight hospital.
-The WHO limits were exceeded fewer than 50 times at five of the locations, and only the monitors at the Lighthouse Clinic and the Guardian Shelter recorded more than 50 instances above the most hazardous limit.
-The composition of emissions recorded ranged significantly, analysed as the ratio PM~10~:PM~2.5~ likely reflects a range of burning styles and/or fuel (waste) type, and no discernible trend was observed..
-Among interview respondents, there was a general  consensus that air quality caused by the burning of waste within the hospital was a problem, although there were significant differences between respondents over their understandings of potential impacts, the effectiveness of various coping mechanisms, and their problematisations linked to the burning of specific waste materials.
-For most, going home or leaving the Queen’s campus was the only pathway to relief, though for patients who can not leave, endurance was the only option.
-This work was originally planned as a before-and-after study:  we envisioned reporting these baseline results compared to follow-up measurements once the long-promised incinerator was commissioned.
-However, nearly 4 years later, the incinerator is rarely functional (partly for technical reasons, partly for financial ones) and the smoke hovering over Queen’s is still thick.
-Waste management has long been known as an environmental health issue, but in the case of open trash burning at a medical facility, it becomes more:  for some of the most vulnerable, especially those with TB, asthma or other chronic disease, it could be a matter of life or death.
+The air quality, as a result of open waste burning at Queen's is poor
+and not suitable for a city, let alone a hospital. Over the course of
+this two-month study, the hazardous limits for both PM parameters were
+exceeded at all locations except for the Lions Sight hospital. The WHO
+limits were exceeded fewer than 50 times at five of the locations, and
+only the monitors at the Lighthouse Clinic and the Guardian Shelter
+recorded more than 50 instances above the most hazardous limit. The
+composition of emissions recorded ranged significantly, analysed as the
+ratio PM~10~:PM~2.5~ likely reflects a range of burning styles and/or
+fuel (waste) type, and no discernible trend was observed.. Among
+interview respondents, there was a general consensus that air quality
+caused by the burning of waste within the hospital was a problem,
+although there were significant differences between respondents over
+their understandings of potential impacts, the effectiveness of various
+coping mechanisms, and their problematisations linked to the burning of
+specific waste materials. For most, going home or leaving the Queen's
+campus was the only pathway to relief, though for patients who can not
+leave, endurance was the only option. This work was originally planned
+as a before-and-after study: we envisioned reporting these baseline
+results compared to follow-up measurements once the long-promised
+incinerator was commissioned. However, nearly 4 years later, the
+incinerator is rarely functional (partly for technical reasons, partly
+for financial ones) and the smoke hovering over Queen's is still thick.
+Waste management has long been known as an environmental health issue,
+but in the case of open trash burning at a medical facility, it becomes
+more: for some of the most vulnerable, especially those with TB, asthma
+or other chronic disease, it could be a matter of life or death.
 
@@
 
-The authors are grateful to the staff, patients and caregivers at QECH for their assistance, willingness to share their views, and importantly, perseverance in delivering the best care possible under difficult circumstances.
-Moreover, we would like to acknowledge the assistance and effort of Open Data Durban, now the Open Cities Lab, for their assistance in constructing and programming the sensors.
+The authors are grateful to the staff, patients and caregivers at QECH
+for their assistance, willingness to share their views, and importantly,
+perseverance in delivering the best care possible under difficult
+circumstances. Moreover, we would like to acknowledge the assistance and
+effort of Open Data Durban, now the Open Cities Lab, for their
+assistance in constructing and programming the sensors.
 
@@
 
-dat_in |> 
+figA1 <- dat_in |> 
   ggplot(aes(x = date_time, y = value)) +
   geom_line(alpha = 0.5) +
-  labs(title = "Hospital Air Quality",
-       subtitle = "PM~2.5~ and PM~10~ measurements in 5-min intervals at eight locations") +
+  #labs(title = "Hospital Air Quality",
+  #     subtitle = "PM~2.5~ and PM~10~ measurements in 5-min intervals at eight locations") +
   ylab(bquote(uq/m^3)) +
+  labs(x = NULL) +
   facet_grid(indicator~location) +
-  theme_bw(base_size = 14) + 
   theme(panel.grid.minor = element_blank(),
@@
 
+figA1
+
+ggsave(filename = here::here("manuscript/figs/fig-A1-double-column.tiff"),
+       plot = figA1,
+       dpi = param_dpi,
+       width = param_width_double,
+       height = param_width_double * width_height_ratio,
+       units = "mm")
+
 ```
@@
     specifications:
+
 -   Broadcom BCM2837 64bit Quad Core Processor powered Single Board
@@
     Amps)
--   Same form factor as the Pi 2 Model B, however the LEDs have
-changed position
+-   Same form factor as the Pi 2 Model B, however the LEDs have changed
+    position
 -   RPI3-MODB-16GB-NOOBS
+
 2.  Nova PM SDS011 Sensor
+
 -   See Data Sheet attachment for technical specifications
+
 3.  MCP3008 microchip
+
 -   analogue-to-digital converter (ADC)
+
 4.  Romoss Sense8+ 30000mAh QC Type-C
+
 -   Power bank battery
+
 5.  Pipe
+
 -   clear rubber pipe attainable from any local hardware store
+
 6.  Box
--   any box that can fit all the components comfortably can be used.
-The box will need to have a whole drilled in it for the pipe.
-
-### Software ‘components’ and data management {.unnumbered}
-
-1. The Raspberry Pi operating system runs off of Rasbian software on a NOOBS scandisk. 
-2.
-Air quality monitoring code is written in Python using the [aqi library](https://pypi.org/project/python-aqi/) (A library of algorithms to convert between AQI value and pollutant concentration).
-The code is stored on the [Open Data Durban GitHub account](https://github.com/opendatadurban/hospital_stations) and is free for all to access.
-To be able to collect data, the device needs to be connected to a reliable wifi network, the SSID and password need to be coded onto the device during step 2 in the building method.
-3. Once the whole device has been put together and placed on site, to access that data we use a combination of programs:
-- [Putty](https://www.putty.org/) - this is an SSH client that we use to gain access to the Raspberry Pi from another computer.
-To be able to access the Raspberry Pi in question, the computer in use and the air quality monitoring device need to be on the same network for the connection to be successful.
-- [Winscp](https://winscp.net/eng/download.php) - this is an FTP client that we use, in conjunction with Putty, to be able to transfer data files from the air quality monitoring device to one’s computer.
-- [Remote.it](https://app.remote.it/) - this is an application that supports remote SSH, so this is used when wanting to access data from a device that is not connected to the same network as the computer in use.
-Remote.it is used together with Putty and Winscp to access the air quality data remotely.
+
+-   any box that can fit all the components comfortably can be used. The
+    box will need to have a whole drilled in it for the pipe.
+
+### Software 'components' and data management {.unnumbered}
+
+1.  The Raspberry Pi operating system runs off of Rasbian software on a
+    NOOBS scandisk.
+2.  Air quality monitoring code is written in Python using the [aqi
+    library](https://pypi.org/project/python-aqi/) (A library of
+    algorithms to convert between AQI value and pollutant
+    concentration). The code is stored on the [Open Data Durban GitHub
+    account](https://github.com/opendatadurban/hospital_stations) and is
+    free for all to access. To be able to collect data, the device needs
+    to be connected to a reliable wifi network, the SSID and password
+    need to be coded onto the device during step 2 in the building
+    method.
+3.  Once the whole device has been put together and placed on site, to
+    access that data we use a combination of programs:
+
+-   [Putty](https://www.putty.org/) - this is an SSH client that we use
+    to gain access to the Raspberry Pi from another computer. To be able
+    to access the Raspberry Pi in question, the computer in use and the
+    air quality monitoring device need to be on the same network for the
+    connection to be successful.
+-   [Winscp](https://winscp.net/eng/download.php) - this is an FTP
+    client that we use, in conjunction with Putty, to be able to
+    transfer data files from the air quality monitoring device to one's
+    computer.
+-   [Remote.it](https://app.remote.it/) - this is an application that
+    supports remote SSH, so this is used when wanting to access data
+    from a device that is not connected to the same network as the
+    computer in use. Remote.it is used together with Putty and Winscp to
+    access the air quality data remotely.
 
@@
 
-1. Set up the Raspberry Pi with the Rasbian software using the NOOBS scandisk. (insert the scandisk into the Raspberry Pi) A comprehensive start up guide can be found here. 
-2.
-Next we load the code onto the Raspberry Pi.
-The code can be found on the Open Data Durban Hospital Stations Repository Github.
-To use it you can follow the steps in the file called Initiate_WS.txt.
-Once all the steps have been completed, the Raspberry Pi is now a device that can be connected to a particulate matter monitoring sensor.
+1.  Set up the Raspberry Pi with the Rasbian software using the NOOBS
+    scandisk. (insert the scandisk into the Raspberry Pi) A
+    comprehensive start up guide can be found here.
+2.  Next we load the code onto the Raspberry Pi. The code can be found
+    on the Open Data Durban Hospital Stations Repository Github. To use
+    it you can follow the steps in the file called Initiate_WS.txt. Once
+    all the steps have been completed, the Raspberry Pi is now a device
+    that can be connected to a particulate matter monitoring sensor.\
 3.  Connect the hardware components
-- Connect the Raspberry Pi to the Nova PM Sensor using the break out board that comes with the Sensor. 
+
+-   Connect the Raspberry Pi to the Nova PM Sensor using the break out
+    board that comes with the Sensor.
 -   Connect the battery to the Raspberry Pi via the micro port
 -   Attach the rubber pipe to the sensor
-- Assemble all components inside the box, with the pipe coming out of the drilled hole
-- To prevent dirt entering the system, close off all openings between the pipe and the box with silicone putty
-- Secure box in an elevated position (not too close to the ground) and with the pipe facing the direction of the inflow of the air under investigation.
+-   Assemble all components inside the box, with the pipe coming out of
+    the drilled hole
+-   To prevent dirt entering the system, close off all openings between
+    the pipe and the box with silicone putty
+-   Secure box in an elevated position (not too close to the ground) and
+    with the pipe facing the direction of the inflow of the air under
+    investigation.
 

### 2023-04-18 manuscript/manuscript.tex
Finalise figures for TIFF export

@@
 }
- \ead{me@me.com} 
+ \ead{heiko.heilgendorff@gmail.com} 
 \author[1]{Marc Kalina%
@@
 every 5 minutes at 8 locations across the QECH for 2 months. Qualitative
-data collection consisted of 57 interviews with patients, caregivers and
+data collection consisted of 56 interviews with patients, caregivers and
 hospital staff (including janitorial and maintenance staff, nurses,
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, frame hidden, boxrule=0pt, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, interior hidden, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 well as an additional 31 interviews conducted with staff, patients, and
-caregivers in January and February 2021. Interviews were
-semi-structured, and included questions on hospital waste management
-practices, perceived best practices, and workplace health, safety, and
-risk. However, the semi-structured nature of the interviews allowed for
-the participants to raise topics of interest, and for unexpected themes
-to emerge. These interviews are supplemented by participatory
-observation recorded within detailed field notes, and supported by
-nearly a decade of sustained involvement and activism by the authors in
-the waste management practices of the hospital.
+caregivers in January and February 2021, spread around the hospital's
+various wards and departments. Interviews were semi-structured, and
+included questions on hospital waste management practices, perceived
+best practices, and workplace health, safety, and risk. The first round
+of interviews included specific questions on air quality, while the
+second round focused on waste management more broadly. However, the
+semi-structured nature of the interviews allowed for the participants to
+raise topics of interest, and for unexpected themes to emerge. These
+interviews are supplemented by participatory observation recorded within
+detailed field notes, and supported by nearly a decade of sustained
+involvement and activism by the authors in the waste management
+practices of the hospital.
 
@@
 Personally, I have never experienced {[}eye discomfort{]} because I get
-vaccinated and I am protected including other staff. However, we realize
+vaccinated and I am protected including other staff. However, we realise
 that the air can badly affect other people and patients who come to this

### 2023-04-18 R/00-setup.R

@@ -20,7 +20,7 @@ axis_title_size = 6
 
 theme_set(
   # Pick a starting theme
-  theme_gray(base_family = "Arial",
+  theme_gray(#base_family = "Arial",
              base_size = base_size) +
     # Add your favourite elements
     theme(
@@ -28,7 +28,7 @@ theme_set(
       axis.line.x = element_blank(),
       axis.line.y = element_blank(),
       axis.ticks = element_blank(),
-      axis.title=element_text(size=axis_title_size),
+      legend.title = element_text(size = base_size),
       #axis.title.x=element_text(hjust=xj),
       #axis.title.y=element_text(hjust=yj),
       panel.background = element_rect(fill = "white"),
@@ -43,21 +43,26 @@ theme_set(
 param_dpi = 1000
 param_width_single = 90
 param_width_double = 190
+width_height_ratio = 190/275
 
 # test
 
-source("R/01-data-preparation.R")
+# source("R/01-data-preparation.R")
 
-dat_in_sum_day |>
-  ggplot(aes(x = date, y = mean, group = indicator, color = indicator)) +
-  geom_point(size = 0.8) +
-  geom_line() +
-  labs(x = NULL) +
-  ylab(bquote(uq/m^3)) +
-  facet_wrap(~location, ncol = 4) +
-  scale_y_continuous(limits = c(0, 400)) +
-  scale_color_brewer(type = "qual", palette = 2)
-
-
-
+#ggplot(data = percent_exposure,
+#       aes(fill = air_quality_who_annual, values = exposure_percent)) +
+#  geom_waffle(color ="white", size = 0.5, n_rows = 5, na.rm = TRUE) +
+#  coord_equal() +
+#  labs(fill = "WHO 2021 targets (annual)") +
+#  scale_fill_brewer(type = "div", palette = "RdYlBu", direction = -1) +
+#  scale_x_discrete(expand=c(0,0)) +
+#  scale_y_discrete(expand=c(0,0)) +
+#  facet_grid(location ~ indicator, switch = "y") +
+#
+#  #facet_wrap(~location, ncol = 2) +
+#  #theme_minimal(base_size = 14) +
+#  # help from: https://stackoverflow.com/questions/34749636/left-justify-text-from-multi-line-facet-labels
+#  theme(strip.text.x = element_text(hjust = 0),
+#        strip.text.y = element_text(vjust = 0))
+#
 

### 2023-04-18 manuscript/manuscript.qmd
Add footnote references

@@
 discussion, with patients, staff, and caregivers struggling to cope,
-when just breathing is, as the quote used in the title of this article,
-'unbearable.'
+when just breathing is, as the quote[^4] used in the title of this
+article, 'unbearable.'
 
@@
     bathing, and all other non-medical services.
+
+[^4]: ID5, 2019-11-18
 
@@
     research and patient care in the area of paediatric malaria,
-    specifically cerebral malaria, for more than 30 years[^4]. Its
+    specifically cerebral malaria, for more than 30 years[^5]. Its
     research infrastructure includes an administrative team, an
@@
 
-[^4]: BMP is credited with the development of the Blantyre Coma Score,
+[^5]: BMP is credited with the development of the Blantyre Coma Score,
     which is used widely.
@@
 Queens, were unable to access or analyse hospital policy, and generally
-tended to rely on "experience" and "human judgement" when handling
-potentially hazardous waste. Moreover, although nearly all doctors and
-nurses had prior education on waste management best practices, cleaners
-began their waste handling duties without any prior waste management
-experience, and although there is an orientation programme for new
-hires, there is no regular refresher training or capacity building
-regarding waste for existing staff, at any level.
+tended to rely on, what one cleaner[^6] described as, "experience" and
+"human judgement" when handling potentially hazardous waste. Moreover,
+although nearly all doctors and nurses had prior education on waste
+management best practices, cleaners began their waste handling duties
+without any prior waste management experience, and although there is an
+orientation programme for new hires, there is no regular refresher
+training or capacity building regarding waste for existing staff, at any
+level.
+
+[^6]: ID44, 2021-01-18
 
@@
 interviewed being able to recall having received a complaint, and in
-turn, complaining to administration. One of the staff members
+turn, complaining to administration. One of the staff[^7] members
 responsible for the burning said that he personally, had received
@@
 repaired.
+
+[^7]: ID3, 2019-11-18
 
@@
 breathing difficulties, and persistent cold and flu. Nausea was also
-mentioned, but was not a commonly described impact. Only one respondent,
-of the 26 total interviewed, did not describe lingering health impacts
-which they could ascribe to the smoke, however, they also described
-having chronic eye irritation, but did not believe the smoke was a
-contributing factor.
+mentioned, but was not a commonly described impact. Only one
+respondent[^8], of the 26 total interviewed, did not describe lingering
+health impacts which they could ascribe to the smoke, however, they also
+described having chronic eye irritation, but did not believe the smoke
+was a contributing factor.
+
+[^8]: ID1, 2019-11-18
 
@@
 belief among several respondents, including several nurses, that the
-smoke could be a cause of the disease itself. As one staff member
-(20/11/2019) stated, "I believe breathing this air for a long time can
-cause Tuberculosis." This, however, was only one of the few instances of
+smoke could be a cause of the disease itself. As one staff member[^9]
+stated, "I believe breathing this air for a long time can cause
+Tuberculosis." This, however, was only one of the few instances of
 misinformation which staff members held regarding air quality and
@@
 Nonetheless, they were concerned about the impacts of the smoke on
-patients and visitors, as one staff member (18/11/2019) expressed:
+patients and visitors, as one staff member[^10] expressed:
+
+[^9]: ID8, 2019-11-20
+
+[^10]: ID3, 2019-11-18
 
@@
 breaks away from hospital grounds in order to catch some breaths of
-fresher air. However, for janitorial staff, inside was not necessarily
-better, as several reported their indoor workspaces as being also
-insufferable for long periods of time from the smell of cleaning agents
-and other chemicals. Furthermore, amongst respondents there was a
+fresher air. However, for janitorial staff[^11], inside was not
+necessarily better, as several reported their indoor workspaces as being
+also insufferable for long periods of time from the smell of cleaning
+agents and other chemicals. Furthermore, amongst respondents there was a
 general disagreement over the effectiveness of personal protective
@@
 impacts of the smoke, describing them as ineffective.
+
+[^11]: ID14, 2019-11-23
 
@@
 the time of the interviews), and biding their time as construction drags
-on; coping as best they can. As one staff member (25/11/2019)
-described,"we are just hoping we will start breathing good air soon,
-when the new incinerator is opened."
+on; coping as best they can. As the same member of the janitorial
+staff[^12] described, "we are just hoping we will start breathing good
+air soon, when the new incinerator is opened."
+
+[^12]: ID14, 2019-11-23
 

### 2023-04-18 manuscript/manuscript.qmd
export docx

@@
 format:
+  docx: default
   elsevier-pdf:
@@
 
-(TODO: REPLACE screenshot)
+(TODO: REPLACE screenshot with actual data)
 
@@
 #| fig-align: center
+#| fig-height: 5
 
@@
 on; coping as best they can. As the same member of the janitorial
-staff[^12] described, "we are just hoping we will start breathing good
+staff[^11] described, "we are just hoping we will start breathing good
 air soon, when the new incinerator is opened."
 
-[^12]: ID14, 2019-11-23
+[^11]: ID14, 2019-11-23
 

### 2023-04-18 manuscript/manuscript.tex
export docx

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, interior hidden, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, breakable, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 patients, staff, and caregivers struggling to cope, when just breathing
-is, as the quote used in the title of this article, `unbearable.'
+is, as the quote\footnote{ID5, 2019-11-18} used in the title of this
+article, `unbearable.'
 
@@
 Queens, were unable to access or analyse hospital policy, and generally
-tended to rely on ``experience'' and ``human judgement'' when handling
-potentially hazardous waste. Moreover, although nearly all doctors and
-nurses had prior education on waste management best practices, cleaners
-began their waste handling duties without any prior waste management
-experience, and although there is an orientation programme for new
-hires, there is no regular refresher training or capacity building
-regarding waste for existing staff, at any level.
+tended to rely on, what one cleaner\footnote{ID44, 2021-01-18} described
+as, ``experience'' and ``human judgement'' when handling potentially
+hazardous waste. Moreover, although nearly all doctors and nurses had
+prior education on waste management best practices, cleaners began their
+waste handling duties without any prior waste management experience, and
+although there is an orientation programme for new hires, there is no
+regular refresher training or capacity building regarding waste for
+existing staff, at any level.
 
@@
 interviewed being able to recall having received a complaint, and in
-turn, complaining to administration. One of the staff members
-responsible for the burning said that he personally, had received
-hundreds of complaints, but was powerless to affect meaningful change,
-aside from burning at different hours, until the incinerator could be
-repaired.
+turn, complaining to administration. One of the staff\footnote{ID3,
+  2019-11-18} members responsible for the burning said that he
+personally, had received hundreds of complaints, but was powerless to
+affect meaningful change, aside from burning at different hours, until
+the incinerator could be repaired.
 
@@
 breathing difficulties, and persistent cold and flu. Nausea was also
-mentioned, but was not a commonly described impact. Only one respondent,
-of the 26 total interviewed, did not describe lingering health impacts
-which they could ascribe to the smoke, however, they also described
-having chronic eye irritation, but did not believe the smoke was a
-contributing factor.
+mentioned, but was not a commonly described impact. Only one
+respondent\footnote{ID1, 2019-11-18}, of the 26 total interviewed, did
+not describe lingering health impacts which they could ascribe to the
+smoke, however, they also described having chronic eye irritation, but
+did not believe the smoke was a contributing factor.
 
@@
 belief among several respondents, including several nurses, that the
-smoke could be a cause of the disease itself. As one staff member
-(20/11/2019) stated, ``I believe breathing this air for a long time can
-cause Tuberculosis.'' This, however, was only one of the few instances
-of misinformation which staff members held regarding air quality and
-health. Another example, voiced by several respondents, included a
-belief that some staff members were immune to ill-health impacts of the
-smoke, because they had received vaccinations from the hospital.
-Nonetheless, they were concerned about the impacts of the smoke on
-patients and visitors, as one staff member (18/11/2019) expressed:
+smoke could be a cause of the disease itself. As one staff
+member\footnote{ID8, 2019-11-20} stated, ``I believe breathing this air
+for a long time can cause Tuberculosis.'' This, however, was only one of
+the few instances of misinformation which staff members held regarding
+air quality and health. Another example, voiced by several respondents,
+included a belief that some staff members were immune to ill-health
+impacts of the smoke, because they had received vaccinations from the
+hospital. Nonetheless, they were concerned about the impacts of the
+smoke on patients and visitors, as one staff member\footnote{ID3,
+  2019-11-18} expressed:
 
@@
 breaks away from hospital grounds in order to catch some breaths of
-fresher air. However, for janitorial staff, inside was not necessarily
-better, as several reported their indoor workspaces as being also
-insufferable for long periods of time from the smell of cleaning agents
-and other chemicals. Furthermore, amongst respondents there was a
-general disagreement over the effectiveness of personal protective
-equipment (PPE), such as face masks, towards mitigating the impacts of
-the smoke. A few staff described pleading to hospital administration for
-such equipment, to no avail. However, other staff members, who do have
-access to PPE, noted that even face masks do little to mitigate the
-impacts of the smoke, describing them as ineffective.
+fresher air. However, for janitorial staff\footnote{ID14, 2019-11-23},
+inside was not necessarily better, as several reported their indoor
+workspaces as being also insufferable for long periods of time from the
+smell of cleaning agents and other chemicals. Furthermore, amongst
+respondents there was a general disagreement over the effectiveness of
+personal protective equipment (PPE), such as face masks, towards
+mitigating the impacts of the smoke. A few staff described pleading to
+hospital administration for such equipment, to no avail. However, other
+staff members, who do have access to PPE, noted that even face masks do
+little to mitigate the impacts of the smoke, describing them as
+ineffective.
 
@@
 the time of the interviews), and biding their time as construction drags
-on; coping as best they can. As one staff member (25/11/2019)
-described,``we are just hoping we will start breathing good air soon,
-when the new incinerator is opened.''
+on; coping as best they can. As the same member of the janitorial
+staff\footnote{ID14, 2019-11-23} described, ``we are just hoping we will
+start breathing good air soon, when the new incinerator is opened.''
 

### 2023-04-18 README.qmd
Write up README

---
format: gfm
execute: 
  echo: false
  warning: false
  freeze: true
editor: 
  markdown: 
    wrap: 72
bibliography: [manuscript/man-qechairquality.bib, manuscript/packages.bib]
---

## Manuscript: "'It Is Unbearable To Breath Here': Air Quality, Open Incineration, And Misinformation In Blantyre, Malawi"

This repository contains a reproducible manuscript for submission to the
journal [Habitat
International](https://www.sciencedirect.com/journal/habitat-international).

Raw data and analysis-ready processed data is available as an R package
at @R-qechairquality.

R Statistical Software version 4.2.1, RStudio IDE version 2023.3.0.386,
and Quarto scientific publishing system version were used for
quantitative data analysis and writing of the manuscript [@R-base;
@allaire2022quartoa; @positteam2023rstudio].

A set of additional R packages were used for data wrangling, analysis,
and visualisation [@ggplot22016; @lubridate2011; @R-dplyr; @R-forcats;
@R-ggplot2; @R-janitor; @R-leaflet; @R-lubridate; @R-mapview; @R-readr;
@R-tidyr; @R-waffle].

The Quarto template for Elsevier journals was used to prepared the final
submitted PDF document.

Figures were prepared using the [Elsevier guidelines for artwork and
media](https://www.elsevier.com/authors/policies-and-guidelines/artwork-and-media-instructions).
The exported figures in TIFF format exceeded 100 MB per file and are not
tracked in this repository. The automated exported figures from the
Quarto markdown file [are accessible in the
repository](https://github.com/Global-Health-Engineering/manuscript-qechairquality/tree/main/manuscript/manuscript_files/figure-pdf).

The data underlying the tables and figures of this manuscript are
contained in:

\[TODO: Add link to data\]

## References

### 2023-04-18 manuscript/manuscript.qmd
Fix email
Number fig and tbl labels

@@
     orcid: 0000-0002-4050-5121
-    email: kwangerlo@gmail.comm
+    email: kwangulero@gmail.com 
     affilations:
@@
 ```{r}
-#| label: tbl-peaks
+#| label: tbl01-peaks
 #| tbl-cap: Peaks for both PM~2.5~ and PM~10~ are the number of data points above the WHO interim target 1 (annual) which is the least stringent; the number of observations recorded (count) is provided for reference.
@@
 
-observations_location |> 
-  left_join(no_peaks_who) |> 
+tbl01_peaks <- observations_location |> 
+  left_join(no_peaks_who)
+
+tbl01_peaks |> 
+  write_csv(file = here::here("data/processed-data/tbl01-peaks.csv"))
+
+tbl01_peaks |> 
   knitr::kable()
+
 ```
@@
 ```{r}
-#| label: fig-24-hour-average
+#| label: fig03-24-hour-average
 #| fig-cap: Average 24-hour PM~10~ and PM~2.5~ at 8 monitoring stations over 2 months
@@
 
+dat_in_sum_day |> 
+  write_csv(here::here("data/processed-data/fig03-24-hour-average.csv"))
+
 ```
@@
 ```{r}
-#| label: fig-pm-ratios
+#| label: fig04-pm-ratios
 #| fig-cap: PM~10~:PM~2.5~ values for each location.
@@
 
+dat_in_ratios |> 
+  write_csv(file = here::here("data/processed-data/fig04-pm-ratios.csv"))
+
 ```
@@
 ```{r}
-#| label: fig-percent-exposure
+#| label: fig05-percent-exposure
 #| fig-cap: Percentage of measured values categorised according to WHO 2021 targets
@@
        units = "mm")
+
+percent_exposure |> 
+  write_csv("data/processed-data/fig05-percent-exposure.csv")
 
@@
 ```{r}
-#| label: fig-five-minute-intervals
+#| label: figA1-five-minute-intervals
 #| fig-cap: PM~2.5~ and PM~10~ values collected every 5 minutes over 10 months at 8 locations.

### 2023-04-18 README.qmd
Re-organise content

@@
 
+## Data
+
 Raw data and analysis-ready processed data is available as an R package
-at @R-qechairquality.
+at: https://github.com/Global-Health-Engineering/qechairquality [@R-qechairquality].
+
+The data underlying the tables and figures of this manuscript are
+contained in this repo in [data/processed-data/](https://github.com/Global-Health-Engineering/manuscript-qechairquality/tree/main/data/processed-data).
+
+## Figures
+
+Figures were prepared using the [Elsevier guidelines for artwork and
+media](https://www.elsevier.com/authors/policies-and-guidelines/artwork-and-media-instructions).
+The exported figures in TIFF format exceeded 100 MB per file and are not
+tracked in this repository. The automated exported figures from the
+Quarto markdown file [are accessible in the
+repository](https://github.com/Global-Health-Engineering/manuscript-qechairquality/tree/main/manuscript/manuscript_files/figure-pdf).
+
+## Software
 
@@
 
-The Quarto template for Elsevier journals was used to prepared the final
+The [Quarto template for Elsevier journals](https://github.com/quarto-journals/elsevier) was used to prepared the final
 submitted PDF document.
 
-Figures were prepared using the [Elsevier guidelines for artwork and
-media](https://www.elsevier.com/authors/policies-and-guidelines/artwork-and-media-instructions).
-The exported figures in TIFF format exceeded 100 MB per file and are not
-tracked in this repository. The automated exported figures from the
-Quarto markdown file [are accessible in the
-repository](https://github.com/Global-Health-Engineering/manuscript-qechairquality/tree/main/manuscript/manuscript_files/figure-pdf).
-
-The data underlying the tables and figures of this manuscript are
-contained in:
-
-\[TODO: Add link to data\]
-
 ## References

### 2023-04-18 manuscript/manuscript.qmd
Add HTML
Add reference

@@
 document that contains the analysis code and the written narrative of
-the manuscript \[TODO: manuscript software citation\].
+the manuscript [@schobitz2023manuscript].
 
@@
 percent_exposure |> 
-  write_csv("data/processed-data/fig05-percent-exposure.csv")
+  write_csv(here::here("data/processed-data/fig05-percent-exposure.csv"))
 

### 2023-05-19 manuscript/manuscript.qmd
Fix citation

@@
 at the hospital's incinerator, located at a central point within the
-hospital campus[^1] [@kalina2020this]. However, the vast majority of the
+hospital campus[^1] [@tilley2020wea]. However, the vast majority of the
 deposited waste is not incinerated, as a new incinerator provided by the
@@
 [^4]: ID5, 2019-11-18
-
+x
 Challenges with solid waste management (SWM) are not unique to Queen's,

### 2023-05-19 manuscript/manuscript.qmd
render

--- 2023-04-18 manuscript/manuscript.qmd
@@
 
-[^3]: Patients are required to bring a caregiver for in-patient
-    hospitals stays, usually a family member, who assist with feeding,
-    bathing, and all other non-medical services.
+[^3]: Patients are required to bring a caregiver for in-patient hospital
+    stays, usually a family member, who assists with feeding, bathing,
+    and all other non-medical services.
 
@@
 pollution within urban areas [@amegah2018proliferation;
-@bulto2020impact]. Although open trashing burning is a problem in LICs
+@bulto2020impact]. Although open trash burning is a problem in LICs
 broadly, for instance, @wiedinmyer2014global estimate that open trash
@@
     concrete floor shelter for sleeping along with a cooking pavilion
-    and toilet/shower block are maintained by a the local NGO Chira
+    and toilet/shower block that are maintained by a local NGO, Chira
     fund. Though simple, the facility is secure and offers one of the
@@
 these shortages have manifested in poor salaries, missed pay cheques,
-failing infrastructure and limited supplies [@kalina2020this]. Regarding
+failing infrastructure and limited supplies [@tilley2020wea]. Regarding
 waste management, these shortages have manifested in a chronic shortage
@@
 including accidental pokes from improperly disposed of sharps, common
-amongst hospital staff. Regardless, separation ends at the point of
-collection. The hospital is not serviced by regular municipal waste
-collection, and as a result, nearly all of the waste produced by the
-hospital, hazardous and non-hazardous, has to be disposed of on-site,
-either within the limited capacity of the hospital incinerator, or more
-commonly, through open burning. As a result, the constantly smouldering
-pile of waste puts forth a continuous cloud of grey smoke, which
-mingling with the dozens of other fires on the grounds, from cooking
-fires and burning garden waste, blankets Queen's in a permanent cloud of
-foul smelling haze.
+amongst hospital staff [@kalina2020this]. Regardless, separation ends at
+the point of collection. The hospital is not serviced by regular
+municipal waste collection, and as a result, nearly all of the waste
+produced by the hospital, hazardous and non-hazardous, has to be
+disposed of on-site, either within the limited capacity of the hospital
+incinerator, or more commonly, through open burning. As a result, the
+constantly smouldering pile of waste puts forth a continuous cloud of
+grey smoke, which mingling with the dozens of other fires on the
+grounds, from cooking fires and burning garden waste, blankets Queen's
+in a permanent cloud of foul smelling haze.
 
@@
 Precision Laser sensor. This sensor measures particles at 2.5 and 10 µm
-in diameter in mg/m^3^. A list of the specific hardware and software
+in diameter in µg/m^3^. A list of the specific hardware and software
 components can be found in Appendix B.
@@
 approximately two metres from the ground with the help of QECH
-maintenance department personnel. This height was chosen to prevent the
-public from tampering with the equipment when they visit the hospital.
-Additionally, we assured that each pipe that inhaled ambient air was
-freely protruding in the building (room) in order to capture the air
-quality.
+maintenance department personnel. TThis height was chosen to both
+capture the approximate breathing zone and to prevent the public from
+tampering with the equipment when they visit the hospital. Additionally,
+we assured that each pipe that inhaled ambient air was freely protruding
+in the building (room) in order to capture the airquality.
 
@@
 Knowing the exact burning locations and burning times, especially at the
-incinerator and the Guardian Shelter, would have helped in better
-identifying the sources and movements of the plumes. However, given the
-size of the campus and it's 24-hour schedule, a much larger research
-team would have been required to quantify all the burning. Finally, The
-humidity, and the potential for interference was not accounted for; the
+incinerator and the cooking times at the Guardian Shelter (though it is
+nearly constant), would have helped in better identifying the sources
+and movements of the plumes. However, given the size of the campus and
+its 24-hour schedule, a much larger research team would have been
+required to quantify all the burning. Finally, the relative humidity,
+and the potential for interference was not accounted for; the
 measurements were not adjusted for humidity which could affect the
@@
 ```{r}
-#| label: tbl01-peaks
+#| label: tbl-peaks
 #| tbl-cap: Peaks for both PM~2.5~ and PM~10~ are the number of data points above the WHO interim target 1 (annual) which is the least stringent; the number of observations recorded (count) is provided for reference.
@@
 characteristic of incomplete combustion and dust that is typical for the
-area and the open burning that occurs there. The residents at the
-Guardian Shelter however are mostly cooking within contained clay stoves
-that have been designed to improve combustion of the coal and wood that
-is burned within them. The fewer number of PM~10~ peaks is a testament
-to this intervention, though the frequency of PM~2.5~ peaks is still
-beyond acceptable.
+area and the open burning that occurs there. Although there is trash
+burning at the Guardian Shelter, the main source of smoke is from the
+constant cooking. The residents at the Guardian Shelter are mostly
+cooking coal and wood that is burned within small "improved" cooking
+stoves that are contained, in a row, within a dedicated cooking block.
+The fewer number of PM10 peaks is a testament to this intervention,
+though the frequency of PM2.5 peaks is still beyond acceptable.
 
@@
 The 24-hour averaged values (logged every 5 minutes) are presented for
-each location, for both PM~10~ and PM~2.5~ in fig-24-hour-average. As
+each location, for both PM~10~ and PM~2.5~ in @fig-24-hour-average. As
 well as dampening local variation and peaks, the WHO air quality
@@
 Shelter are both consistently higher for both parameters and the two
-averages closely followed the same general trends. Though the peaks at
-the Lighthouse Clinic were higher than the Guardian Shelter, the low
-values were consistently lower, indicating more times of little or no
-burning, unlike the Guardian Shelter emissions which were relatively
-constant. However, further analysis of 12 hour averages (8:00-15:59
-(working hours) and 16:00-7:59 (evening)) did not indicate clear
-differences between the time periods; stated differently, the data did
-not clearly point to more or less burning in the day or at night.
+averages closely follow the same general trends. Though the peaks at the
+Lighthouse Clinic were higher than the Guardian Shelter, the low values
+were consistently lower, indicating more times of little or no burning,
+unlike the Guardian Shelter emissions which were relatively constant.
+However, further analysis of 12 hour averages (8:00-15:59 (working
+hours) and 16:00-7:59 (evening)) did not indicate clear differences
+between the time periods; stated differently, the data did not clearly
+point to more or less burning in the day or at night.
 
 ```{r}
-#| label: fig03-24-hour-average
+#| label: fig-24-hour-average
 #| fig-cap: Average 24-hour PM~10~ and PM~2.5~ at 8 monitoring stations over 2 months
@@
 ```{r}
-#| label: fig04-pm-ratios
+#| label: fig-pm-ratios
 #| fig-cap: PM~10~:PM~2.5~ values for each location.
@@
 ```{r}
-#| label: fig05-percent-exposure
+#| label: fig-percent-exposure
 #| fig-cap: Percentage of measured values categorised according to WHO 2021 targets
@@
 Unsurprisingly, a majority of the air quality at the Lighthouse Clinic
-and the Guardian Shelter falls at, or below, interim target 1. What is
+and the Guardian Shelter falls at, or above, interim target 1. What is
 less expected is that even though there are fewer days of hazardous or
@@
 Project and Lions Sight have the largest percentages of days with air
-quality meeting the interim target 4. Interim target 2 rated air
-occupies an unexpectedly large proportion of the days at all locations,
-indicating that although there are a few very toxic, and a few quite
-good days, the majority of the air is actually neither, but still cause
-for concern.
+quality meeting the interim target 4, which is expected as these
+locations are further away from the main burning site. Interim target 2
+rated air occupies an unexpectedly large proportion of the days at all
+locations, indicating that although there are a few very toxic, and a
+few quite good days, the majority of the air is actually neither, but
+still cause for concern.
 
@@
 on; coping as best they can. As the same member of the janitorial
-staff[^11] described, "we are just hoping we will start breathing good
+staff[^12] described, "we are just hoping we will start breathing good
 air soon, when the new incinerator is opened."
 
-[^11]: ID14, 2019-11-23
+[^12]: ID14, 2019-11-23
 
@@
 recorded more than 50 instances above the most hazardous limit. The
+results for the Lighthouse Clinic are directly related to the main trash
+burning site, while the worrisome air quality at the Guardian Shelter is
+primarily from near constant cooking with coal. Simple improvements to
+the cooking block (e.g. Whirlybird vents) could at least vent the smoke
+above the roof and improve the air quality at human level. The
 composition of emissions recorded ranged significantly, analysed as the
 ratio PM~10~:PM~2.5~ likely reflects a range of burning styles and/or
-fuel (waste) type, and no discernible trend was observed.. Among
+fuel (waste) type, and no discernible trend was observed. Among
 interview respondents, there was a general consensus that air quality
@@
 circumstances. Moreover, we would like to acknowledge the assistance and
-effort of Open Data Durban, now the Open Cities Lab, for their
+effort of the Open Cities Lab (formerly Open Data Durban), for their
 assistance in constructing and programming the sensors.
@@
 ```{r}
-#| label: figA1-five-minute-intervals
+#| label: fig-five-minute-intervals
 #| fig-cap: PM~2.5~ and PM~10~ values collected every 5 minutes over 10 months at 8 locations.
@@
     algorithms to convert between AQI value and pollutant
-    concentration). The code is stored on the [Open Data Durban GitHub
-    account](https://github.com/opendatadurban/hospital_stations) and is
-    free for all to access. To be able to collect data, the device needs
-    to be connected to a reliable wifi network, the SSID and password
-    need to be coded onto the device during step 2 in the building
-    method.
+    concentration). To be able to collect data, the device needs to be
+    connected to a reliable wifi network, the SSID and password need to
+    be coded onto the device during step 2 in the building method.
 3.  Once the whole device has been put together and placed on site, to

### 2023-05-19 manuscript/manuscript.tex
render

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
 \fi
@@
 \makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
 \makeatother
@@
 \@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
 \makeatother
@@
 }
- \ead{kwangerlo@gmail.comm} 
+ \ead{kwangulero@gmail.com} 
 \author[1]{Lars Schöbitz%
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, breakable, sharp corners]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, enhanced, sharp corners, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 hospital campus\footnote{Sharps are handled and disposed of separately.}
-\citep{kalina2020this}. However, the vast majority of the deposited
-waste is not incinerated, as a new incinerator provided by the Ministry
-of Health in late 2019, is only able to handle a small percentage of the
+\citep{tilley2020wea}. However, the vast majority of the deposited waste
+is not incinerated, as a new incinerator provided by the Ministry of
+Health in late 2019, is only able to handle a small percentage of the
 hospital's waste. Furthermore, the incinerator is frequently inoperable,
@@
 leaves and other garden refuse. Furthermore, caregivers\footnote{Patients
-  are required to bring a caregiver for in-patient hospitals stays,
-  usually a family member, who assist with feeding, bathing, and all
+  are required to bring a caregiver for in-patient hospital stays,
+  usually a family member, who assists with feeding, bathing, and all
   other non-medical services.} who reside separately on hospital
@@
 pollution within urban areas
-\citep{amegah2018proliferation, bulto2020impact}. Although open trashing
+\citep{amegah2018proliferation, bulto2020impact}. Although open trash
 burning is a problem in LICs broadly, for instance,
@@
   gender-separated concrete floor shelter for sleeping along with a
-  cooking pavilion and toilet/shower block are maintained by a the local
-  NGO Chira fund. Though simple, the facility is secure and offers one
-  of the only free/accessible toilet facilities for visitors on the QECH
-  campus.
+  cooking pavilion and toilet/shower block that are maintained by a
+  local NGO, Chira fund. Though simple, the facility is secure and
+  offers one of the only free/accessible toilet facilities for visitors
+  on the QECH campus.
 \end{itemize}
@@
 these shortages have manifested in poor salaries, missed pay cheques,
-failing infrastructure and limited supplies \citep{kalina2020this}.
+failing infrastructure and limited supplies \citep{tilley2020wea}.
 Regarding waste management, these shortages have manifested in a chronic
@@
 infected waste, including accidental pokes from improperly disposed of
-sharps, common amongst hospital staff. Regardless, separation ends at
-the point of collection. The hospital is not serviced by regular
-municipal waste collection, and as a result, nearly all of the waste
-produced by the hospital, hazardous and non-hazardous, has to be
-disposed of on-site, either within the limited capacity of the hospital
-incinerator, or more commonly, through open burning. As a result, the
-constantly smouldering pile of waste puts forth a continuous cloud of
-grey smoke, which mingling with the dozens of other fires on the
-grounds, from cooking fires and burning garden waste, blankets Queen's
-in a permanent cloud of foul smelling haze.
+sharps, common amongst hospital staff \citep{kalina2020this}.
+Regardless, separation ends at the point of collection. The hospital is
+not serviced by regular municipal waste collection, and as a result,
+nearly all of the waste produced by the hospital, hazardous and
+non-hazardous, has to be disposed of on-site, either within the limited
+capacity of the hospital incinerator, or more commonly, through open
+burning. As a result, the constantly smouldering pile of waste puts
+forth a continuous cloud of grey smoke, which mingling with the dozens
+of other fires on the grounds, from cooking fires and burning garden
+waste, blankets Queen's in a permanent cloud of foul smelling haze.
 
@@
 Precision Laser sensor. This sensor measures particles at 2.5 and 10 µm
-in diameter in mg/m\textsuperscript{3}. A list of the specific hardware
+in diameter in µg/m\textsuperscript{3}. A list of the specific hardware
 and software components can be found in Appendix B.
@@
 reproducible document that contains the analysis code and the written
-narrative of the manuscript {[}TODO: manuscript software citation{]}.
+narrative of the manuscript \citep{schobitz2023manuscript}.
 
@@
 observations recorded (count) is provided for reference.}\tabularnewline
-\toprule()
+\toprule\noalign{}
 location & n & pm10 & pm2.5 \\
-\midrule()
+\midrule\noalign{}
 \endfirsthead
-\toprule()
+\toprule\noalign{}
 location & n & pm10 & pm2.5 \\
-\midrule()
+\midrule\noalign{}
 \endhead
+\bottomrule\noalign{}
+\endlastfoot
 Administration & 10900 & 171 & 0 \\
@@
 uMoyo & 10637 & 703 & 0 \\
-\bottomrule()
 \end{longtable}
@@
 and dust that is typical for the area and the open burning that occurs
-there. The residents at the Guardian Shelter however are mostly cooking
-within contained clay stoves that have been designed to improve
-combustion of the coal and wood that is burned within them. The fewer
-number of PM\textsubscript{10} peaks is a testament to this
-intervention, though the frequency of PM\textsubscript{2.5} peaks is
-still beyond acceptable.
+there. Although there is trash burning at the Guardian Shelter, the main
+source of smoke is from the constant cooking. The residents at the
+Guardian Shelter are mostly cooking coal and wood that is burned within
+small ``improved'' cooking stoves that are contained, in a row, within a
+dedicated cooking block. The fewer number of PM10 peaks is a testament
+to this intervention, though the frequency of PM2.5 peaks is still
+beyond acceptable.
 
@@
 each location, for both PM\textsubscript{10} and PM\textsubscript{2.5}
-in fig-24-hour-average. As well as dampening local variation and peaks,
-the WHO air quality guidelines are also based on 24-hour averages which
-are the standard against which the health risks can be judged.
+in Figure~\ref{fig-24-hour-average}. As well as dampening local
+variation and peaks, the WHO air quality guidelines are also based on
+24-hour averages which are the standard against which the health risks
+can be judged.
 
@@
 Shelter are both consistently higher for both parameters and the two
-averages closely followed the same general trends. Though the peaks at
-the Lighthouse Clinic were higher than the Guardian Shelter, the low
-values were consistently lower, indicating more times of little or no
-burning, unlike the Guardian Shelter emissions which were relatively
-constant. However, further analysis of 12 hour averages (8:00-15:59
-(working hours) and 16:00-7:59 (evening)) did not indicate clear
-differences between the time periods; stated differently, the data did
-not clearly point to more or less burning in the day or at night.
+averages closely follow the same general trends. Though the peaks at the
+Lighthouse Clinic were higher than the Guardian Shelter, the low values
+were consistently lower, indicating more times of little or no burning,
+unlike the Guardian Shelter emissions which were relatively constant.
+However, further analysis of 12 hour averages (8:00-15:59 (working
+hours) and 16:00-7:59 (evening)) did not indicate clear differences
+between the time periods; stated differently, the data did not clearly
+point to more or less burning in the day or at night.
 
@@
 recorded more than 50 instances above the most hazardous limit. The
+results for the Lighthouse Clinic are directly related to the main trash
+burning site, while the worrisome air quality at the Guardian Shelter is
+primarily from near constant cooking with coal. Simple improvements to
+the cooking block (e.g.~Whirlybird vents) could at least vent the smoke
+above the roof and improve the air quality at human level. The
 composition of emissions recorded ranged significantly, analysed as the
@@
 of burning styles and/or fuel (waste) type, and no discernible trend was
-observed.. Among interview respondents, there was a general consensus
+observed. Among interview respondents, there was a general consensus
 that air quality caused by the burning of waste within the hospital was
@@
   algorithms to convert between AQI value and pollutant concentration).
-  The code is stored on the
-  \href{https://github.com/opendatadurban/hospital_stations}{Open Data
-  Durban GitHub account} and is free for all to access. To be able to
-  collect data, the device needs to be connected to a reliable wifi
-  network, the SSID and password need to be coded onto the device during
-  step 2 in the building method.
+  To be able to collect data, the device needs to be connected to a
+  reliable wifi network, the SSID and password need to be coded onto the
+  device during step 2 in the building method.
 \item

### 2023-05-19 R/01-data-preparation.R
Reproduce targets table

@@ -6,6 +6,8 @@
 
 library(readr)
 library(dplyr)
+library(tabulizer)
+library(stringr)
 library(forcats)
 library(lubridate)
 library(ggplot2)
@@ -168,6 +170,40 @@ dat_air_quality_sum_who_2021 <- dat_air_quality |>
   mutate(exposure_hrs = round(exposure_mins / 60)) |>
   select(-n)
 
-# data export -------------------------------------------------------------
+
+# extract table from WHO pdf ----------------------------------------------
+
+
+
+
+who_df <- extract_tables(file = "data/raw-data/9789240034228-eng.pdf", pages = 159)
+
+who_target_tidy <- who_df[[1]] |>
+  as_tibble() |>
+  separate(V4, into = c("interim_target2", "interim_target3"),
+           sep = " ", remove = TRUE) |>
+  rename(
+    pollutant = V1,
+    averageing_time = V2,
+    interim_target1 = V3,
+    interim_target4 = V5,
+    AQG_level = V6
+  ) |>
+  slice(-3:-1) |>
+  mutate_if(is.character, na_if, "") |>
+  mutate_if(is.character, na_if, "–") |>
+  fill(pollutant, .direction = "down") |>#pull(pollutant)
+  mutate(pollutant = str_replace(pollutant, "SO , μg/m32", "SO, μg/m3")) |>
+  mutate(pollutant = str_replace(pollutant, "μg", "µg")) |>
+  mutate(pollutant = str_replace(pollutant, "m 3", "m3")) |>
+  mutate(averageing_time = case_when(
+    averageing_time == "24-houra" ~ "24-hour",
+    averageing_time == "Peak seasonb" ~ "Peak season",
+    averageing_time == "8-houra" ~ "8-hour",
+    TRUE ~ averageing_time
+  )) |>
+  mutate(across(interim_target1:AQG_level, as.numeric))
+
+
 
 

### 2023-05-19 R/01-data-preparation.R
Store as long table

@@ -202,8 +202,14 @@ who_target_tidy <- who_df[[1]] |>
     averageing_time == "8-houra" ~ "8-hour",
     TRUE ~ averageing_time
   )) |>
-  mutate(across(interim_target1:AQG_level, as.numeric))
+  mutate(across(interim_target1:AQG_level, as.numeric)) |>
+  pivot_longer(cols = interim_target1:interim_target4,
+               values_to = "concentration",
+               names_to = "interim_target") |>
+  mutate(interim_target = str_replace(interim_target, "interim_target", ""))
 
+who_target_tidy |>
+  write_csv("data/processed-data/who-air-quality-guidelines-2021-targets-table.csv")
 
 
 

### 2023-05-19 R/00-setup.R
Add gt

@@ -6,6 +6,7 @@
 
 library(ggplot2)
 library(scales)
+library(gt)
 
 # theme adapted from:
 # https://github.com/teunbrand/ggplot_tricks

### 2023-05-19 R/01-data-preparation.R

@@ -184,7 +184,7 @@ who_target_tidy <- who_df[[1]] |>
            sep = " ", remove = TRUE) |>
   rename(
     pollutant = V1,
-    averageing_time = V2,
+    averaging_time = V2,
     interim_target1 = V3,
     interim_target4 = V5,
     AQG_level = V6
@@ -196,11 +196,12 @@ who_target_tidy <- who_df[[1]] |>
   mutate(pollutant = str_replace(pollutant, "SO , μg/m32", "SO, μg/m3")) |>
   mutate(pollutant = str_replace(pollutant, "μg", "µg")) |>
   mutate(pollutant = str_replace(pollutant, "m 3", "m3")) |>
-  mutate(averageing_time = case_when(
-    averageing_time == "24-houra" ~ "24-hour",
-    averageing_time == "Peak seasonb" ~ "Peak season",
-    averageing_time == "8-houra" ~ "8-hour",
-    TRUE ~ averageing_time
+  mutate(pollutant = str_replace(pollutant, "PM 10", "PM10")) |>
+  mutate(averaging_time = case_when(
+    averaging_time == "24-houra" ~ "24-hour",
+    averaging_time == "Peak seasonb" ~ "Peak season",
+    averaging_time == "8-houra" ~ "8-hour",
+    TRUE ~ averaging_time
   )) |>
   mutate(across(interim_target1:AQG_level, as.numeric)) |>
   pivot_longer(cols = interim_target1:interim_target4,
@@ -208,8 +209,12 @@ who_target_tidy <- who_df[[1]] |>
                names_to = "interim_target") |>
   mutate(interim_target = str_replace(interim_target, "interim_target", ""))
 
+
 who_target_tidy |>
   write_csv("data/processed-data/who-air-quality-guidelines-2021-targets-table.csv")
+
+
+
 
 
 

### 2023-05-19 manuscript/manuscript.tex
Add WHO targets table

@@
 
+\usepackage{booktabs}
+\usepackage{caption}
+\usepackage{longtable}
 \makeatletter
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, enhanced, sharp corners, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, enhanced, sharp corners, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 \citep{worldhealthorganization2021who}. The relevant particulate matter
-targets are presented in \textbf{?@tbl-who-targets}.
-
-(TODO: REPLACE screenshot with actual data)
-
-\begin{table}
-
-\caption{\textbf{?(caption)}}\begin{minipage}[t]{\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics[width=2.04in,height=\textheight]{img/image-replace-table.png}
-
-}
-
-}
-
-\end{minipage}%
-
-\end{table}
+targets are presented in Table~\ref{tbl-who-targets}.
+
+\hypertarget{tbl-who-targets}{}
+\setlength{\LTpost}{0mm}
+\begin{longtable}{l|rrrrr}
+\caption{\label{tbl-who-targets}Recommended long- and short-term AQG (Air Quality Guidelines) levels and
+interim targets. }\tabularnewline
+
+\toprule
+\multicolumn{1}{l}{} & \multicolumn{4}{c}{\textbf{Interim target}} &  \\ 
+\cmidrule(lr){2-5}
+\multicolumn{1}{l}{} & 1 & 2 & 3 & 4 & AQG level \\ 
+\midrule
+\multicolumn{6}{l}{Annual} \\ 
+\midrule
+PM$_{2.5}$, µg/m$_{3}$ & $35.0$ & $25.0$ & $15.0$ & $10.0$ & $5.0$ \\ 
+PM$_{10}$, µg/m$_{3}$ & $70.0$ & $50.0$ & $30.0$ & $20.0$ & $15.0$ \\ 
+\midrule
+\multicolumn{6}{l}{24-hour\textsuperscript{\textit{1}}} \\ 
+\midrule
+PM$_{2.5}$, µg/m$_{3}$ & $75.0$ & $50.0$ & $37.5$ & $25.0$ & $15.0$ \\ 
+PM$_{10}$, µg/m$_{3}$ & $150.0$ & $100.0$ & $75.0$ & $50.0$ & $45.0$ \\ 
+\bottomrule
+\end{longtable}
+\begin{minipage}{\linewidth}
+\textsuperscript{\textit{1}}99th percentile (i.e. 3-4 exceedance days per year).\\
+\end{minipage}
 

### 2023-05-19 R/01-data-preparation.R

@@ -173,10 +173,7 @@ dat_air_quality_sum_who_2021 <- dat_air_quality |>
 
 # extract table from WHO pdf ----------------------------------------------
 
-
-
-
-who_df <- extract_tables(file = "data/raw-data/9789240034228-eng.pdf", pages = 159)
+who_df <- extract_tables(file = here::here("data/raw-data/9789240034228-eng.pdf"), pages = 159)
 
 who_target_tidy <- who_df[[1]] |>
   as_tibble() |>
@@ -211,8 +208,7 @@ who_target_tidy <- who_df[[1]] |>
 
 
 who_target_tidy |>
-  write_csv("data/processed-data/who-air-quality-guidelines-2021-targets-table.csv")
-
+  write_csv(here::here("data/processed-data/who-air-quality-guidelines-2021-targets-table.csv"))
 
 
 

### 2023-05-19 manuscript/manuscript-qech-air-quality-tilley.qmd
Rename file

--- 2023-05-19 manuscript/manuscript.qmd
@@
 
-(TODO: REPLACE screenshot with actual data)
-
 ```{r}
 #| label: tbl-who-targets
-#| tbl-cap: Particulate matter targets from [@worldhealthorganization2021who]
-
-knitr::include_graphics(here::here("manuscript/img/image-replace-table.png"))
+#| tbl-cap: Recommended long- and short-term AQG (Air Quality Guidelines) levels and interim targets.
+
+who_target_tidy |>
+  pivot_wider(names_from = "interim_target",
+              values_from = "concentration") |>
+  filter(str_detect(pollutant, "PM")) |>
+  relocate(AQG_level, .after = "4") |>
+  group_by(averaging_time) |>
+  # prepare table
+  gt(rowname_col = "pollutant") |>
+  tab_spanner(label = md("**Interim target**"), columns = `1`:`4`) |>
+  cols_label(
+    averaging_time = "Averaging time",
+    AQG_level = "AQG level"
+  ) |>
+  fmt_number(columns = `1`:AQG_level, decimals = 1) |>
+  tab_style(
+    locations = cells_column_labels(),
+    style = list(cell_text(weight = "bold"))
+  ) |>
+  tab_style(locations = cells_row_groups(),
+            style = list(cell_text(weight = "bold"))) |>
+  text_transform(locations = cells_stub(),
+                 fn = function(x) { 
+                   x |> 
+                     str_replace_all("2.5", "$_{2.5}$") |> 
+                     str_replace_all("3", "$_{3}$") |> 
+                     str_replace_all("10", "$_{10}$") }
+                 ) |> 
+  tab_footnote(
+    footnote = "99th percentile (i.e. 3-4 exceedance days per year).",
+    locations = cells_row_groups(group = "24-hour")
+  )
 

### 2023-05-19 manuscript/manuscript-qech-air-quality-tilley.tex
Rename file

--- 2023-05-19 manuscript/manuscript.tex
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, enhanced, sharp corners, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, interior hidden, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi
 
@@
 
-{\centering \includegraphics{manuscript_files/figure-pdf/fig-24-hour-average-1.pdf}
+{\centering \includegraphics{manuscript-qech-air-quality-tilley_files/figure-pdf/fig-24-hour-average-1.pdf}
 
@@
 
-{\centering \includegraphics{manuscript_files/figure-pdf/fig-pm-ratios-1.pdf}
+{\centering \includegraphics{manuscript-qech-air-quality-tilley_files/figure-pdf/fig-pm-ratios-1.pdf}
 
@@
 
-{\centering \includegraphics{manuscript_files/figure-pdf/fig-percent-exposure-1.pdf}
+{\centering \includegraphics{manuscript-qech-air-quality-tilley_files/figure-pdf/fig-percent-exposure-1.pdf}
 
@@
 
-{\centering \includegraphics{manuscript_files/figure-pdf/fig-five-minute-intervals-1.pdf}
+{\centering \includegraphics{manuscript-qech-air-quality-tilley_files/figure-pdf/fig-five-minute-intervals-1.pdf}
 

### 2023-05-19 manuscript/latex/manuscript-qech-air-quality-tilley.tex
Submitted manuscript

--- 2023-05-19 manuscript/manuscript-qech-air-quality-tilley.tex
@@
 \makeatother
-\journal{Habitat International}
+\journal{Habitat International (Declarations of interest: none.)}
 \ifLuaTeX
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, interior hidden, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, enhanced, sharp corners]}{\end{tcolorbox}}\fi
 

### 2023-05-19 manuscript/manuscript-qech-air-quality-tilley-credit.qmd
Submitted manuscript

---
title: "'It Is Unbearable To Breath Here': Air Quality, Open Incineration, And Misinformation In Blantyre, Malawi"
author:
  - name: Elizabeth Tilley
    orcid: 0000-0002-2095-9724
    email: tilleye@ethz.ch
    affiliations: 
        - id: ethz
          name: ETH Zurich
          department: Department of Mechanical and Process Engineering
          address: Global Health Engineering, Clausiusstrasse 37
          city: Zurich
          country: Switzerland
          postal-code: 8092
    attributes:
        corresponding: true
  - name: Hope Chilunga
    orcid: 0000-0003-4216-6319
    email: hchilunga@poly.ac.mw
    affiliations: 
        - id: mubas
          name: Malawi University of Business and Applied Sciences
          department: Department of Environmental Health
          city: Blantyre
          country: Malawi
  - name: Jonathan Kwangulero
    orcid: 0000-0002-4050-5121
    email: kwangulero@gmail.com 
    affilations:
        - ref: mubas
  - name: Lars Schöbitz
    orcid: 0000-0003-2196-5015
    email: lschoebitz@ethz.ch
    affiliations:
        - ref: ethz
  - name: Saloni Vijay
    orcid: 0000-0001-5883-5308
    email: svijay@ethz.ch
    affiliations:
        - ref: ethz
  - name: Heiko Heilgendorff
    orcid: 0009-0004-1389-5541
    email: heiko.heilgendorff@gmail.com
    affiliations: 
        - id: ukzn
          name: University of KwaZulu-Natal
          department: School of Mathematics, Statistics and Computer Science
          city: Durban
          country: South Africa
  - name: Marc Kalina
    orcid: 0000-0002-6335-3845
    email: mkalina@ethz.ch
    affiliations:
        - ref: ethz
date: last-modified
bibliography: [man-qechairquality.bib, packages.bib]
format:
  docx: default
execute: 
  echo: false
  warning: false
  freeze: true
editor: 
  markdown: 
    wrap: 72
editor_options: 
  chunk_output_type: console
---

# CRediT roles

```{r}

library(tidyverse)

df_credit <- read.csv("https://raw.githubusercontent.com/Global-Health-Engineering/manuscript-qechairquality/main/data/credit.csv") |> 
  unite("name", first_name:surname, sep = "")

```

```{r}
credit_sentence <- larnsceR::credit_sentence(data = df_credit, group_var = name)
```

`r credit_sentence$sentence`

### 2023-05-25 manuscript/manuscript-qech-air-quality-tilley.qmd
Submitted manuscript
Update email
Add line numbers, remove highlights

@@
     orcid: 0009-0004-1389-5541
-    email: heiko.heilgendorff@gmail.com
+    email: heiko@opencitieslab.org
     affiliations: 
@@
   elsevier-pdf:
+    include-in-header: 
+      text: '\usepackage{lineno}\linenumbers'
     keep-tex: true
     journal:
-      name: Habitat International
+      name: "Environment International"
       formatting: review
@@
       cite-style: authoryear
-      highlights:
-        - Low-cost monitors were used to assess air quality at Malawi’s largest hospital
-        - Open waste burning is physically extensive and temporally frequent across the hospital campus
-        - Air quality measured as PM~2.5~ and PM~10~ consistently exceeded even the most hazardous levels 
-        - The most exposed locations were where guardians of patients stay and the HIV clinic
 execute: 

### 2023-05-25 manuscript/manuscript-qech-air-quality-tilley.tex
Add line numbers, remove highlights

@@
 \usepackage{longtable}
+\usepackage{lineno}\linenumbers
 \makeatletter
@@
 \makeatother
-\journal{Habitat International}
+\journal{Environment International}
 \ifLuaTeX
@@
 }
- \ead{heiko.heilgendorff@gmail.com} 
+ \ead{heiko@opencitieslab.org} 
 \author[1]{Marc Kalina%
@@
 
-\begin{highlights}
-\item Low-cost monitors were used to assess air quality at Malawi's
-largest hospital\item Open waste burning is physically extensive and
-temporally frequent across the hospital campus\item Air quality measured
-as PM\textsubscript{2.5} and PM\textsubscript{10} consistently exceeded
-even the most hazardous levels\item The most exposed locations were
-where guardians of patients stay and the HIV clinic
-\end{highlights}
-
 \begin{keyword}
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, interior hidden, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, breakable, frame hidden, sharp corners, interior hidden]}{\end{tcolorbox}}\fi
 

### 2023-05-25 manuscript/manuscript-qech-air-quality-tilley.tex
Fix typo

@@
 \hypersetup{
-  pdftitle={`It Is Unbearable To Breath Here': Air Quality, Open Incineration, And Misinformation In Blantyre, Malawi},
+  pdftitle={`It Is Unbearable To Breathe Here': Air Quality, Open Incineration, And Misinformation In Blantyre, Malawi},
   pdfauthor={Elizabeth Tilley; Hope Chilunga; Jonathan Kwangulero; Lars Schöbitz; Saloni Vijay; Heiko Heilgendorff; Marc Kalina},
@@
 \begin{frontmatter}
-\title{`It Is Unbearable To Breath Here': Air Quality, Open
+\title{`It Is Unbearable To Breathe Here': Air Quality, Open
 Incineration, And Misinformation In Blantyre, Malawi}
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, breakable, frame hidden, sharp corners, interior hidden]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, interior hidden, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
 

### 2023-08-23 manuscript/manuscript-qech-air-quality-tilley.tex
Update figures to follow author guidelines

@@
 \makeatletter
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi
+\makeatother
+\makeatletter
 \makeatother
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, interior hidden, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
-
     \hypertarget{introduction}{%

### 2023-08-23 R/00-setup.R

@@ -16,7 +16,7 @@ axis_title_just = "rt"
 xj <- switch(tolower(substr(axis_title_just, 1, 1)), b=0, l=0, m=0.5, c=0.5, r=1, t=1)
 yj <- switch(tolower(substr(axis_title_just, 2, 2)), b=0, l=0, m=0.5, c=0.5, r=1, t=1)
 
-base_size = 7
+base_size = 8
 axis_title_size = 6
 
 theme_set(
@@ -41,10 +41,10 @@ theme_set(
 
 # paramaters for ggsave
 
-param_dpi = 1000
-param_width_single = 90
-param_width_double = 190
-width_height_ratio = 190/275
+param_dpi = 300
+param_width_single = 85
+param_width_double = 180
+width_height_ratio = 180/275
 
 # test
 

### 2023-09-26 manuscript/manuscript-qech-air-quality-tilley.qmd
Fix typo
Fix issue with table

@@
 ---
-title: "'It Is Unbearable To Breath Here': Air Quality, Open Incineration, And Misinformation In Blantyre, Malawi"
+title: "'It Is Unbearable To Breathe Here': Air Quality, Open Incineration, And Misinformation In Blantyre, Malawi"
 author:
@@
 
-no_peaks_who <- dat_in |> 
+no_peaks_who
+
+dat_in |> 
   mutate(peaks = case_when(
-    indicator == "pm25" & value >  35 ~ TRUE,
+    indicator == "pm2.5" & value >  35 ~ TRUE,
     indicator == "pm10" & value >  70 ~ TRUE,

### 2023-09-26 manuscript/manuscript-qech-air-quality-tilley.tex
Fix issue with table

@@
 \makeatletter
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, sharp corners, frame hidden, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt]}{\end{tcolorbox}}\fi
 \makeatother

### 2023-10-05 manuscript/manuscript-qech-air-quality-tilley.qmd
Show figure facet label in subscript

@@
 
-no_peaks_who
-
-dat_in |> 
+no_peaks_who <- dat_in |> 
   mutate(peaks = case_when(
@@
 
+# Create a custom labeller function
+custom_labeller <- function(labels) {
+  lapply(labels, function(x) bquote(.(x)))
+}
+
 percent_exposure <- dat_air_quality_sum_who_2021 |> 
@@
     TRUE ~ exposure_percent
-  ))
+  )) |> 
+  mutate(indicator = case_when(
+    indicator == "pm10" ~ "PM[10]",
+    indicator == "pm2.5" ~ "PM[2.5]"
+  )) |> 
+  ungroup()
 
@@
 
-fig05 <- 
-ggplot(data = percent_exposure,
+fig05 <- ggplot(data = percent_exposure,
        aes(fill = air_quality_who_annual, values = exposure_percent)) +
@@
   scale_y_discrete(expand=c(0,0)) +
-  facet_grid(location ~ indicator, switch = "y") +
+  facet_grid(location~indicator, switch = "y", 
+             labeller = labeller(location = custom_labeller, indicator = label_parsed)) +
 
@@
   write_csv(here::here("data/processed-data/fig05-percent-exposure.csv"))
+
+```
+
+```{r}
+# Load required libraries
+library(ggplot2)
+
+# Create example data
+df <- data.frame(
+  x = rep(1:3, each = 10),
+  y = rnorm(30),
+  group = rep(c("A[1]", "B[2]", "C[3]"), each = 10),
+  category = rep(c("Cat 1", "Cat 2"), 15)
+)
+
+# Create a custom labeller function
+custom_labeller <- function(labels) {
+  lapply(labels, function(x) bquote(.(x)))
+}
+
+# Create a ggplot with subscript labels in facets
+p <- ggplot(df, aes(x, y)) +
+  geom_point() +
+  facet_grid(category ~ group, labeller = labeller(category = custom_labeller, group = label_parsed))
+
+# Print the plot
+print(p)
 

### 2023-10-05 manuscript/manuscript-qech-air-quality-tilley.tex
Show figure facet label in subscript

@@
 \makeatletter
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, sharp corners, frame hidden, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, enhanced, frame hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
 \makeatother
@@
 \endlastfoot
-Administration & 10900 & 171 & 0 \\
-Guardian Shelter & 11032 & 4529 & 0 \\
-Lighthouse Clinic & 11472 & 4365 & 0 \\
-Lions Sight & 11749 & 91 & 0 \\
-Malaria & 12031 & 617 & 0 \\
-Mercy James & 12193 & 736 & 0 \\
-Ward 6B & 11861 & 1113 & 0 \\
-uMoyo & 10637 & 703 & 0 \\
+Administration & 10900 & 171 & 94 \\
+Guardian Shelter & 11032 & 4529 & 5314 \\
+Lighthouse Clinic & 11472 & 4365 & 3891 \\
+Lions Sight & 11749 & 91 & 14 \\
+Malaria & 12031 & 617 & 578 \\
+Mercy James & 12193 & 736 & 749 \\
+Ward 6B & 11861 & 1113 & 1236 \\
+uMoyo & 10637 & 703 & 399 \\
 \end{longtable}
@@
 \end{figure}
+
+\includegraphics{manuscript-qech-air-quality-tilley_files/figure-pdf/unnamed-chunk-9-1.pdf}
 
